### 2015-11-21 R/data_explore_pt2.Rmd

Exploring possible correlation between total species range area and percent overlap of smaller range within the larger:

``` {r pct overlap vs area, echo = FALSE, warning = FALSE, message = FALSE}

scatter_area_vs_overlap <- ggplot(spp_list, aes(sm_perc, total_area, color = sm_range)) + 
  ggtheme_plot + 
  geom_point() + 
  geom_abline() + 
  labs(y = 'Total Area (km^2)', 
       x = 'Percent overlap smaller range in larger', 
       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')

lm_eqn <- function(df){
  m  <- lm(total_area ~ sm_perc, spp_list)
  eq <- substitute(italic('y') == a + b %.% italic('x')*","~~italic(r)^2~"="~r2, 
                   list(a  = format(coef(m)[1], digits = 2), 
                        b  = format(coef(m)[2], digits = 2), 
                        r2 = format(summary(m)$r.squared, digits = 3)))
  as.character(as.expression(eq))             
}

scatter_area_vs_overlap <- scatter_area_vs_overlap + 
  geom_text(x = 40, y = 300000000, label = lm_eqn(spp_list), parse = TRUE)

print(scatter_area_vs_overlap)

# summary(lm(total_area ~ sm_perc, spp_list))

ggsave(file.path(dir_fig, 'AreavsPercOverlap.png'), width = 6, height = 4)

```

***

Percent coverage as function of area, clipping to area below 100,000,000 km^2^:

``` {r, echo = FALSE, warning = FALSE, message = FALSE}
# remove all area greater than 100000000

spp_small_df <- spp_list %>% 
  filter(total_area < 100000000)

scatter_a_vs_overlap_small <- ggplot(spp_small_df, aes(total_area, sm_perc, color = sm_range)) + 
  ggtheme_plot + 
  geom_point() + 
  labs(x = 'Total Area (km^2)', 
       y = 'Percent overlap smaller range in larger', 
       title = 'Size of species range vs. percent overlap of \nsmaller range with larger')

ggsave(file.path(dir_fig, 'AreavsPercOverlap_2.png'), width = 6, height = 4)

print(scatter_a_vs_overlap_small)

```

Percent coverage as a function of log(area)
``` {r, echo = FALSE, warning = FALSE, message = FALSE}
# Log transform area

scatter_logA_vs_overlap <- ggplot(spp_list, aes(log_total_area, sm_perc, color = sm_range)) + 
  ggtheme_plot + 
  geom_point() + 
  labs(x = 'Log(total Area)', 
       y = 'Percent overlap smaller range in larger', 
       title = 'Size of species range (log) vs. percent overlap of \nsmaller range with larger')

ggsave(file.path(dir_fig, 'AreavsPercOverlap_log.png'), width=6, height=4)

print(scatter_logA_vs_overlap)
```

``` {r examine regressions, eval = FALSE, echo = FALSE, warning = FALSE}

A_mdl      <- lm(total_area ~ sm_perc, spp_list)
#                Estimate Std. Error t value Pr(>|t|)    
#   (Intercept) -16285410    2858388  -5.697  1.4e-08 ***
#   sm_perc        533370      41702  12.790  < 2e-16 ***
#   Multiple R-squared:  0.07732,	Adjusted R-squared:  0.07685

A_mdl_clip <- lm(total_area ~ sm_perc, l)
#               Estimate Std. Error t value Pr(>|t|)    
#   (Intercept)  2993939     810560   3.694 0.000227 ***
#   sm_perc       153523      11952  12.844  < 2e-16 ***
#   Multiple R-squared:  0.07955,	Adjusted R-squared:  0.07906 
logA_mdl   <- lm(log_total_area ~ sm_perc, spp_list)
#                Estimate Std. Error t value Pr(>|t|)    
#   (Intercept) 14.775325   0.096446   153.2   <2e-16 ***
#   sm_perc      0.016885   0.001407    12.0   <2e-16 ***
#   Multiple R-squared:  0.0687,	Adjusted R-squared:  0.06823 
```

***

###(5) Examining data broken into quartiles to see if smaller patterns emerge
``` {r examine quartiles, eval = FALSE, echo = FALSE, warning = FALSE}

summary(spp_list$total_area)
#      Min.   1st Qu.    Median      Mean   3rd Qu.      Max.
#     55760   3032000  10790000  18500000  22670000 372100000

qtiles <- c(quantile(spp_list$total_area, 0.0, na.rm = TRUE),
            quantile(spp_list$total_area, .25, na.rm = TRUE),
            quantile(spp_list$total_area, .50, na.rm = TRUE),
            quantile(spp_list$total_area, .75, na.rm = TRUE),
            quantile(spp_list$total_area, 1.0, na.rm = TRUE))

qtile_mdl <- list()
for (i in 1:4) {
  area_clip <- spp_list %>%
    filter(total_area >= qtiles[i] & total_area < qtiles[i + 1])
  qtile_mdl[[i]] <- lm(total_area ~ sm_perc, area_clip)

  qtile_plot <- ggplot(area_clip, aes(total_area, sm_perc, color = sm_range)) + 
    ggtheme_plot + 
    geom_point() + 
    labs(x = 'Total Area (km^2)', 
         y = 'Percent overlap smaller range in larger', 
         title = sprintf('Size of species range (quartile %s) vs. percent overlap of \nsmaller range with larger', i))

  ggsave(file.path(dir_fig, paste0('AreavsOverlap_quantile_', i, '.png')), width = 6, height = 4)
  print(qtile_plot)
}
# First quartile regression:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  1423917     101053  14.091   <2e-16 ***
#     sm_perc         1188       1555   0.764    0.445    
#     Multiple R-squared:  0.001198,	Adjusted R-squared:  -0.0008529 
#
# second quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept)  7011028     275943  25.408  < 2e-16 ***
#     sm_perc       -17725       4228  -4.192 3.28e-05 ***
#     Multiple R-squared:  0.03483,	Adjusted R-squared:  0.03285 
#
# third quartile:
#                 Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) 12813279     507856  25.230  < 2e-16 ***
#     sm_perc        64710       8159   7.931  1.5e-14 ***
#     Multiple R-squared:  0.1146,	Adjusted R-squared:  0.1128 
#
# fourth quartile:
#                   Estimate Std. Error t value Pr(>|t|)    
#     (Intercept) -101736351   17259544  -5.894 7.04e-09 ***
#     sm_perc        1911286     215216   8.881  < 2e-16 ***
#     Multiple R-squared:  0.1399,	Adjusted R-squared:  0.1381

# This avenue of exploration looks like a bust.

```

***