### 2015-12-12 R/data_explore_global_maps.Rmd
more on manuscript...

###Global representation of species

By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.

``` {r setup, message = FALSE, eval = FALSE}
library(rasterVis)
library(ggplot2)
library(maps)
library(raster)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)
library(readr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
dir_fig <- file.path(getwd(), 'figures')

source('R/data_explore_fxns.R')

data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'

spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(sm_perc)) ### will ditch the ones where no IUCN polygon exists

```

``` {r load data sets, echo = FALSE, message = FALSE}
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           col_types = 'cdddd', progress = TRUE) %>%
  as.data.frame() %>%
  filter(presence != 5) ### removes extinct subpopulations

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)

### LOICZID cell ID numbers - used in plotting rasters
loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'
```

```{r global map matched species, echo = FALSE, warning = FALSE, message = FALSE}

### find total species in each dataset for later percent calcs
# am_n_tot   <- am_spp_cells %>% 
#   filter(am_sid %in% unique(spp_list$am_sid)) %>%
#   dplyr::select(am_sid) %>% 
#   unique() %>% nrow()
# iucn_n_tot <- iucn_spp_cells %>% 
#   filter(sciname %in% unique(spp_list$sciname)) %>%
#   dplyr::select(sciname) %>% 
#   unique() %>% nrow()
### good - the same number

### Number of aquamaps species per cell
am_n_per_cell <- am_spp_cells %>%
  filter(am_sid %in% unique(spp_list$am_sid)) %>%
  group_by(loiczid) %>%
  summarize(count_am = n()) %>%
#   summarize(count_am = n(),
#             pct_am = count_am/am_n_tot) %>%
  as.data.frame()

am_sp_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count_am')], 
                  by = 'loiczid', 
                  which = 'count_am', 
                  subsWithNA = TRUE)

### Number of IUCN species per cell
iucn_n_per_cell <- iucn_spp_cells %>%
  group_by(loiczid) %>%
  summarize(count_iucn = n()) %>%
#   summarize(count_iucn = n(),
#             pct_iucn = count_iucn/iucn_n_tot) %>%
  as.data.frame()

iucn_sp_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count_iucn')], 
                    by = 'loiczid', 
                    which = 'count_iucn', 
                    subsWithNA = TRUE)

### Plot both maps
myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))

levelplot(am_sp_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset', 
          colorkey = list( at = seq(0, 1063, length = 15)))

levelplot(iucn_sp_map, par.settings = myTheme, 
          main = 'IUCN Dataset', 
          colorkey = list(at = seq(0, 1063, length = 15)))

```

Now look at full aquamaps species

```{r all_aquamaps, eval = FALSE, echo = FALSE}

am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
  as.data.frame()
# ### am_spp_cell_big is missing loiczid info so bring in hcaf to match
# 
# hcaf <- fread(file.path(dir_anx, 'raw/aquamaps_2014/tables/hcaf.csv')) %>% #bring this in just to get loiczid
#   as.data.frame() %>%
#   dplyr::select(CsquareCode, LOICZID) %>%
#   rename(loiczid = LOICZID)

### find total spp to allow for percent calcs later
am_all_n_tot <- am_spp_cells_big %>% 
  dplyr::select(SpeciesID) %>% 
  unique() %>% nrow()

all_am <- am_spp_cells_big %>%
  dplyr::select(SpeciesID, CsquareCode) %>%
  group_by(CsquareCode) %>%
  summarize(all_count = n(),
            all_pct   = all_count/am_all_n_tot) %>%
  as.data.frame()

all <- all_am %>%
  left_join(hcaf, by = 'CsquareCode')

all_am_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_count')], 
                   by = 'loiczid', 
                   which = 'all_count', 
                   subsWithNA = TRUE)

levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
```

![image](figures/all_aquamaps_map.png)

``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
### find distribution by percent of total species
am_sp_pct_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'pct_am')], 
                      by = 'loiczid', 
                      which = 'pct_am', 
                      subsWithNA = TRUE)
iucn_sp_pct_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'pct_iucn')], 
                        by = 'loiczid', 
                        which = 'pct_iucn', 
                        subsWithNA = TRUE)
### find difference in percent between the two distribution maps; should be small,
### if species are well-matched, since it's all the same species
diff_pct_map    <- iucn_sp_pct_map - am_sp_pct_map

### plot maps with distributions as percentages.  Should it be 0 - 1, or 
### 0 - max value?
levelplot(am_sp_pct_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset pct', 
          colorkey = list( at = seq(0, 1, length = 15)))
levelplot(iucn_sp_pct_map, par.settings = myTheme, 
          main = 'IUCN Dataset pct', 
          colorkey = list(at = seq(0, 1, length = 15)))
### this is the difference one. Should it be min value to max value?
levelplot(diff_pct_map, par.settings = myTheme, 
          main = 'IUCN pct - AM pct', 
          colorkey = list(at = seq(-1, 1, length = 15)))

### Do similar, but with the full aquamaps list (should also do this
### against the full IUCN list...)
all_am_pct_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_pct')], 
                   by = 'loiczid', 
                   which = 'all_pct', 
                   subsWithNA = TRUE)

diff_pct_map2    <- iucn_sp_pct_map - all_am_pct_map
levelplot(all_am_pct_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset, pct')
levelplot(diff_pct_map2, par.settings = myTheme, main = 'iucn (overlap only) - aquamaps (all)')

```