### 2015-12-12 ms_outline.Rmd

---
title: "A comparative analysis of AquaMaps and IUCN species distribution maps"
author: "JA, CO, BH"
output:
  html_document:
    toc: true
    number_sections: true
    theme: "spacelab"
    highlight: "haddock" 
---

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# library(ggplot2)
# library(maps)
# library(readr)
# library(raster)
# library(data.table)
library(tidyr)
library(dplyr)
library(stringr)

# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
#            'Darwin'  = '/Volumes/data_edit',
#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

# dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
# dir_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
# dir_fig <- file.path(getwd(), 'figures')
```

``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
### use data_explore.Rmd to create this file, if it does not yet exist.

data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(popn_category)) %>% ### filter out unassessed species
  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)

### generate other lists

spp_all_file <- 'data/spp_all_2015.csv'
spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)

iucn_marine_file <- 'data/spp_iucn_marine.csv'
spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)

iucn_maps_file <- 'data/spp_iucn_maps.csv'
spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
  rename(sciname = binomial) %>%
  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp

spp_am_file <- 'data/spp_am.csv'
spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)

```

## Main message

Understanding the differences in marine species range maps from IUCN and AquaMaps, and the implications of using one or both.

<!--
##Working Abstract  
* Importance of understanding spatial distribution of species
    * help identify hotspots for global biodiversity 
    * priorities for conservation
    * marine spatial planning
    * assessing biodiversity 
    * monitor impacts (human, natural)
    * global-scale assessments
    * poor overlap of species lists between the two data sets - sometimes helpful to use both
* Challenges
    * difficulty of monitoring marine species in space
    * even more difficult to monitor over time
        * can Aquamaps, as a bioclimatic envelope model, be used to predict changes or used as proxy for species shifts if envelopes are known to have shifted? __can we use this as trend? or does all that count as pressures? seems like human factors can drastically change the trend that wouldn't be captured here__
        * Does that additional data in Aquamaps include predictions for 2050, 2100 (bioclimatic envelopes) - how is that reflected in the species distributions?
* IUCN vs Aquamaps
    * spatial distribution efforts range from modeled to expert input
    * distribution data for 1900+ species found in both data sets
    * differences in predictions of species presence in both distribution and area
        * Aquamaps validation and benefits from _Ready et al. 2010_:
          *  We present and evaluate AquaMaps, a presence-only species distribution modelling system that allows the incorporation of expert knowledge about habitat usage and was designed for maximum output of standardized species range maps at the global scale. In the marine environment there is a significant challenge to the production of range maps due to large biases in the amount and location of occurrence data for most species. AquaMaps is compared with traditional presence-only species distribution modelling methods to determine the quality of outputs under equivalently automated conditions.  _The effect of the inclusion of expert knowledge to AquaMaps is also investigated._ 
          * Model outputs were tested internally, through data partitioning, and externally against independent survey data to determine the ability of models to predict presence versus absence. Models were also tested externally by assessing correlation with independent survey estimates of relative species abundance. AquaMaps outputs compare well to the existing methods tested, and inclusion of expert knowledge results in a general improvement in model outputs. The transparency, speed and adaptability of the AquaMaps system, as well as the existing online framework which allows expert review to compensate for sampling biases and thus improve model predictions are proposed as additional benefits for public and research use alike.
-->

Species range maps are useful in conservation and management, as well as understanding the spatial distribution of movement and overlap of species in different habitats globally. Marine species prove more difficult to map based on their inherent nature of being difficult to observe and track. Two main sources for marine species are the IUCN and AquaMaps. These two datasets differ in methodology and final products, yet the differences between the two are not well documented _(though Aquamaps, and probably IUCN, each probably have some validation studies)_. AquaMaps uses a computer model based on preferential environmental parameters to predict species locations and probability of occurrence, while the IUCN uses expert insight to assign spatial boundaries around known locations of marine species. Approximately 1,900 species are found in both datasets and there are very large differences between these maps. Depending on data usage, choosing one of these datasets over the other can lead to potentially large differences in results as exemplified here by the Ocean Health Index as an example.

-----

# Introduction

Mapping species abundances and distributions is a critical component to understanding global biodiversity and detecting patterns and trends through time. There have been multiple efforts to map terrestrial, freshwater and marine species from local to global scales. These pieces of information inform our understanding of taxonomic assemblages, biogeographical limits to species ranges, suitable habitat for species.

These two spatial datasets have been used to analyze species status globally (Halpern et al. 2012, Selig et al. 2013), global biodiversity patterns and trends, predicting range shifts (Molinos et al. 2015),and setting conservation priorities (Klein et al. 2015). Yet there have been no comprehensive comparisons of these two datasets to identify the differences and provide insight into how using one or the other can influence the outcome of an analysis. With impactful conclusions being drawn from these large-scale studies on global biodiversity and conservation, it's imperative that the data used in these analyses are appropriate and the data decisions are informed.

The IUCN Red List of Threatened Species includes more than 19,000 assessed marine species, 
<!--define what "assessed" means? include DD? keeping those in for now--> but fewer than 4,500 <!--11975 maps in those shapefiles, but fewer than 4500 correspond to assessed marine species... and only 2500 of those overall correpond to AquaMaps species--> of these have associated spatial information. AquaMaps has developed range maps for over 22,800 species. Only 2,200 species are mapped in both datasets. This gap in overlapping species between these two datasets further highlights the need for understanding each of these datasets. Both IUCN and AquaMaps have gaps in marine species information and a significant portion of available data may be overlooked if limited to just one of these datasets.

Species range maps provided by AquaMaps (Kaschner et al. 2013) are model-based, relying on environmental parameters in addition to known occurrences and expert input.  AquaMaps use information about individual species environmental tolerances for salinity, temperature, depth, productivity, and where applicable, sea ice and distance from land, to create a presence-only gridded range map at a resolution of 0.5 degree cells with cell values ranging from 0 to 1.00. These values represent the *relative environmental suitability* for the species, with 0 representing unsuitable habitat and 1.00 being the most suitable habitat. Once produced, each range map maintains a dynamic element that allows individuals to adjust parameters based on their expert knowledge to improve the model-generated map.

Three comparisons presented here:

* Differences in overall species representation - by taxa and by spatial distribution - how well does each set of maps represent the global distribution of species, the global distribution of biodiversity
* alignment between maps for species in common - how well do the maps align, when compared at a species-by-species level
* sensitivity analysis to AquaMaps presence threshold - diff reports use diff thresholds; what is the likely effect? is there a "correct" threshold?
    * this seems like a sub-analysis to justify choice of 0% threshold etc.  
    * ___That graph should be box plots at each threshold, not simply mean and/or median.___

# Understanding each dataset separately

## IUCN species distribution maps

The IUCN publishes species distribution maps for as collections of shapefiles for a range of taxonomic groups, including both terrestrial and marine organisms.  Species spatial distributions are determined by experts based upon known occurrences of the species and expert understanding of range and habitat preferences.  These maps are typically dependent on observation records from databases ___such as GBIF other examples?___, and IUCN provides general guidelines to facilitate consistency in how the maps are generated.

The IUCN presents these maps as 'limits of distibution'; they are meant to indicate that a species probably occurs within the defined polygon, but are not meant to indicate that the species is evenly distributed everywhere within the polygon.  While the polygons roughly define regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.

IUCN publishes spatial distribution map sets only for taxonomic groups whjich have been "comprehensively assessed," i.e. in which at least 90% of the species within the taxon group have been evaluated.  While this mitigates sampling bias within taxa, it also means that entire taxonomic groups remain unavailable until they have met this threshold of comprehensive assessment.  As of December 2015, IUCN had published species distribution maps for nearly 12,000 marine species across 26 taxonomic groups, though fewer than 4,500 of these have been formally assessed under the IUCN Red List of Threatened Species.

* _use as a reference:_ 
    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources

## AquaMaps standardized range maps of marine species

AquaMaps (Kaschner et al. 2013) offers species range maps based on modeled relative environmental suitability.  For each assessed species, data from observations and expert knowledge are used to generate bioclimatic envelopes for sea surface temperature, salinity, depth, productivity, and where applicable, sea ice cover and/or distance to land.  The species' suite of bioclimatic envelopes is compared to a map of environmental attributes on a 0.5 degree global grid, creating a global map of ___probability of occurrence___ (or ___relative environmental suitability?___ (see Kaschner et al 2006 and 2011)), with cell values ranging from 0 (unsuitable) to 1.00 (highly suitable).  These maps can then be tweaked by experts altering the parameters and regenerating the modeled map. As of December 2015, nearly 23,000 species have been mapped, though only 5,600 of these species have been evaluated for the IUCN Red List. ___Expert mode to further refine models?___ See (Kaschner et al. 2013) for more detailed methods.

* ___how is expert insight used? how frequently? new data set includes "reviewed" field___ expert insight seems to be used to generate the environmental suitability envelope for species, to provide sightings info that use used in training the model, and to exclude species from regions of known absence; but the model generates the maps. However, some species maps are reviewed by experts (approved? changed? what happens in the review)

## Comparison between AM and IUCN

* Differences in intent: capture all observations and be inclusive as possible, vs model suitable habitats
* Differences in methodology - expert-defined range maps (over-predicting) vs sightings (under-predicting) vs models: 
from __Kaschner et al 2011__:  _Similar to the trade-offs of different habitat prediction modeling approaches [11], these two methods lie on ___opposite ends of a spectrum from potentially overpredicting expert-derived (range maps) to potentially underpredicting (empirical sighting surveys) range sizes____
    * are there other sources (non-IUCN and AM sources) that can back this up?
* Differences in the interpretation: probability/RES, vs. presence "probably within this polygon, but not everywhere within" - apples to oranges.

-----

# Analysis of taxonomic and regional inclusion

## Overlap between assessed species

Due to differences in both methodology and intent, there are significant differences in the coverage afforded by the species distribution maps published by AquaMaps and those published by the IUCN.  This is immediately clear from an examination of the number of species maps offered by each source, and the small proportion of species with maps from both sources.

| Map set   | n~total~                | n~Red_List~                                                       |
| :-------: | ----------------------: | :---------------------------------------------------------------: | 
| AquaMaps  | `r nrow(spp_am)`        | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_marine$sciname))` | 
| IUCN      | `r nrow(spp_iucn_maps)` | all                                                               | 
| Both      | `r nrow(spp_am %>% filter(sciname %in% spp_iucn_maps$sciname))` | all                       |

___Should we break it down by "reviewed" Aquamaps as well?___

The AquaMaps species list includes all species available as of December 2015.  The IUCN species list includes all species maps available directly from the IUCN Red List website (http://www.iucnredlist.org/technical-documents/spatial-data) as of December 2015, filtered to include only species with habitat listed as "marine."  For the purposes of this analysis, we did not include IUCN range maps for bird species, as those range maps are hosted separately by BirdLife International.

## Taxonomic distribution between data sets

![](figures/bar_by_taxa_all.png)

![](figures/bar_by_taxa_assessed.png)

As IUCN only releases map sets when a taxonomic group (on the scale of order or family) has been comprehensively assessed, spatial data for many taxonomic classes are not available, and within a class, the assessed sub-groups may not represent the entire class.  For example, as of this writing, IUCN has released no spatial data for class Elasmobranchii (cartilaginous fishes including sharks and rays); and while IUCN offers a large number of maps within class Actinopterygii (ray-finned bony fishes), the available maps include only a few primarily tropical taxonomic sub-groups, such as wrasses, damselfish, butterflyfish, tunas, and billfishes.  However, IUCN's criterion of comprehensive assessment greatly reduces the risk of sample bias within the bounds of the assessed taxonomic groups.

The release of AquaMaps distribution maps is not limited in the same way, and maps are available across a much larger range of taxonomic classes; however, there is no guarantee that the list of species included within each class is a representative cross section of the entire class.  ___How do they choose their species to assess?___

## Spatial global distribution between data sets

___main messages:___

* IUCN seems to be weighted toward tropics; AM might be more evenly distributed?
* This difference in distribution might come into play when selecting a data set for a more regionally-focused study
* ... or for global studies, it might be desirable to use both data sets to get better global distribution?
* Small number of overlapping species indicates that each offers data on distinct spp
* This section is really supporting the idea that it is critical to understanding the differences, as both data sets support one another

Using predefined species groups by the IUCN, it is clear to see that tropical species - specifically wrasses, corals, blennies and groupers, make up the majority of species found in both. This is driven more by the IUCN dataset since the number of species with spatial data is directly attributed to the species that have assessments. ___This could be shown using the n = XXX on the species group quadrant breakdown chart too.___

___Total species occurrences for each data set: all AM, all IUCN, and overlap (one? the other? occurrences of both?)___

![](figures/all_aquamaps_map.png)

_Relevance of this_ - unequal distribution of assessed species - strengths and weaknesses of each; benefits of using both maybe?  save for discussion

Each data set offers spatial distribution information for large numbers of species.  However, the data sets vary in terms of taxonomic coverage and regional coverage.  For spatial assessments of biodiversity, the choice of one data set over the other is likely to create significantly different results.  For studies confined to a narrow range of taxa or to a narrow spatial scale, one data set may offer an advantage over the other in the number of species maps available.  For global scale biodiversity studies, however, the selection of one data set over the other will entail tradeoffs in spatial coverage, taxonomic breadth, and taxonomic depth. ___find a reference that describes what makes a "good" data set for global biodiversity, e.g. OHI___

-----

# Analysis of spatial alignment by species

## Defining spatial alignment between the two data sets

AquaMaps and IUCN marine species distribution maps ostensibly describe the same information, and while differences in methodology and input data will inevitably lead to differences, we would expect to see strong similarities between range maps for any given species.

To facilitate our analysis, we selected the subset of marine species that have range maps in both IUCN (filtering out ranges identified as extinct) and AquaMaps current native distribution (n = `r nrow(spp_map_pairs)`), and used genus and species binomials to identify paired maps.  We then rasterized the IUCN shapefiles to match the Aquamaps half-cell raster.  We flattened each map by assigning to each cell a simple "presence/absence" value:

* For IUCN data, we defined "presence" within a cell as any non-zero overlap with a range map polygon.
* For AquaMaps data, we defined presence as any cell with non-zero probability ___(or RES?)___.  
* In both cases, these criteria for presence will maximize the number of cells included.

Using these criteria, we found that for `r round(nrow(spp_map_pairs %>% filter(sm_range == 'AM')) / nrow(spp_map_pairs) * 100, 1)`% of species, the IUCN distribution maps indicated a larger species range than the range indicated by AquaMaps.

Overlaying distribution maps from the two data sets for a given species, we calculated two dimensions of spatial alignment: distribution match, which we defined as the ratio of the smaller range that overlapped with the larger range; and area match, which we defined as the ratio of the smaller range to the larger range.

$$Match_{dist} = \frac{A_{overlap}}{A_{larger}} * 100\%$$

$$Match_{area} = \frac{A_{smaller}}{A_{larger}} * 100\%$$

For a species whose distribution is well understood, and whose range is described in both data sets, we would expect to see a value near 100% for each dimension of alignment, indicating near-total inclusion of the smaller range within the larger (distribution alignment) and very similar extent of range (area alignment).

## Comparing spatial alignment between the two data sets

For each of the `r nrow(spp_map_pairs)` species whose range is described by both data sets, we calculated the distribution match and the area ratio, and plotted the species map pairs on a scatter plot of distribution alignment vs area alignment (Figure 1).

![](figures/AreaRatiovsPercOverlap.png)

___Figure XXX: Distribution alignment vs area alignment for `r nrow(spp_map_pairs)` species mapped in both Aquamaps and IUCN species distribution maps.___

___ditch the species identifiers here?  or highlight ones that are used in discussion of possible mechanistic explanations for differences?___

We divided the plot into quadrants based on the median values for each axis to differentiate between four different qualities of alignment:

``` {r calc medians and quadrant counts, echo = FALSE}
dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
```

| Description              |  Distribution match  |     Area match       |
| :----------------------: | :------------------: | :------------------: |
| q1: Well-aligned         | >= `r dist_median` % | >= `r area_median` % |
| q2: Distribution-aligned |  < `r dist_median` % | >= `r area_median` % |
| q3: Area-aligned         | >= `r dist_median` % |  < `r area_median` % |
| q4: Poorly-aligned       |  < `r dist_median` % |  < `r area_median` % |

The upper right quadrant (quadrant 1) comprises species whose maps largely agree (better than median value) in both spatial distribution and the area of described ranges.  For these species, the expert-drawn range map and the bioclimatic envelope model produce similar predictions of species distribution. 
(n = `r q1_count`; `r round(q1_count/nrow(spp_map_pairs) * 100, 1)` %)

The other quadrants each suggest a different story.  Species map pairs that fall within the upper left quadrant (quadrant 2: well-matched in distribution but poorly-matched in area) indicate species in which the range of the smaller map falls generally within the larger map, but the larger map may include more generous buffers or may include areas unrepresented in the smaller map. 
(n = `r q2_count`; `r round(q2_count/nrow(spp_map_pairs) * 100, 1)` %)

The lower right quadrant (quadrant 3: well-matched in area but poorly matched in distribution) includes species for which the two maps generally agree in range area, but disagree in where those ranges occur.  Disagreement in distribution seems to indicate a greater issue than disagreement in area extent, so we categorized map pairs in this quadrant as being more problematic than map pairs in either of the upper quadrants. 
(n = `r q3_count`; `r round(q3_count/nrow(spp_map_pairs) * 100, 1)` %)

The lower left quadrant (quadrant 4: poorly matched in both dimensions) indicates species for which the map pairs fail to agree in both area and distribution. 
(n = `r q4_count`; `r round(q4_count/nrow(spp_map_pairs) * 100, 1)` %)

![](figures/barchart_spp_gp_quads.png)

Figure 2 divides the `r nrow(spp_map_pairs)` species map pairs into taxonomic groups as classified by the published IUCN range maps, and examines the breakdown of each taxonomic group by quadrant.

___move to discussion?___ We were not surprised to find that map pairs for marine mammals, tunas, and billfishes were well-matched (quadrants 1 and 2); many of these species are global-ranging pelagic organisms, with some ranges encompassing nearly the entire ocean.  Mangroves are also not a surprise, being immobile and easily mapped by remote sensing methods.

_Map pairs for corals, most of which fall into quadrant 2, ***hypothesis...*** generally agree well in distribution, but in many cases IUCN ranges include buffers that extend much farther from coastlines than AquaMaps modeled ranges, which use seafloor depth as a limiting factor._

___Other hypotheses/case studies to highlight possible mechanistic differences?___

# AquaMaps: Effect of changing "presence" threshold on area representation

![](figures/AMThreshold.png) ___add a mean/median vline on this?___ do we need this figure? it looks cool though... why broken down by risk category? Ditch NE category since most studies probably care about the species health as well as distribution?

___Identify several studies that used different thresholds - OHI, MPA gap analysis, AquaMaps mammal distribution___ to justify this micro-analysis

![](figures/scatter_AMThreshold_mockup.png)
Effect of selecting different thresholds on total "present" area - linearish, no big steps or drops to indicate sudden change in tradeoff.  Do regression - shallow slope indicates that sensitivity is less than 1:1.

-----

# Implications

## Application to OHI

The global Ocean Health Index (Halpern et al. 2012), an index made up of 10 goals, utilizes both of these datasets to inform the  Biodiversity goal. The biodiversity goal is made up of two parts; Species and Habitats. The species subgoal uses the IUCN range map data and threat status to calculate average species status in global ocean regions around the world. Spatial range maps are developed by experts for species that are assessed. The assessment process happens in phases and only [X] number of marine species have been given spatial range maps. AquaMaps has spatial range maps for over 17,000 species. For those species that are missing spatial information from IUCN, AquaMaps data was used. If a species was assigned an IUCN threat status but did not have spatial range information in either dataset, it was excluded.

This decision to use IUCN data preferentially to AquaMaps data was made based on the fact that experts are developing the IUCN maps and AquaMaps data is based off a computer model, which in some cases is then consulted by an expert and appropriately altered.

Between the two databases, there are 1,923 species that have a range map in both. Figure [X] shows the change in status score for the Species Subgoal within the global Ocean Health Index under three different scenarios.

![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)

<!--
![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)
-->

Scenario 1 shows change in score if IUCN maps were still preferentially used over AquaMaps, but the AquaMaps threshold was set to 1% rather than 40% probability of occurrence. Scenario 2 shows the change in score if we switched preferences for AquaMaps over IUCN data while keeping the 40% probability of occurrence threshold. The final Scenario is the same as scenario 2 but without a probability of occurrence threshold.

It is important to remember that although only ___[X (1,923/total number of species)]%___ of all species included in the species goal are found in both datasets, there are still significant differences when electing to choose one dataset over the other.

-----

## Application to MPA Gap Analysis

___need more info on these options? they both sound interesting though___
- could talk about MPA gap analysis

# Conclusion

-----

#References

__Key AquaMaps publications:__

Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF

Ready, J., K. Kaschner, A.B. South, P.D Eastwood, T. Rees, J. Rius, E. Agbayani, S. Kullander and R. Froese (2010). Predicting the distributions of marine organisms at the global scale. Ecological Modelling 221(3): 467-478. PDF

__Papers based on AquaMaps:__ ___Use these for section 4?___

Jones, M.C., S.R. Dyeb, J.K. Pinnegar and W.W.L. Cheung (2012). Modelling commercial fish distributions: Prediction and assessment using different approaches. Ecological Modelling 225(2012): 133-145. PDF

Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF

Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D. (2014). Manual of marine nd coastal datasets of biodiversity importance. May 2014 release. Cambridge (UK): UNEP World Conservation Monitoring Centre. 28 pp. (+ 4 annexes totalling 174 pp. and one e-supplement). PDF

#References

García Molinos, Jorge, Benjamin S. Halpern, David S. Schoeman, Christopher J. Brown, Wolfgang Kiessling, Pippa J. Moore, John M. Pandolfi, Elvira S. Poloczanska, Anthony J. Richardson, and Michael T. Burrows. “Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015). doi:10.1038/nclimate2769.

Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L. McLeod, Jameal F. Samhouri, Steven K. Katona, Kristin Kleisner, et al. “An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no. 7413 (August 30, 2012): 615–20. doi:10.1038/nature11397.

<!--
Halpin, P.N., A.J. Read, E. Fujioka, B.D. Best, B. Donnelly, L.J. Hazen, C. Kot, K. Urian, E. LaBrecque, A. Dimatteo, J. Cleary, C. Good, L.B. Crowder, and K.D. Hyrenbach. 2009. OBIS-SEAMAP: The world data center for marine mammal, sea bird, and sea turtle distributions. Oceanography 22(2):104-115
-->

Klein, Carissa J., Christopher J. Brown, Benjamin S. Halpern, Daniel B. Segan, Jennifer McGowan, Maria Beger, and James E.M. Watson. “Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539. doi:10.1038/srep17539.

Selig, Elizabeth R., Catherine Longo, Benjamin S. Halpern, Benjamin D. Best, Darren Hardy, Cristiane T. Elfes, Courtney Scarborough, Kristin M. Kleisner, and Steven K. Katona. “Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no. 4 (April 11, 2013): e60284. doi:10.1371/journal.pone.0060284.