### 2015-12-14 R/data_explore_global_maps.Rmd
testing out AM vs IUCN difference level plot

###Global representation of species

By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.

``` {r setup, message = FALSE, eval = FALSE}
library(rasterVis)
library(ggplot2)
library(maps)
library(raster)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)
library(readr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')
dir_fig <- file.path(getwd(), 'figures')

source('R/data_explore_fxns.R')

data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'

spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(sm_perc)) ### will ditch the ones where no IUCN polygon exists

```

``` {r load data sets, echo = FALSE, message = FALSE}
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           col_types = 'cdddd', progress = TRUE) %>%
  as.data.frame() %>%
  filter(presence != 5) ### removes extinct subpopulations

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)

### LOICZID cell ID numbers - used in plotting rasters
loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'
```

```{r global map matched species, echo = FALSE, warning = FALSE, message = FALSE}

### find total species in each dataset for later percent calcs
# am_n_tot   <- am_spp_cells %>% 
#   filter(am_sid %in% unique(spp_list$am_sid)) %>%
#   dplyr::select(am_sid) %>% 
#   unique() %>% nrow()
# iucn_n_tot <- iucn_spp_cells %>% 
#   filter(sciname %in% unique(spp_list$sciname)) %>%
#   dplyr::select(sciname) %>% 
#   unique() %>% nrow()
### good - the same number

### Number of aquamaps species per cell
iucn_pair_rast_file <- 'rasters/iucn_pair_rast.tif'
am_pair_rast_file   <- 'rasters/am_pair_rast.tif'

if(!file.exists(iucn_pair_rast_file) | !file.exists(am_pair_rast_file)) {
  am_n_per_cell <- am_spp_cells %>%
  filter(am_sid %in% unique(spp_list$am_sid)) %>%
  group_by(loiczid) %>%
  summarize(count_am = n()) %>%
#   summarize(count_am = n(),
#             pct_am = count_am/am_n_tot) %>%
  as.data.frame()

  am_sp_map <- subs(loiczid_raster, am_n_per_cell[ , c('loiczid', 'count_am')], 
                    by = 'loiczid', 
                    which = 'count_am', 
                    subsWithNA = TRUE)

  ### Number of IUCN species per cell
  iucn_n_per_cell <- iucn_spp_cells %>%
    group_by(loiczid) %>%
    summarize(count_iucn = n()) %>%
  #   summarize(count_iucn = n(),
  #             pct_iucn = count_iucn/iucn_n_tot) %>%
    as.data.frame()

  iucn_sp_map <- subs(loiczid_raster, iucn_n_per_cell[ , c('loiczid', 'count_iucn')], 
                      by = 'loiczid', 
                      which = 'count_iucn', 
                      subsWithNA = TRUE)

  writeRaster(iucn_sp_map, filename = iucn_pair_rast_file, overwrite = TRUE)
  writeRaster(am_sp_map, filename = am_pair_rast_file, overwrite = TRUE)
} else {
  am_sp_map   <- raster(am_pair_rast_file)
  iucn_sp_map <- raster(iucn_pair_rast_file)
}

### Plot both maps
myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))

levelplot(am_sp_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset', 
          colorkey = list( at = seq(0, 1063, length = 15)))

levelplot(iucn_sp_map, par.settings = myTheme, 
          main = 'IUCN Dataset', 
          colorkey = list(at = seq(0, 1063, length = 15)))

```

Now look at full aquamaps species

```{r all_aquamaps, eval = FALSE, echo = FALSE}

all_am_rast_file   <- 'rasters/am_all_rast.tif'

if(!file.exists(am_all_rast_file)) {
  am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
    as.data.frame(stringsAsFactors = FALSE)

  ### find total spp to allow for percent calcs later
  # am_all_n_tot <- am_spp_cells_big %>% 
  #   dplyr::select(SpeciesID) %>% 
  #   unique() %>% nrow()

  all_am <- am_spp_cells_big %>%
    group_by(loiczid) %>%
    summarize(all_count = n()) %>%
  #   summarize(all_count = n(),
  #             all_pct   = all_count/am_all_n_tot) %>%
    as.data.frame()

  all_am_map <- subs(loiczid_raster, all_am[ , c('loiczid', 'all_count')], 
                     by = 'loiczid', 
                     which = 'all_count', 
                     subsWithNA = TRUE)

  writeRaster(all_am_map, filename = am_all_rast_file, overwrite = TRUE)

} else {
  all_am_map   <- raster(all_am_rast_file)
}

levelplot(all_am_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset')
```

Now look at full IUCN species

```{r all_aquamaps, eval = FALSE, echo = FALSE}

all_iucn_rast_file   <- 'rasters/am_all_rast.tif'

if(!file.exists(iucn_all_rast_file)) {
  iucn_cells_file    <- file.path(dir_data_iucn, 'iucn_cells_2015.csv')
  iucn_spp_cells_all <- fread(iucn_cells_file) %>%
        as.data.frame(stringsAsFactors = FALSE)

  ### find total spp to allow for percent calcs later
  # am_all_n_tot <- am_spp_cells_big %>% 
  #   dplyr::select(SpeciesID) %>% 
  #   unique() %>% nrow()

  all_iucn <- iucn_spp_cells_all %>%
    group_by(loiczid) %>%
    summarize(all_count = n()) %>%
  #   summarize(all_count = n(),
  #             all_pct   = all_count/am_all_n_tot) %>%
    as.data.frame()

  all_iucn_map <- subs(loiczid_raster, all_iucn[ , c('loiczid', 'all_count')], 
                     by = 'loiczid', 
                     which = 'all_count', 
                     subsWithNA = TRUE)

  writeRaster(all_iucn_map, filename = all_iucn_rast_file, overwrite = TRUE)

} else {
  all_iucn_map   <- raster(iucn_all_rast_file)
}

levelplot(all_iucn_map, par.settings = myTheme, main = 'Entire IUCN Dataset')
```

![image](figures/all_aquamaps_map.png)

``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
### rescale values relative to maximum cell value

am_map_max <- max(all_am_map, na.rm = TRUE)
all_am_pct_map <- all_am_map/am_map_max

iucn_map_max <- max(all_iucn_map, na.rm = TRUE)
all_iucn_pct_map <- all_iucn_map/iucn_map_max

diff_pct_map    <- all_am_pct_map - all_iucn_pct_map

### plot maps with distributions as percentages.  Should it be 0 - 1, or 
### 0 - max value?
levelplot(am_sp_pct_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset pct', 
          colorkey = list( at = seq(0, 1, length = 15)))
levelplot(iucn_sp_pct_map, par.settings = myTheme, 
          main = 'IUCN Dataset pct', 
          colorkey = list(at = seq(0, 1, length = 15)))
### this is the difference one. Should it be min value to max value?
levelplot(diff_pct_map, par.settings = myTheme, 
          main = 'IUCN pct - AM pct', 
          colorkey = list(at = seq(-1, 1, length = 15)))

### Do similar, but with the full aquamaps list (should also do this
### against the full IUCN list...)
all_am_pct_map <- subs(loiczid_raster, all[ , c('loiczid', 'all_pct')], 
                   by = 'loiczid', 
                   which = 'all_pct', 
                   subsWithNA = TRUE)

diff_pct_map2    <- iucn_sp_pct_map - all_am_pct_map
levelplot(all_am_pct_map, par.settings = myTheme, main = 'Entire Aquamaps Dataset, pct')
levelplot(diff_pct_map2, par.settings = myTheme, main = 'iucn (overlap only) - aquamaps (all)')

```