### 2015-12-17 data_explore.Rmd
updated draft to send to Ben

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(maps)
library(readr)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(getwd(), 'figures')

source('R/data_explore_fxns.R')
```

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           col_types = 'cdddd', progress = TRUE) %>%
  as.data.frame() %>%
  filter(presence != 5) ### removes extinct subpopulations

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), progress = TRUE)

### LOICZID cell ID numbers - used in plotting rasters
loiczid_raster_file  <- 'shiny/data/loiczid_raster.grd'
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'
```

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'

if (!file.exists(data_file)) {
  ### species info list - scinames, IUCN species ID, AM species ID, category etc
  spp_list_base <- read.csv('data/spp_am_v_iucn_2015_inclDD.csv', stringsAsFactors = FALSE)

  create_spp_list(spp_list_base, data_file)
}

spp_list <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(sm_perc)) 
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?
spp_list <- spp_list %>%
  left_join(read.csv('sciname_lookups/spp_groupnames.csv', stringsAsFactors = FALSE),
            by = 'spp_group')

```

``` {r child = 'R/data_explore_pt1.Rmd', eval = FALSE}
```
<!-- # bar chart: number of spp per group
# scatter plot: area compare IUCN vs AM, including log plot and plot by species group
# rangemap: Chromis alpha -->

``` {r, child = 'R/data_explore_pt2.Rmd', eval = FALSE}
```

``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

spp_list <- spp_list %>%
  mutate(area_ratio = am_area / iucn_area,
         area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
         area_ratio = area_ratio * 100)

### define 25% and 75% quartiles for both x and y axes
area_ratio_quartiles <- quantile(spp_list$area_ratio, na.rm = TRUE)[2:4]
pct_overlap_quartiles <- quantile(spp_list$sm_perc, na.rm = TRUE)[2:4]

### define representative species from each quadrant, for over-plotting
# spp_quads <- c('q1' = 'Kajikia albida', 
#                'q2' = 'Galaxea cryptoramosa', 
#                'q4' = 'Acanthurus nigroris', 
#                'q4a' = 'Metanephrops australiensis', 
#                'q3' = 'Conus magnificus')

# spp_rando <- c("Lagocephalus lagocephalus",
#                "Makaira nigricans",
#                "Pentacheles laevis",
#                "Stereomastis suhmi",
#                "Stereomastis surda") 
# 
### truncate the species names for the representative species
# spp_list_label <- spp_list %>%
#   filter(sciname %in% spp_rando) %>%
#   separate(sciname, c('g', 's'), sep = ' ', remove = FALSE) %>%
#   mutate(sciname_label = paste(str_sub(g, 1, 1), '. ', s, sep = '')) %>%
#   dplyr::select(-g, -s)

### mongo plot time
scatter_arearatio_vs_overlap <- ggplot(spp_list %>% 
                                         transform(popn_category = reorder(popn_category, category_score)),
                                       aes(x = area_ratio, y = sm_perc, 
                                           color = popn_category,
                                           size = lg_area_pct)) + 
  ggtheme_plot + 
  ### color the quadrant backgrounds:
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "green4")  + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax = 100, alpha = .3, fill= "greenyellow") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax = 100, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orange") + 
    annotate("rect", xmin = area_ratio_quartiles[2],  xmax =   0, 
                     ymin = pct_overlap_quartiles[2], ymax =   0, alpha = .3, fill= "orangered2") + 
  geom_point() + 
  scale_colour_manual(values = c('LC' = 'green4', 
                                 'NT' = 'greenyellow', 
                                 'VU' = 'yellow', 
                                 'EN' = 'orange', 
                                 'CR' = 'orangered2',
                                 'DD' = 'grey30')) +
  geom_vline(x = area_ratio_quartiles[c(1, 3)],
             color = 'grey50', linetype = 'dashed') +
  geom_hline(y = pct_overlap_quartiles[c(1, 3)],
             color = 'grey50', linetype = 'dashed') +
#   geom_point(data = spp_list_label, shape = 4, color = 'grey30', size = 4) + 
#   geom_text(data = spp_list_label, 
#             aes(label = sciname_label), 
#             color = 'grey30',
#             fontface = 'bold.italic',
#             size = 2,
#             hjust = .25,
#             vjust = 1.5) +
  labs(x = 'Area match: \n(range area smaller : range area larger) * 100%', 
       y = 'Distribution match: \n % overlap of smaller range w/ larger', 
       title = 'Area match vs distribution match',
       color = 'Risk', size = '% ocean range')

ggsave(file.path(dir_fig, 'AreaRatiovsPercOverlap.png'), width = 6, height = 4)

print(scatter_arearatio_vs_overlap)

```

```{r plot_spp_from_quadrants, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

for (i in 1:length(spp_quads)) { # i = 1
  spp <- spp_quads[i]
  spp_quad <- names(spp_quads)[i]

  spp_plot <- plot_rangemap(spp)
  print(spp_plot)

  ggsave(filename = file.path(dir_fig, sprintf('rangemap_%s_%s.png', spp_quad,
                                               str_replace(tolower(spp), ' ', '_'))),
         plot = spp_plot, 
         width = 8, height = 4, dpi = 200)
}

```

``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}

spp_list1 <- spp_list %>%
  group_by(spp_group_text) %>%
  mutate(n_spp = n()) %>%
  ungroup()

### Create species lists by quadrant
spp_list_q1 <- spp_list1 %>%
  filter(area_ratio >= median(area_ratio) & sm_perc >= median(sm_perc))
spp_list_q2 <- spp_list1 %>%
  filter(area_ratio < median(area_ratio) & sm_perc >= median(sm_perc))
spp_list_q3 <- spp_list1 %>%
  filter(area_ratio >= median(area_ratio) & sm_perc < median(sm_perc))
spp_list_q4 <- spp_list1 %>%
  filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))

### Summarize quadrant species list by species group (e.g. CORALS1)
spp_gp_q1 <- spp_list_q1 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q1 = n())
spp_gp_q2 <- spp_list_q2 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q2 = n())
spp_gp_q3 <- spp_list_q3 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q3 = n())
spp_gp_q4 <- spp_list_q4 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q4 = n())

### Join species group by quadrant into a single dataframe
spp_gp_quadrants <- spp_gp_q1 %>%
  full_join(spp_gp_q2, by = c('spp_group_text', 'n_spp')) %>%
  full_join(spp_gp_q3, by = c('spp_group_text', 'n_spp')) %>%
  full_join(spp_gp_q4, by = c('spp_group_text', 'n_spp')) %>%
  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
  mutate(quad = str_replace(quad, 'n_spp_', ''),
         pct_quad = n_quad/n_spp,
         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))

spp_gp_quadrants <- spp_gp_quadrants %>%
  left_join(spp_gp_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(spp_group_text, pct_q1 = pct_quad),
            by = 'spp_group_text') %>%
  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
         quad = str_replace(quad, 'q3', 'q3 area-matched'),
         quad = str_replace(quad, 'q4', 'q4 poorly matched'))

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'bin', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
  coord_flip() +
  labs(x = 'Species Group', 
       y = 'Relative number of species per quadrant', 
       title = 'Species group breakdown by quadrant',
       fill = 'Match quadrant')

print(barchart_spp_gp_quads)
ggsave(plot = barchart_spp_gp_quads,
       filename = file.path(dir_fig, 'barchart_spp_gp_quads.png'),
       width = 6, height = 4)

```

``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}

spp_list1 <- spp_list %>%
  group_by(popn_category) %>%
  mutate(n_spp = n()) %>%
  ungroup()

### Create species lists by quadrant
spp_list_q1 <- spp_list1 %>%
  filter(area_ratio >= median(area_ratio) & sm_perc >= median(sm_perc))
spp_list_q2 <- spp_list1 %>%
  filter(area_ratio < median(area_ratio) & sm_perc >= median(sm_perc))
spp_list_q3 <- spp_list1 %>%
  filter(area_ratio >= median(area_ratio) & sm_perc < median(sm_perc))
spp_list_q4 <- spp_list1 %>%
  filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))

### Summarize quadrant species list by species group (e.g. CORALS1)
spp_risk_q1 <- spp_list_q1 %>%
  group_by(popn_category) %>%
  summarize(n_spp = first(n_spp), n_spp_q1 = n())
spp_risk_q2 <- spp_list_q2 %>%
  group_by(popn_category) %>%
  summarize(n_spp = first(n_spp), n_spp_q2 = n())
spp_risk_q3 <- spp_list_q3 %>%
  group_by(popn_category) %>%
  summarize(n_spp = first(n_spp), n_spp_q3 = n())
spp_risk_q4 <- spp_list_q4 %>%
  group_by(popn_category) %>%
  summarize(n_spp = first(n_spp), n_spp_q4 = n())

### Join species group by quadrant into a single dataframe
spp_risk_quadrants <- spp_risk_q1 %>%
  full_join(spp_risk_q2, by = c('popn_category', 'n_spp')) %>%
  full_join(spp_risk_q3, by = c('popn_category', 'n_spp')) %>%
  full_join(spp_risk_q4, by = c('popn_category', 'n_spp')) %>%
  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
  mutate(quad = str_replace(quad, 'n_spp_', ''),
         pct_quad = n_quad/n_spp,
         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))

spp_risk_quadrants <- spp_risk_quadrants %>%
  left_join(spp_risk_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(popn_category, pct_q1 = pct_quad),
            by = 'popn_category') %>%
  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
         quad = str_replace(quad, 'q3', 'q3 area-matched'),
         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
  left_join(spp_list %>%
              dplyr::select(popn_category, cat_rank = category_score) %>%
              unique() %>%
              mutate(cat_rank = ifelse(popn_category == 'DD', -1, cat_rank)),
            by = 'popn_category')

### Plot the bar chart
barchart_spp_risk_quads <- ggplot(spp_risk_quadrants %>% 
                                  transform(popn_category = reorder(popn_category, cat_rank)), 
                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  geom_bar(stat = 'bin', alpha = .5) +
  scale_fill_manual(values = c('q1 well-matched'   = 'green4', 
                               'q2 dist-matched'   = 'greenyellow', 
                               'q3 area-matched'   = 'orange', 
                               'q4 poorly matched' = 'orangered2')) +
  geom_text(aes(label = sprintf('n = %s', n_spp), y = .05), hjust = 0, size = 4) +
  coord_flip() +
  labs(x = 'Risk category', 
       y = 'Relative number of species per quadrant', 
       title = 'Risk category breakdown by quadrant',
       fill = 'Match quadrant')

print(barchart_spp_risk_quads)
ggsave(plot = barchart_spp_risk_quads,
       filename = file.path(dir_fig, 'barchart_spp_risk_quads.png'),
       width = 6, height = 4)

```

``` {r child = 'R/data_explore_pt3.Rmd', eval = FALSE}
```
<!-- # Multiple species maps per quadrant -->

***

Breakdown maps by taxa

``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
### use data_explore.Rmd to create this file, if it does not yet exist.

data_file <- 'data/spp_list_w_area_percent_noextinct_2015.csv'
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(popn_category)) %>% ### filter out unassessed species
  filter(!is.na(sm_perc))           ### filter out NA maps (usually IUCN maps with no polygons)

### generate other lists

spp_all_file <- 'data/spp_all_2015.csv'
spp_all <- read.csv(spp_all_file, stringsAsFactors = FALSE)

iucn_marine_file <- 'data/spp_iucn_marine.csv'
spp_iucn_marine <- read.csv(iucn_marine_file, stringsAsFactors = FALSE)

iucn_maps_file <- 'data/spp_iucn_maps.csv'
spp_iucn_maps <- read.csv(iucn_maps_file, stringsAsFactors = FALSE) %>%
  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
  filter(sciname %in% spp_iucn_marine$sciname)   ### ditch freshwater spp

spp_am_file <- 'data/spp_am.csv'
spp_am <- read.csv(spp_am_file, stringsAsFactors = FALSE)

```

``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE}

sciname_lookup <- read.csv('sciname_lookups/spp_iucn_all.csv', stringsAsFactors = FALSE)

iucn_maps_class <- spp_iucn_maps %>% 
  left_join(sciname_lookup %>% 
              select(sciname, class) %>%
              unique(), 
            by = 'sciname')

taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
  filter(!is.na(class)) %>% ### don't count the unknowns
  mutate(class = str_to_title(class),
         n_iucn_tot = n()) %>%        ### length of whole maps list
  group_by(class) %>%
  summarize(n_iucn_class_all  = n(),
            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))

taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
  filter(!is.na(class)) %>% ### don't count the unknowns
  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
  mutate(class = str_to_title(class),
         n_iucn_solo = n()) %>%       ### length of IUCN only list
  group_by(class) %>%
  summarize(n_iucn_class_solo = n(),
            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))

taxa_am_all <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  mutate(class = str_to_title(class),
         n_am_tot = n()) %>%        ### length of whole AM maps list
  group_by(class) %>%
  summarize(n_am_class_all  = n(),
            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))

taxa_am_solo <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
  mutate(class = str_to_title(class),
         n_am_solo = n()) %>%       ### length of AM only list
  group_by(class) %>%
  summarize(n_am_class_solo = n(),
            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))

taxa_both <- spp_am %>%
  filter(sciname %in% spp_iucn_maps$sciname) %>%
  filter(!class %in% c('Not assigned', 'null')) %>%
  mutate(class = str_to_title(class),
         n_both_tot = n()) %>%
  group_by(class) %>%
  summarize(n_class_both = n(),
            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))

taxa_am_all_assessed <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(sciname %in% spp_iucn_marine$sciname) %>%
  mutate(class = str_to_title(class),
         n_am_tot = n()) %>%        ### length of whole AM maps list
  group_by(class) %>%
  summarize(n_am_class_ass  = n(),
            p_am_class_ass  = round((n_am_class_ass/first(n_am_tot) * 100), 2))

taxa_am_solo_assessed <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(sciname %in% spp_iucn_marine$sciname) %>%
  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
  mutate(class = str_to_title(class),
         n_am_solo = n()) %>%       ### length of AM only list
  group_by(class) %>%
  summarize(n_am_class_solo_ass = n(),
            p_am_class_solo_ass = round((n_am_class_solo_ass/first(n_am_solo) * 100), 2))

taxa_df <- taxa_am_all %>%
  full_join(taxa_am_solo,   by = 'class') %>%
  full_join(taxa_am_all_assessed,  by = 'class') %>%
  full_join(taxa_am_solo_assessed, by = 'class') %>%
  full_join(taxa_iucn_all,  by = 'class') %>%
  full_join(taxa_iucn_solo, by = 'class') %>%
  full_join(taxa_both,      by = 'class') %>%
  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))

knitr::kable(taxa_df %>%
               select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both))
```

``` {r plot bar charts by taxa, echo = FALSE, message = FALSE, warning = FALSE}
### bar charts comparing number of species in each class for each data set

taxa_df_long1 <- taxa_df %>%
  select(class, AquaMaps = n_am_class_all, IUCN = n_iucn_class_all, Both = n_class_both) %>%
  gather(dataset, n_spp, AquaMaps:Both)
bar_taxa1 <- ggplot(taxa_df_long1, aes(x = class, y = n_spp)) +
  ggtheme_plot + 
  geom_bar(stat = 'identity') +
  facet_grid(dataset ~ ., scales = 'free_y') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1))
print(bar_taxa1)

taxa_df_long2 <- taxa_df %>%
  select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
  gather(dataset, n_spp, AquaMaps:Both) %>%
  group_by(class) %>%
  mutate(n_total = sum(n_spp, na.rm = TRUE),
         percent = n_spp/n_total * 100)
bar_taxa2 <- ggplot(taxa_df_long2, aes(x = class, y = percent, fill = dataset)) +
  ggtheme_plot + 
  geom_bar(stat = 'identity', alpha = .5) +
  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
  geom_text(aes(label = sprintf('n = %s', n_total), y = .05), hjust = 0, size = 4, angle = 90, hjust = 1) +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  labs(x = 'Class', y = 'Percent of mapped species',
       title = 'Distribution of species maps by taxonomic group (class)')
print(bar_taxa2)
ggsave(filename = file.path(dir_fig, 'bar_by_taxa_all.png'), height = 4, width = 8)

taxa_df_long3 <- taxa_df %>%
  select(class, AquaMaps = n_am_class_solo_ass, IUCN = n_iucn_class_solo, Both = n_class_both) %>%
  gather(dataset, n_spp, AquaMaps:Both) %>%
  group_by(class) %>%
  mutate(n_total = sum(n_spp, na.rm = TRUE),
         percent = n_spp/n_total * 100)
bar_taxa3 <- ggplot(taxa_df_long3, aes(x = class, y = percent, fill = dataset)) +
  ggtheme_plot + 
  geom_bar(stat = 'identity', alpha = .5) +
  scale_fill_manual(values = c('AquaMaps' = 'red', 'Both' = 'mediumorchid', 'IUCN' = 'blue')) +
  geom_text(aes(label = sprintf('n = %s', n_total), y = .05), hjust = 0, size = 4, angle = 90, hjust = 1) +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) + 
  labs(x = 'Class', y = 'Percent of mapped and assessed species',
       title = 'Distribution of species maps by taxonomic group (class)')
print(bar_taxa3)
ggsave(filename = file.path(dir_fig, 'bar_by_taxa_assessed.png'), height = 4, width = 8)

```

***

``` {r, child = 'R/data_explore_global_maps.Rmd', eval = FALSE}
```

``` {r, child = 'R/data_explore_am_thresh.Rmd', eval = FALSE}
```

## Changes in SPP for scenarios

Comparing three scenarios to the current (v2015) global model for SPP:

* Baseline is: preferring IUCN rangemaps over Aquamaps rangemaps; Aquamaps 'presence' threshold is set at 40% or greater.
* Scenario 1: preferring IUCN rangemaps over Aquamaps; AM presence threshold set at >= 1% (effectively, all occurrences).
* Scenario 2: preferring AM maps over IUCN; AM presence >= 40%.
* Scenario 3: preferring AM over IUCN; AM presence >= 1%.

Boxplots:

![Absolute change in SPP score](figures/SPP_scenarios/boxplot_abs_diff.png)

![Relative change in SPP score](figures/SPP_scenarios/boxplot_rel_diff.png)