### 2015-12-18 R/data_explore_am_thresh.Rmd
adding ribbon plot for threshold

###(6) Looking at how many and what type (threat category) of species we are losing by setting a threshold of 0.4 on the aquamaps species

This looks at the full Aquamaps dataset... should it be limited to species with valid categories?

```{r aquamaps probabilities, warning = FALSE, message = FALSE, cache.lazy = FALSE, cache = TRUE, verbose = FALSE, results = FALSE, echo = FALSE, include = FALSE}

# get spp_cells raw data for aquamaps
dir_aquamaps <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO/raw/aquamaps_2014')

spp_cells <- read_csv(file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), 
                      col_types = '_ccd__', progress = TRUE)

# spp_cells <- fread(input = file.path(dir_aquamaps, 'tables/ohi_hcaf_species_native.csv'), verbose = FALSE, header = TRUE, 
#                   colClasses = c('NULL', 'character', NA, 'numeric', 'NULL', 'NULL')) #1 min 49 sec

setnames(spp_cells, c('SpeciesID', 'CsquareCode', 'probability'), c('species_id', 'csquare_code', 'probability'))

spp_cells <- as.data.frame(spp_cells)

spp <- fread(input = file.path(dir_aquamaps, 'tables/ohi_speciesoccursum.csv'), 
             verbose = FALSE, header = TRUE) %>%
  mutate(sciname = paste(Genus, Species, sep = ' '),
         iucn_code = str_replace(toupper(iucn_code), 'LR/', ''),
         iucn_code = str_replace(iucn_code, 'N.E.', 'NE'))
spp <- as.data.frame(spp)

```

```{r hist_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}

spp_df3 <- spp_cells %>%
  group_by(species_id) %>%
  summarize(cells_0.4 = sum(probability >= 0.4),
            cells_t = sum(probability >= 0)) %>%
  left_join(spp, by = c('species_id' = 'SPECIESID')) %>%
  mutate(prop = cells_0.4 / cells_t,
         iucn_code = ifelse(is.na(iucn_code), 'NE', iucn_code)) %>%
  filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')) %>%
  # filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')) %>%
  select(sciname, species_id, cells_0.4, cells_t, prop, iucn_code)

iucn_levels = c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')
spp_df3 <- spp_df3 %>% 
  transform(iucn_code = factor(iucn_code, levels = iucn_levels))

hist_area_vs_threshold <- ggplot(spp_df3, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3)))

ggsave('figures/AMThreshold.png', width = 6, height = 4)

print(hist_area_vs_threshold)

spp_df3a <- spp_df3 %>% 
  filter(!iucn_code %in% c('DD', 'NE'))

hist_area_vs_threshold2 <- ggplot(spp_df3a, aes(prop, fill = iucn_code)) + 
  ggtheme_plot + 
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 0.05, alpha = .7) + 
  scale_fill_manual(values = c('LC' = 'green4', 
                               'NT' = 'greenyellow', 
                               'VU' = 'orange', 
                               'EN' = 'orangered2', 
                               'CR' = 'red4',
                               'DD' = 'grey60',
                               'NE' = 'grey70')) +
  labs(x = 'Proportion of range area included when selecting a threshold of 0.4', 
       y = sprintf('Proportion of all aquamaps species (n = %s)', nrow(spp_df3a)))

ggsave('figures/AMThreshold2.png', width = 6, height = 4)

print(hist_area_vs_threshold2)

```

```{r scatter_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}

### create dataframe that gives me all species, their total number of cells, 
### total number of cells >= 0.4 prob, difference between the two and threat

threshold_df <- data.frame()
for (i in seq(0, 1, .05)) { # i = .05
  temp_df <- spp_cells %>%
    group_by(species_id) %>%
    summarize(thresh       = i,
              cells_thresh = sum(probability >= thresh),
              cells_total  = sum(probability >= 0),
              proportion   = cells_thresh/cells_total) %>%
    left_join(spp %>%
#              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE')), 
              filter(toupper(iucn_code) %in% c('LC', 'NT', 'VU', 'EN', 'CR', 'EX')), 
            by = c('species_id' = 'SPECIESID')) %>%
    select(sciname, species_id, thresh, cells_thresh, cells_total, proportion, iucn_code)

  threshold_df <- threshold_df %>%
    bind_rows(temp_df)
}

thresh_df_summary <- threshold_df %>%
  group_by(thresh) %>%
  summarize(mean_prop   = mean(proportion, na.rm = TRUE),
            median_prop = median(proportion, na.rm = TRUE),
            quant09 = quantile(proportion, .09),
            quant25 = quantile(proportion, .25),
            quant75 = quantile(proportion, .75),
            quant91 = quantile(proportion, .91))

scatter_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = mean_prop)) + 
  ggtheme_plot + 
  geom_point(aes(y = median_prop), color = 'grey60') +
  geom_point() + 
  geom_abline(slope = -1, intercept = 1, color = 'red') + 
  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
  ylim(0, 1) +
  labs(title = 'Mean (median) species range after presence threshold cutoff',
       x = 'Presence threshold cutoff', 
       y = 'Mean (median) relative species range')

ggsave('figures/scatter_AMThreshold.png', width = 6, height = 4)

print(scatter_area_vs_threshold)

```

```{r boxplot_lost_area_w_threshold, warning = FALSE, message = FALSE, echo = FALSE}

area_mean_model <- lm(mean_prop ~ thresh, data = thresh_df_summary)
area_median_model  <- lm(median_prop ~ thresh,  data = thresh_df_summary)

ribbon_area_vs_threshold <- ggplot(thresh_df_summary, aes(x = thresh, y = median_prop)) + 
  ggtheme_plot + 
#  geom_point(position = position_jitter(w = .1), aes(color = scenario), alpha = .5) +

  geom_ribbon(aes(ymax = quant91, ymin = quant09), alpha = .5, color = NA, fill = 'grey80') +
  geom_ribbon(aes(ymax = quant75, ymin = quant25), alpha = .5, color = NA, fill = 'grey70') +
  geom_point(color = 'grey20') + 
  geom_point(aes(y = mean_prop), color = 'darkred') + 
  geom_abline(intercept = area_mean_model[['coefficients']][1],
              slope     = area_mean_model[['coefficients']][2],
              color = 'red') + 
  geom_vline(x = 0.40, color = 'blue', linetype = 'dashed') +
  geom_vline(x = 0.50, color = 'blue', linetype = 'dashed') +
  geom_vline(x = 0.60, color = 'blue', linetype = 'dashed') +
  annotate('text', x = .40, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'Ocean Health Index: 40%') + 
  annotate('text', x = .50, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'MPA Gap Analysis: 50%') + 
  annotate('text', x = .60, y = .05, hjust = 0, vjust = -.5,
           angle = 90, size = 2, color = 'grey20',
           label = 'Marine mammals: 60%') + 
  annotate('text', x = .75, y = .85,
           size = 2, color = 'grey20',
           label = 'R^2 = .995') + 
  ylim(0, 1) +
  labs(title = 'Relative species range after presence threshold',
       x = 'Presence threshold cutoff', 
       y = 'Mean (median) relative species range')

print(ribbon_area_vs_threshold)

ggsave('figures/ribbon_AMThreshold.png', width = 6, height = 4)

```