### 2016-02-17 data_explore/data_prep.Rmd
Updating numbers etc in manuscript... done with "data poor" section?

---
title: 'Data Prep: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(readr)
library(data.table)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')
dir_data_remote <- file.path(dir_N, 'git-annex/globalprep/_raw_data')
scenario <- 'v2016'

if(basename(getwd()) != 'data_explore') setwd('~/github/IUCN-AquaMaps/data_explore')
# source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

Need to:

gather spp_all, spp_am, spp_iucn, spp_iucn_maps, and spp_am_raw into local data folder.

``` {r pull_files_to_local, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### pull spp_iucn_marine.csv from git-annex and bring to local repo
spp_iucn <- read_csv(file.path(dir_anx, scenario, 'int/spp_iucn_cleaned.csv'))
spp_iucn <- spp_iucn %>%
  dplyr::select(sciname, iucn_sid, iucn_category, 
                popn_trend, parent_sid, subpop_sid) %>%
  unique()
write_csv(spp_iucn, file.path(dir_data, 'spp_iucn.csv'))

### pull spp_iucn_maps_all.csv from git-annex and bring to local repo
spp_iucn_maps <- read_csv(file.path(dir_anx, scenario, 'int/spp_iucn_maps_all.csv')) %>%
  rename(iucn_sid = id_no)

spp_iucn_maps <- spp_iucn_maps %>%
  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
  unique() %>%
  arrange(iucn_sid)

### At this point, still some duplicated iucn_sids, due to several
### names matching a single ID.  Not sure which names might be in
### AquaMaps though, so leave 'em in.

write_csv(spp_iucn_maps, file.path(dir_data, 'spp_iucn_maps.csv'))

### pull cleaned spp_am data from git-annex and bring to local repo
spp_am <- read_csv(file.path(dir_anx, scenario, 'int/spp_am_cleaned.csv'))
write_csv(spp_am, file.path(dir_data, 'spp_am.csv'))

### pull raw aquamaps species list from git-annex and bring to local repo
spp_am_raw <- read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv'))
write_csv(spp_am_raw, file.path(dir_data, 'speciesoccursum.csv'))

```

```{r get_spp_all, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### get spp_all_complete from git-annex, bring to local repo.  The spp_all_complete
### file was created in ohiprep/globalprep/SPP_ICO/data_prep_SPP.R, running the
### function create_spp_master_lookup(source_pref = 'iucn', fn_tag = '_complete', reload = TRUE)
spp_all_remote <- read_csv(file.path(dir_anx, scenario, 'int/spp_all.csv'))
spp_all_remote_clean <- spp_all_remote %>%
  filter(!str_detect(spatial_source, 'subpop')) %>%
  filter(!str_detect(spatial_source,  'alias'))
### At this point, plenty of duplicates:
### * This code returns 984 duplicated names:
###     length(spp_all_remote_clean$sciname) - length(unique(spp_all_remote_clean$sciname))
### * some are AM subpops which were ignored in processing the remote spp_all file.
###   Note: All subpops have an identified parent, so can filter them out without
###   losing any species.
###   code to verify this:
###     z_parent <- spp_all_remote_clean %>% 
###       filter(!is.na(subpop_sid) & spatial_source == 'am')
###     z_subpop <- spp_all_remote_clean %>% 
###       filter(!is.na(parent_sid) & spatial_source == 'am')
###     nrow(z_subpop %>% filter(!sciname %in% z_parent$sciname))
### * some are duplicated IUCN parents (e.g. Dermochelys coriacea) that will be 
###   eliminated when select(-iucn_subpop) column and then unique()

### filter out spatial_source == 'am' that also have parent_sid (i.e.
### these are AM-mapped subpopulations); then: 
### select away the iucn_subpop, parent_sid, and subpop_sid (among others),
### these are creating false unique rows.
spp_all <- spp_all_remote_clean %>%
  filter(!(spatial_source == 'am' & !is.na(parent_sid))) %>%
  dplyr::select(am_sid, sciname, reviewed, iucn_sid, popn_trend,
                popn_category, category_score, spp_group, 
                id_no, spatial_source, name_verified) %>%
  unique()
### At this point, still some duped scinames - these are ones with multiple IUCN
### SIDs that are not designated as parents or subpops.  We will brute-force
### this by just eliminating the duplicated rows, keeping only the first instance.
### By arranging by iucn_sid, we will keep the lowest value, which seems to
### generally correspond to the parent...
spp_all <- spp_all %>%
  arrange(iucn_sid)
spp_all <- spp_all[!duplicated(spp_all$sciname), ]

length(spp_all$sciname) - length(unique(spp_all_remote_clean$sciname))
### no duped scinames - same length as original (cleaned) list

### But, some of the extracted IUCN maps don't appear on this list! add 'em in.
spp_iucn_maps_new <- read_csv(file.path(dir_data, 'spp_iucn_maps.csv')) %>%
  filter(!sciname %in% spp_all$sciname) %>%
  filter(!str_detect(sciname, 'subpop')) %>% ### a Hawaiian subpop... ditch it!
  select(-iucn_subpop)

spp_all1 <- spp_all %>%
  full_join(spp_iucn_maps_new)

### Also, some of the AM listings have disappeared due to name conflicts
### - aliases, etc.
spp_am_new <- read_csv(file.path(dir_data, 'speciesoccursum.csv')) %>%
  mutate(sciname = paste(str_trim(genus), str_trim(species)),
         iucn_sid = as.integer(iucn_id)) %>%
  dplyr::select(am_sid = speciesid, sciname, reviewed, iucn_sid,
                popn_category = iucn_code) %>%
  filter(!am_sid %in% spp_all$am_sid) %>%
  mutate(spatial_source = 'am',
         popn_category = ifelse(popn_category == 'LR/lc', 'LC', popn_category),
         popn_category = ifelse(popn_category == 'LR/nt', 'NT', popn_category),
         popn_category = ifelse(popn_category == 'N.E.',  'NE', popn_category))

### 545 new spp_am not in spp_all from remote source, by am_sid.
### 24 of these, however, list iucn_sids that match existing iucn_sid.
### * for species with iucn_sid that match, and no am_sid, add am_sid to 
###   existing record (use existing values for other fields; drop new values).
### * for species with iucn_sid that match, and with a different am_sid, no action.
### * for species with non-matching or NA iucn_sid, just add to list.
spp_am_new_iucnmatch <- spp_am_new %>% 
  filter(!is.na(iucn_sid) & iucn_sid %in% spp_all$iucn_sid) %>%
  dplyr::select(iucn_sid, am_sid_new = am_sid, reviewed_new = reviewed)
spp_all2 <- spp_all1 %>%
  left_join(spp_am_new_iucnmatch, by = 'iucn_sid') %>%
  mutate(am_sid   = ifelse(is.na(am_sid) & !is.na(am_sid_new), am_sid_new, am_sid),
         reviewed = ifelse(is.na(am_sid) & !is.na(am_sid_new), reviewed_new, reviewed)) %>%
  select(-am_sid_new, -reviewed_new)
spp_all3 <- spp_all2 %>%
  full_join(spp_am_new %>%
              filter(!iucn_sid %in% spp_am_new_iucnmatch$iucn_sid),
            by = c("am_sid", "sciname", "reviewed", "iucn_sid", "popn_category", "spatial_source"))

write_csv(spp_all3, file.path(dir_data, 'spp_all.csv'))

```

```{r get_other_vars, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}
### create spp_am_v_iucn: the whole maplist narrowed down to just the overlapping species.
### Overlaps detected from spatial_source containing 'iucn' (b/c source_pref = 'iucn' above),
### that also have a non-NA AquaMaps ID code.
spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))

spp_am_v_iucn <- spp_all %>%
  filter(str_detect(spatial_source, 'iucn') & !is.na(am_sid)) %>%
  filter(!is.na(id_no)) %>%
  filter(!str_detect(spp_group, 'MARINEFISH')) %>%
  unique()
# length(unique(spp_am_v_iucn$am_sid)); length(unique(spp_am_v_iucn$iucn_sid))
### NOTE: 2345 unique scientific names; but 2283 unique am_sid and 2249 unique iucn_sid.
### Some unique species (as identified by am_sid and iucn_sid) have multiple
### scinames.  At this point, not really important what the sciname is?
spp_am_v_iucn1 <- spp_am_v_iucn %>%
  group_by(am_sid, iucn_sid) %>%
  mutate(dupes = n() - 1,
         first_instance = sciname == first(sciname)) %>%
  filter(!(dupes > 0 & !first_instance)) %>%
  select(-dupes, -first_instance) %>%
  ungroup()
# length(unique(spp_am_v_iucn1$am_sid)); length(unique(spp_am_v_iucn1$iucn_sid))

### What species are represented by the duplicated IUCN ID numbers?  These
### are species that IUCN experts lumped together (via synonyms etc) 
### but are treated as separate by AquaMaps (the reverse does not seem
### to happen)...  There are 64 AM species IDs and only 30 IUCN species IDs.  
### How to treat in the mapping?
### * lump AM species together and plot against a single IUCN species map?
###   That would occur in the area determining function - group by iucn_sid
###   and map all am_sids against that iucn_sid
x <- spp_am_v_iucn1 %>%
  group_by(iucn_sid) %>%
  filter(n() > 1)
# length(unique(x$am_sid)); length(unique(x$iucn_sid))

write_csv(spp_am_v_iucn1, file.path(dir_data, 'spp_am_v_iucn.csv'))
```

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE}

spp_am_v_iucn <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
message('Building IUCN species to cell table.  This might take a few minutes.')
iucn_map_files <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
iucn_spp_cells <- lapply(iucn_map_files, data.table::fread) %>%
  bind_rows()
iucn_spp_cells <- iucn_spp_cells %>%
  rename(iucn_sid = id_no, loiczid = LOICZID) %>%
  mutate(iucn_sid = as.integer(iucn_sid),
         presence = as.integer(presence),
         loiczid  = as.integer(loiczid)) %>%
  filter(iucn_sid %in% spp_am_v_iucn$iucn_sid) %>%
  filter(presence != 5)
### length(unique(iucn_spp_cells$iucn_sid)) results in 2013 species found...
### some species with no cells reported.  2249 listed in the am v iucn file.
write_csv(iucn_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'))

message('Loading AquaMaps species to cell table.  This might take a few minutes.')
am_spp_cells <- fread(file.path(dir_data_remote, 'aquamaps/v2015/csv/hcaf_sp_native_trunc.csv'), 
                         showProgress = TRUE)
am_spp_cells <- am_spp_cells %>%
  rename(am_sid = speciesid, prob = probability) %>%
  filter(am_sid %in% spp_am_v_iucn$am_sid)
### length(unique(am_spp_cells$am_sid)) results in 2283 species found - great!
write_csv(am_spp_cells, file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'))

```

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
### Create files for species list with area info, both a complete list and a trimmed list

loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
loiczid_raster       <- raster::raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'

data_file <- file.path(dir_data, 'spp_list_w_area_all.csv')
reload <- FALSE
if (!file.exists(data_file) | reload) {
  ### species info list - scinames, IUCN species ID, AM species ID, category etc
  spp_list_base <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

  create_spp_list(spp_list_base, data_file, am_spp_cells, iucn_spp_cells, loiczid_raster)
}

spp_list <- read_csv(data_file) 
### At this point, spp_list zero-area IUCNs due to missing (?) polygons (maybe points?)

spp_list <- spp_list %>%
  mutate(sm_perc = ifelse(is.na(sm_perc), 0, sm_perc)) %>%
  filter(!area_iucn == 0)
  ### will ditch the ones where no IUCN polygon exists - or are these species
  ### whose ranges only occur in cells beyond the LOICZID grid?

### attach group name text field, and occurcells field
spp_list <- spp_list %>%
  left_join(x <- read_csv(file.path(dir_data, 'sciname_lookups/spp_groupnames.csv')),
            by = 'spp_group') %>%
  left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>% 
              select(am_sid = speciesid, occurcells),
            by = 'am_sid')

cat_score_df <- data.frame(popn_category  = c('LC', 'NT', 'VU', 'EN', 'CR', 'EX', 'DD', 'NE'),
                           category_score = c( 0.0,  0.2,  0.4,  0.6,  0.8,  1.0,   NA,  NA))
spp_list <- spp_list %>%
  left_join(cat_score_df, by = 'popn_category')

write_csv(spp_list, file.path(dir_data, 'spp_list_w_area_trimmed.csv'))

```

``` {r counts_spp_all_observations, echo = FALSE}
spp_all <- read_csv(file.path(dir_data, 'spp_all.csv'))

# library(rgbif)

# system.time({
#   x <- occ_search(scientificName = spp_all$sciname[1:100], fields = 'minimal', limit = 1)
#   ### seems linear - 10 names ~ 2.5 sec, 100 names ~ 25 sec, 1000 names, ~250 sec
#   ### so 30000 names ~ 7500 sec or 2 hours
# })
# y <- sapply(x, function(x) x$meta$count)
# gbif_occur <- data.frame(sciname = names(y), gbif_occ = y)
# 
# gbif_occur <- gbif_occur %>% 
#   left_join(spp_all %>%
#               select(iucn_sid, am_sid, sciname),
#             by = 'sciname') %>%
#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
#                        select(am_sid = speciesid, occurcells),
#                      by = 'am_sid')

# plot(ln_obs ~ ln_occurcells, data = gbif_occur)
# plot(obs ~ occurcells, data = gbif_occur)
# print(summary(lm(obs ~ occurcells, data = gbif_occur)))

# write_csv(gbif_occur, file.path(dir_data, 'gbif_occurrence.csv'))
gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))

# source('obis_count.R')
# scinames <- spp_all$sciname
# 
# obis_occur <- obis_count(scinames)
# obis_occur <- obis_count(scinames[26500:26600], verbose = TRUE)
# 
# obis_occur <- obis_occur %>% 
#   left_join(spp_all %>%
#               select(iucn_sid, am_sid, sciname),
#             by = 'sciname') %>%
#   left_join(read_csv(file.path(dir_data_remote, 'aquamaps/v2015/csv/speciesoccursum.csv')) %>%
#                        select(am_sid = speciesid, occurcells),
#                      by = 'am_sid')
# 
# write_csv(obis_occur, file.path(dir_data, 'obis_occurrence.csv'))
obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))

totes_occur <- obis_occur %>%
  left_join(gbif_occur) %>%
  rowwise() %>%
  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE))

```

__Possible "data poor" indicators:__

* (of matched species) how does bottom decile get distributed across quadrants
* use mean of OBIS and GBIF, and just report average (or median?) occurrences across all species in each quadrant
* also use similar method for AM occurcells

``` {r data_poor_analysis, echo = FALSE, eval = FALSE}
spp_quads <- read_csv(file.path(dir_data, 'spp_list_quads.csv')) %>% 
  select(am_sid, iucn_sid, occurcells, quad) %>%
  unique()
### This contains all mapped species, discarding those with no actual IUCN maps

spp_occ_counts <- spp_quads %>%
  left_join(totes_occur) %>%
  group_by(iucn_sid, am_sid, quad) %>%
  summarize(occurcells = mean(occurcells, na.rm = TRUE), occ = mean(mean_occ, na.rm = TRUE)) %>%
  group_by(iucn_sid, quad) %>%
  summarize(occurcells = sum(occurcells, na.rm = TRUE), occ = sum(occ, na.rm = TRUE)) %>%
  ungroup()

btm_dec_occurcells <- spp_occ_counts$occurcells %>% quantile(c(.1, .25))
btm_dec_occ <- spp_occ_counts$occ %>% quantile(c(.1, .25))

quad_count_means <- spp_occ_counts %>%
  mutate(quad = ifelse(quad == 'q4', 'q4', 'q123')) %>%
  group_by(quad) %>%
  summarize(mean_occurcells = mean(occurcells),
            mean_occ        = mean(occ),
            med_occurcells  = median(occurcells),
            med_occ         = median(occ),
            dp_am_10 = sum(occurcells <= btm_dec_occurcells[1])/n(),
            dp_am_25 = sum(occurcells <= btm_dec_occurcells[2])/n(),
            dp_gb_10 = sum(occ <= btm_dec_occ[1])/n(),
            dp_gb_25 = sum(occ <= btm_dec_occ[2])/n())

write_csv(quad_count_means, file.path(dir_data, 'quads_data_poor.csv'))

```