### 2016-02-25 mpa_analysis/marine_prep.Rmd

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(rgdal)
library(maptools)
library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

``` {r set_up_spatially_marine_raster}

### This raster (at 0.01 degree resolution) contains flags on whether a cell
### is "spatially marine" - i.e. overlaps (center?) with an ocean area.
### This is separate from the WDPA "marine" flag.

### as long as we're here, let's set up some objects for WGS 84 projection
p4s_wgs84 <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
crs_wgs84 <- CRS(p4s_wgs84)

rast_marine_file <- file.path(dir_git, 'mpa_analysis/rast_marine.tif')
reload <- FALSE
if(reload == TRUE) {
### import OHI regions polygons, to determine spatially-marine areas
  poly_marine_file <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters/poly_marine1.shp')
  if(!file.exists(poly_marine_file)) {
    dir_rgn   <- file.path(dir_N, 'git-annex/globalprep/spatial/v2015/data')
    lyr_rgn <- 'regions_gcs'
    poly_rgn <- readShapePoly(file.path(dir_rgn, lyr_rgn), 
                               proj4string = crs_wgs84)

    ### filter out all polygons that are land or landlocked, and flag 
    ### the rest as spatially-marine = 1
    poly_marine <- poly_rgn[!str_detect(poly_rgn@data$rgn_typ, 'land'), ]

    # writePolyShape(poly_marine, poly_marine_file)
    writeOGR(obj    = poly_marine,
             dsn    = path.expand(dirname(poly_marine_file)),
             layer  = str_replace(basename(poly_marine_file), '.shp', ''),
             driver = 'ESRI Shapefile',
             overwrite_layer = TRUE,
             verbose = TRUE)
  }

  library(gdalUtils)
  rast_marine <- gdal_rasterize(
      src_datasource = poly_marine_file,
      dst_filename   = rast_marine_file, 
        # destination for output
      a = 'rgn_id', 
        # the attribute in the shapefile to be assigned to the cell values
      te = c(-180, -90, 180, 90), 
        # extents for output raster
      tr = c(.01, .01),  
        # resolution for output raster
      tap = TRUE, 
        # target aligned pixels - align coords of extent of output to values of -tr
      a_nodata = NA, 
        # nodata value for raster; otherwise they will be filled in with zeroes
      output_Raster = TRUE, 
        # return output as a RasterBrick? 
      verbose = TRUE)
}

```

``` {r simplify_polys_function}
poly_simplify <- function(poly_set, thresh, verbose = FALSE) {
  if(class(poly_set) == 'list') poly_set <- poly_set[[1]]
#   if(verbose) {
#     message('Counting subpolys in poly set:')
#     poly_counts <- sapply(poly_set@polygons, function(x) length(x@Polygons))
#     print(poly_counts)
#   }

  ### filter out tiny sub-polygons -----
  ### create area list for filtering.  Area will be in square degrees... 
  polys_area <- lapply(poly_set@polygons, function(x) sapply(x@Polygons, function(y) y@area))

  if(verbose) {
    message('Polygon area quantiles:')
    print(quantile(unlist(polys_area)))
  }

  ### check to make sure threshold 
#   if(verbose) {
#     message('Smallest "main" subpolygon, i.e. smallest EEZ, in this set:')
#     poly_small <- min(unlist(lapply(polys_area, function(x) (max(x)))))
#     if(poly_small < thresh) message('Smallest poly smaller than threshold value of ', thresh)
#     for (j in 1:length(polys_area)) {
#       message(sprintf('rgn %s: largest polygon %.5f (%s)', poly_set@data$rgn_id[j], 
#                       max(polys_area[[j]]), poly_set@data$rgn_nam[j]))
#     }
#   }

  mainPolys <- lapply(polys_area, function(x) which(x > thresh))

  for(i in 1:length(mainPolys)) {
    if(length(mainPolys[[i]]) >= 1 && mainPolys[[i]][1] >= 1) {
      poly_set@polygons[[i]]@Polygons  <- poly_set@polygons[[i]]@Polygons[mainPolys[[i]]]
      poly_set@polygons[[i]]@plotOrder <- 1:length(poly_set@polygons[[i]]@Polygons)
    }
  }

  plot_order_max <- max(poly_set@plotOrder)
  poly_list_length <- length(poly_set@polygons)
  message('Plot order max: ', plot_order_max, '; length poly list: ', poly_list_length)

  if(plot_order_max != poly_list_length) {
    message('NOT SAME LENGTH?!')
    stop()
    # poly_set@plotOrder <- 1:length(poly_set@polygons)
    # is plotOrder necessary to be continuous?  that would explain this line.
  }

  if(verbose) {
#     message('Counting subpolys in modified poly set:')
#     poly_counts <- sapply(poly_set@polygons, function(x) length(x@Polygons))
#     print(poly_counts)

    polys_area_new <- lapply(poly_set@polygons, function(x) sapply(x@Polygons, function(y) y@area))

    message('Polygon area quantiles after adjustment:')
    print(quantile(unlist(polys_area_new)))
  }

  return(poly_set)
}

```