### 2016-03-01 mpa_analysis/mpa_gap_analysis.Rmd
let's get some of this up on line...

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

# MPA gap analysis

Several steps to this:

* set up WDPA database in a raster format for easy use
* perform gap analysis for WDPA-MPA data base (which year?) vs AquaMaps 2014, with 50% threshold - replicate the original analysis
* reperform the analysis with latest/greatest WDPA-MPA database, and AquaMaps 2015, with 50% threshold and 0% threshold
* perform the analysis with latest/greatest WDPA-MPA database, and IUCN shapefiles

## set up WDPA raster

* Create polygon subset of WDPA database
    * STATUS = 'Designated'
    * IUCN_CAT = I - IV
    * Simplify polygons to make it easier for rasterization
        * cut any main polygons less than 10% of a raster cell.  Raster
          cells are .01 x .01 degree, so area = .0001; cut any less than 1e-5.
          Do the same for subpolygons within a larger polygon.
    * Inverse order by `STATUS_YR`, so oldest is at the end (for `gdal_rasterize()`)
* To avoid double-counting for overlapping polygons - rasterize WDPA to fine scale first
    * less than 1km^2^ was mentioned in paper as resolution of MPA data.  
      Rasterize to 0.01 degree x 0.01 degree  
        * this results in max cell size of 1.11 km x 1.11 km at the equator.
    * Use STATUS_YR as field; see note above about ordering polygons
      for use in `gdal_rasterize()`

## Set up marine raster

* Create polygons subset of OHI regions, for only marine regions
* Rasterize to same resolution as above.

## Create raster of WDPA IUCN I-IV and spatially marine

* Intersect the two rasters so only protected areas within spatially marine
  areas are included.

## Aggregate protected area layer to LOICZID scale (0.5 deg)

* Use raster::extract? each LOICZID gets 2500 subcells noting NA (not
  protected) or a value (protected); turn this into proportion of
  LOICZID cell that is protected.
* Also need to do this to find the proportion of LOICZID cell that is
  spatially marine - if not fully marine, the protected proportion will
  be larger; but the area will be smaller (from AquaMaps, or recalc)
    * e.g. a single LOICZID is 60% marine (1500 marine cells), and 48%
      protected (1200 protected cells, which are by def marine).  So
      protected proportion = 48/60 = 80% protection of the marine area
      within this cell.

* AquaMaps layer: use as is
* To analyze for each species:
    * Summarize total area and total protected area (for a given protection scenario) for the species
    * Include species extinction risk in there too?

``` {r load_rasters}

dir_rast <- file.path(dir_anx, 'explore_am_v_iucn/mpa_rasters')
rast_mpa_mar_file <- file.path(dir_rast, 'rast_mpa_mar_i-iv.tif')
# rast_mpa_mar_file <- file.path(dir_rast, 'wdpa_all_mar_klein_i-iv.tif')
rast_mar_file <- file.path(dir_rast, 'rast_marine.tif')

if(!file.exists(rast_mpa_mar_file)) {
#   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa_i-iv-des.tif')
   rast_mpa_file <- file.path(dir_rast, 'WDPA_poly_Jan2015_mpa_i-iv-all.tif')

  rast_mar <- raster(rast_mar_file)
  rast_mpa <- raster(rast_mpa_file)

  rast_mpa_mar <- mask(rast_mpa, rast_mar, 
                       filename = rast_mpa_mar_file, 
                       overwrite = TRUE,
                       progress = 'text')
} else {
  rast_mar            <- raster(rast_mar_file)
  names(rast_mar)     <- 'rgn_id'
  rast_mpa_mar        <- raster(rast_mpa_mar_file)
  names(rast_mpa_mar) <- 'status_yr'
}

```

``` {r aggregate_to_loiczid}
rast_loiczid_file   <- file.path(dir_git, 'data/rasters/loiczid_raster.grd')
rast_loiczid        <- raster(rast_loiczid_file)
names(rast_loiczid) <- 'loiczid'

rast_loiczid_01_file <- file.path(dir_rast, 'rast_loiczid_01.tif')
if(!file.exists(rast_loiczid_01_file)) {
  rast_loiczid_01 <- resample(rast_loiczid, rast_mpa_mar, 
                            method = 'ngb',
                            progress = 'text',
                            filename = rast_loiczid_01_file)
} else {
  rast_loiczid_01 <- raster(rast_loiczid_01_file)
}
```

``` {r summarize counts of marine and MPAs by cell}

### Count marine cells in each LOICZID cell
mar_df_file <- file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv')
if(!file.exists(mar_df_file)) {
  is_marine_df <- zonal(rast_mar, 
                        z = rast_loiczid_01, 
                        fun = 'count',
                        progress = 'text') %>%
    as.data.frame() %>%
    rename(loiczid = zone)
  write_csv(is_marine_df, mar_df_file)
}

### Count all protected cells in each LOICZID cell
### NOTE: loses status_yr values - so filter by year for each iteration
yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')
mpa_cvs_files <- c(file.path(dir_git, 
                             sprintf('mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv', 
                                     yr_list)))
# mpa_cvs_files <- c(file.path(dir_git, 
#                              sprintf('mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv', 
#                                      yr_list)))
yr_list1 <- yr_list[!file.exists(mpa_cvs_files)]

if(length(yr_list1) > 0) {
  message('Creating marine protected area rasters for years:')
  message('  ', paste(yr_list1, collapse = ', '))
  for (yr in yr_list1) { # yr <- yr_list1[1]
    message('filtering MPAs to those designated prior to ', str_replace(yr, 'y', ''))
    rast_is_mpa <- rast_mpa_mar
    data_year <- as.integer(str_replace(yr, 'y', ''))
    values(rast_is_mpa)[values(rast_mpa_mar) >= data_year] <- NA
      ### NOTE: 2015 MPA dataset includes up to 2014; 2014 dataset goes
      ###   through 2013; etc.
    message('counting protected cells per LOICZID for ', yr)
    is_mpa_df <- zonal(rast_is_mpa, 
                       z = rast_loiczid_01, 
                       fun = 'count',
                       progress = 'text') %>%
      as.data.frame() %>%
      rename(loiczid = zone)
    write_csv(is_mpa_df, 
              mpa_cvs_files[yr_list == yr])
  }
}
### These files get loaded later; no need to hog memory by loading them here.

```

First step:

* Create data frame of LOICZID vs MPA coverage and marine area
    * note: using raster-generated marine area, not AquaMaps reported marine
      area - need to look into how different these are.
        * 7% of cells differ in calculated vs reported area
        * most of these differences cluster very near zero

For each species:

* for each "present" LOICZID cell
    * determine total area: overall cell area (in km2) * marine area (cells/2500)
    * determine protected area: overall cell area (in km2) * protected area (cells/2500)
* sum total area and protected area across all "present" LOICZID cells
* report: 
    * ratio of (protected area:total area) (dimensionless, percentage)
    * total marine area and total protected area (km2)
    * species name, ID, taxonomic group, extinction risk

``` {r combine_mpa_areas}

dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv/')
yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')

### get AquaMaps file of cells to translate cell IDs to area and ocean area
hcaf <- read_csv(file.path(dir_am_data, 'hcaf_truncated.csv')) %>%
  select(loiczid, cellarea, oceanarea)

### attach the AquaMaps areas to the calculated percentages of marine area
### note: AquaMaps calculated areas differ from my calculated areas - see below
cell_df <- read_csv(file.path(dir_git, 'mpa_analysis/celldata/loiczid_marine.csv'),
  col_types = 'dd') %>%
  rename(marine_ct = count) %>%
  left_join(hcaf, by = 'loiczid') %>%
  mutate(marine_pct  = marine_ct/2500,
         marine_area = marine_pct * cellarea) %>%
  select(-marine_ct)

### load each analysis year (in yr_list) and join them all together
### to note protection status year of each cell.

### Use this to point toward the 'i-iv-all' files (for all STATUS) or
### 'i-iv-des' (for STATUS == 'Designated')
mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_i-iv_%s.csv'
# mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_klein_i-iv_%s.csv'
# mpa_file_base <- 'mpa_analysis/celldata/loiczid_mpa_i-iv-des_%s.csv'

for (yr in yr_list) { # yr <- yr_list[1]
  fn <- file.path(dir_git, sprintf(mpa_file_base, yr))
  df <- read_csv(fn, col_types = 'dd')
  names(df)[2] <- yr
  cell_df <- cell_df %>%
    left_join(df, by = 'loiczid')
  cell_df[ , paste(yr, '_pct', sep = '')]  <- cell_df[ , yr]/2500
  cell_df[ , paste(yr, '_area', sep = '')] <- cell_df[ , paste(yr, '_pct', sep = '')] * cell_df$cellarea

  ### remove intermediate columns
  cell_df[ , yr] <- NULL
  cell_df[ , paste(yr, '_pct', sep = '')] <- NULL
}

# ### check diffs in oceanarea vs marine_area
# x <- cell_df %>% 
#   filter(oceanarea != marine_area) %>%
#   select(loiczid, oceanarea, marine_area, cellarea)
# ### 18158 cells out of 259200 = 7%
# x <- x %>% mutate(pct_diff = (marine_area - oceanarea)/cellarea)
# hist(x$pct_diff)
# ### most differences cluster right around zero.  Not going to worry about it!

```

# recreate Klein results using aquamaps 2014 data

``` {r recreate Klein results}

area_mpa <- cell_df %>%
  summarize(marine_area  = sum(marine_area), 
            mpa_tot_1990 = sum(y1990_area), 
            mpa_tot_1995 = sum(y1995_area), 
            mpa_tot_2000 = sum(y2000_area), 
            mpa_tot_2005 = sum(y2005_area), 
            mpa_tot_2010 = sum(y2010_area), 
            mpa_tot_2014 = sum(y2014_area), 
            mpa_tot_2015 = sum(y2015_area))

### Recreate Klein results, using 2014 AquaMaps
mpa_klein_am2014 <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc_2014.csv'))
mpa_klein_am2014 <- mpa_klein_am2014 %>%
  left_join(cell_df, by = 'loiczid')

### filter to the appropriate threshold...
thresh <- 0.50

mpa_klein_am2014 <- mpa_klein_am2014 %>%
  filter(probability >= thresh)

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
mpa_klein_am2014 <- mpa_klein_am2014 %>%
  group_by(speciesid) %>%
  summarize(marine_range = sum(marine_area),
            mpa2014 = sum(y2014_area),
            pct2014 = mpa2014/marine_range)

mpa_klein_am2014 <- mpa_klein_am2014 %>%
  mutate(cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014))

mpa_klein_orig <- mpa_klein_am2014 %>%
  group_by(cat2014) %>%
  summarize(n_tot = nrow(mpa_klein_am2014), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

### Not filtering on status, but all in shapefile are status == 'Designated'
#   cat2014 n_tot n_cat pct_cat
#       gap 17344   282     1.6
#      0-2% 17344  7728    44.6
#      2-5% 17344  7738    44.6
#     5-10% 17344  1188     6.8
#      >10% 17344   408     2.4

```

# run analysis using 2015 aquamaps

``` {r apply loiczid area df to species list}

### Assemble the AM species-cell df with info on marine area and 
### protected area per cell
am_spp_cells <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv'))
am_spp_cells <- am_spp_cells %>%
  left_join(cell_df, by = 'loiczid')

### filter to the appropriate threshold...
thresh <- 0.50

am_spp_cells_filt <- am_spp_cells %>%
  filter(probability >= thresh)

### group by speciesid, then summarize sum(marine area), 
### sum(protected area), ratio
am_spp_mpas_filt <- am_spp_cells_filt %>%
  group_by(speciesid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa1990 = sum(y1990_area),
            mpa1995 = sum(y1995_area),
            mpa2000 = sum(y2000_area),
            mpa2005 = sum(y2005_area),
            mpa2010 = sum(y2010_area),
            mpa2014 = sum(y2014_area),
            mpa2015 = sum(y2015_area),
            pct1990 = mpa1990/marine_range,
            pct1995 = mpa1995/marine_range,
            pct2000 = mpa2000/marine_range,
            pct2005 = mpa2005/marine_range,
            pct2010 = mpa2010/marine_range,
            pct2014 = mpa2014/marine_range,
            pct2015 = mpa2015/marine_range)
```

``` {r redo klein for new AquaMaps with 2015 mpas}

area_mpa <- cell_df %>%
  summarize(marine_area  = sum(marine_area), 
            mpa_tot_1990 = sum(y1990_area), 
            mpa_tot_1995 = sum(y1995_area), 
            mpa_tot_2000 = sum(y2000_area), 
            mpa_tot_2005 = sum(y2005_area), 
            mpa_tot_2010 = sum(y2010_area), 
            mpa_tot_2014 = sum(y2014_area), 
            mpa_tot_2015 = sum(y2015_area))

#   marine_area mpa_tot_1990 mpa_tot_1995 mpa_tot_2000 mpa_tot_2005 mpa_tot_2010 mpa_tot_2014 mpa_tot_2015
#         (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)        (dbl)
# 1   362657307     353355.3     534137.2     750217.9     935137.1      1140272      2888600      2896875

### Redo analysis using 50% threshold, but new AquaMaps database
### * select just 2014 and 2015, at 50% threshold, and group by
###   same categories that the original paper did.
mpa_klein_am2015 <- am_spp_mpas_filt %>%
  select(speciesid, marine_range, mpa2014, mpa2015, pct2014, pct2015)

mpa_klein_am2015 <- mpa_klein_am2015 %>%
  mutate(cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014),
         cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

mpa_klein_sum2014 <- mpa_klein_am2015 %>%
  group_by(cat2014) %>%
  summarize(n_tot = nrow(mpa_klein_am2015), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

mpa_klein_sum2015 <- mpa_klein_am2015 %>%
  group_by(cat2015) %>%
  summarize(n_tot = nrow(mpa_klein_am2015), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

# > mpa_klein_sum2014
#   cat2014 n_tot n_cat pct_cat
#       gap 22885   797     3.5
#      0-2% 22885 11038    48.2
#      2-5% 22885  8713    38.1
#     5-10% 22885  1650     7.2
#      >10% 22885   687     3.0
# 
# > mpa_klein_sum2015
#   cat2015 n_tot n_cat pct_cat
#       gap 22885   797     3.5
#      0-2% 22885 10990    48.0
#      2-5% 22885  8747    38.2
#     5-10% 22885  1661     7.3
#      >10% 22885   690     3.0

```

``` {r check results for 1990 through 2010 at 50%}
mpa_spp <- am_spp_mpas_filt %>%
  select(speciesid, marine_range, mpa1990:mpa2010, pct1990:pct2010)

mpa_spp <- mpa_spp %>%
  mutate(cat1990 = ifelse(pct1990 == 0,                       'gap', NA),
         cat1990 = ifelse(pct1990 > 0    & pct1990 <= 0.02,  '0-2%', cat1990),
         cat1990 = ifelse(pct1990 > 0.02 & pct1990 <= 0.05,  '2-5%', cat1990),
         cat1990 = ifelse(pct1990 > 0.05 & pct1990 <= 0.10, '5-10%', cat1990),
         cat1990 = ifelse(pct1990 > 0.10,                    '>10%', cat1990),
         cat1995 = ifelse(pct1995 == 0,                       'gap', NA),
         cat1995 = ifelse(pct1995 > 0    & pct1995 <= 0.02,  '0-2%', cat1995),
         cat1995 = ifelse(pct1995 > 0.02 & pct1995 <= 0.05,  '2-5%', cat1995),
         cat1995 = ifelse(pct1995 > 0.05 & pct1995 <= 0.10, '5-10%', cat1995),
         cat1995 = ifelse(pct1995 > 0.10,                    '>10%', cat1995),
         cat2000 = ifelse(pct2000 == 0,                       'gap', NA),
         cat2000 = ifelse(pct2000 > 0    & pct2000 <= 0.02,  '0-2%', cat2000),
         cat2000 = ifelse(pct2000 > 0.02 & pct2000 <= 0.05,  '2-5%', cat2000),
         cat2000 = ifelse(pct2000 > 0.05 & pct2000 <= 0.10, '5-10%', cat2000),
         cat2000 = ifelse(pct2000 > 0.10,                    '>10%', cat2000),
         cat2005 = ifelse(pct2005 == 0,                       'gap', NA),
         cat2005 = ifelse(pct2005 > 0    & pct2005 <= 0.02,  '0-2%', cat2005),
         cat2005 = ifelse(pct2005 > 0.02 & pct2005 <= 0.05,  '2-5%', cat2005),
         cat2005 = ifelse(pct2005 > 0.05 & pct2005 <= 0.10, '5-10%', cat2005),
         cat2005 = ifelse(pct2005 > 0.10,                    '>10%', cat2005),
         cat2010 = ifelse(pct2010 == 0,                       'gap', NA),
         cat2010 = ifelse(pct2010 > 0    & pct2010 <= 0.02,  '0-2%', cat2010),
         cat2010 = ifelse(pct2010 > 0.02 & pct2010 <= 0.05,  '2-5%', cat2010),
         cat2010 = ifelse(pct2010 > 0.05 & pct2010 <= 0.10, '5-10%', cat2010),
         cat2010 = ifelse(pct2010 > 0.10,                    '>10%', cat2010))

mpa_spp_1990 <- mpa_spp %>%
  group_by(cat1990) %>%
  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

mpa_spp_1995 <- mpa_spp %>%
  group_by(cat1995) %>%
  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

mpa_spp_2000 <- mpa_spp %>%
  group_by(cat2000) %>%
  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

mpa_spp_2005 <- mpa_spp %>%
  group_by(cat2005) %>%
  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

mpa_spp_2010 <- mpa_spp %>%
  group_by(cat2010) %>%
  summarize(n_tot = nrow(mpa_spp), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

#   cat1990 n_tot n_cat pct_cat
#       gap 22889   925     4.0
#      0-2% 22889 21569    94.2
#      2-5% 22889   376     1.6
#     5-10% 22889    18     0.1
#      >10% 22889     1     0.0
# 
#   cat1995 n_tot n_cat pct_cat
#       gap 22889   878     3.8
#      0-2% 22889 21111    92.2
#      2-5% 22889   817     3.6
#     5-10% 22889    70     0.3
#      >10% 22889    13     0.1
#
#   cat2000 n_tot n_cat pct_cat
#       gap 22889   801     3.5
#      0-2% 22889 20973    91.6
#      2-5% 22889  1004     4.4
#     5-10% 22889    92     0.4
#      >10% 22889    19     0.1
#
#   cat2005 n_tot n_cat pct_cat
#       gap 22889   608     2.7
#      0-2% 22889 19828    86.6
#      2-5% 22889  2174     9.5
#     5-10% 22889   205     0.9
#      >10% 22889    74     0.3
# 
#   cat2010 n_tot n_cat pct_cat
#       gap 22889   578     2.5
#      0-2% 22889 18670    81.6
#      2-5% 22889  3278    14.3
#     5-10% 22889   288     1.3
#      >10% 22889    75     0.3

argh <- am_spp_mpas_filt %>%
  select(-(mpa1990:mpa2015)) %>%
  gather(key = year, value = pct, pct1990:pct2015) %>%
  mutate(year = str_replace(year, 'pct', '')) %>%
  group_by(year)

hist_plot <- ggplot(data = argh, aes(x = pct, color = year)) + 
  geom_density() +
  scale_x_continuous(limits = c(0, .05))
#  scale_x_log10()

hist_plot
ggsave(file.path(dir_git, 'figures/mpa_analysis_dist.png'))
```

### Original study (Klein 2015) vs new analyses

| study           | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :-------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015| IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, mpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| .               | .                  | .            | .    | .     | .     | .     | .    |
| am2015, mpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 2.4% | 50.3% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.4% | 50.4% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2010 | IUCN I-IV, Sp. Mar |   1,137,142  | 2.5% | 81.6% | 14.3% |  1.3% | 0.3% |
| am2015, mpa2005 | IUCN I-IV, Sp. Mar |     932,817  | 2.7% | 86.6% |  9.5% |  0.9% | 0.3% |
| am2015, mpa2000 | IUCN I-IV, Sp. Mar |     748,172  | 3.5% | 91.6% |  4.4% |  0.4% | 0.1% |
| am2015, mpa1995 | IUCN I-IV, Sp. Mar |     532,098  | 3.8% | 92.2% |  3.6% |  0.3% | 0.1% |
| am2015, mpa1990 | IUCN I-IV, Sp. Mar |     351,592  | 4.0% | 94.2% |  1.6% |  0.1% | 0.0% |

max protected species = Fis-160882

``` {r iucn_only}
### compare results using only IUCN species maps
dir_iucn_data <- '/var/data/ohi/git-annex/globalprep/_raw_data/iucn_spp'
iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_2015.csv')
iucn_spp_cells <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
  as.data.frame(stringsAsFactors = FALSE)

# iucn_spp_cells <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv'))
iucn_spp_cells <- iucn_spp_cells %>%
  left_join(cell_df, by = 'loiczid')

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
iucn_spp_sum <- iucn_spp_cells %>%
  group_by(iucn_sid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

iucn_spp_sum <- iucn_spp_sum %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

iucn_spp_sum2 <- iucn_spp_sum %>%
  group_by(cat2015) %>%
  summarize(n_tot = nrow(iucn_spp_sum), n_cat = n(), pct_cat = round(n_cat/n_tot * 100, 1))

```

``` {r combine IUCN and AM with 0% thresh, AM on top}

### use iucn_spp_sum from above, and create am_spp_sum for 0% thresh.
iucn_spp_sum

am_spp_mpas_0 <- am_spp_cells %>%
  group_by(speciesid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

am_spp_sum <- am_spp_sum %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

# combine by bringing in species list with IUCN SID mapped to AM SID;
# use all with AM SID, and then add in IUCN spp that are not mapped to an AM SID

```

``` {r get_iucn_cells}
reload = TRUE
  dir_iucn_data <- '/var/data/ohi/git-annex/globalprep/_raw_data/iucn_spp'
  iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_2015.csv')
  if(!file.exists(iucn_cells_file) | reload) {
    message(sprintf('Building IUCN species to cell table.  This might take a few minutes.\n'))
    iucn_map_files      <- file.path(dir_anx, 'iucn_intersections', list.files(file.path(dir_anx, 'iucn_intersections')))
    iucn_cells_spp_list <- lapply(iucn_map_files, read.csv) # read each into dataframe within a list
    iucn_cells_spp      <- bind_rows(iucn_cells_spp_list)   # combine list of dataframes to single dataframe
    # This creates a full data frame of all IUCN species, across all species groups, for all cells.
    # Probably big...
    names(iucn_cells_spp) <- tolower(names(iucn_cells_spp))
    write_csv(iucn_cells_spp, iucn_cells_file)
  } else {
    iucn_cells_spp <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
      as.data.frame(stringsAsFactors = FALSE)
  }

}
```