### 2016-03-03 ms_5_refs.Rmd

# References

___still need to check citations, format properly, put into order___

<!--
Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed.

References. References must be in PNAS style. Only published or in-press papers and books may be cited in the reference list. For references to in-press articles, please confirm with the cited journal that the article is in press and include a DOI number and online publication date. Unpublished abstracts of papers presented at meetings or references to "data not shown" are not permitted. References should be cited in numerical order as they appear in text, and all references, including for the SI, should be included in the main manuscript file. Because tables and figures will be inserted in the text where first cited, references in these sections should be numbered accordingly. Include the full title for each cited article. All authors (unless there are more than five) should be named in the citation. If there are more than five, list the first author's name followed by et al. Provide inclusive page ranges for journal articles and book chapters. Cite databases in the text or as footnotes.

Journal articles are cited as follows:

10. Neuhaus J-M, Sitcher L, Meins F, Jr, Boller T (1991) A short C-terminal sequence is necessary and sufficient for the targeting of chitinases to the plant vacuole. Proc Natl Acad Sci USA 88(22):10362–10366.
-->

__Key AquaMaps publications:__

Kaschner, K., R. Watson, A. W. Trites, D. Pauly (2006). Mapping world-wide distributions of marine mammal species using a relative environmental suitability (RES) model.  Marine Ecology Progress Series 316: 285–310. _check this citation journal name..._ _This outlines the basic RES methodology - AM development_

Kaschner, K., D.P. Tittensor, J. Ready, T Gerrodette and B. Worm (2011). Current and Future Patterns of Global Marine Mammal Biodiversity. PLoS ONE 6(5): e19653. PDF _just what the title says - AM development, presence threshold 60%, also analyzes richness as a function of threshold_

Ready, J., K. Kaschner, A.B. South, P.D Eastwood, T. Rees, J. Rius, E. Agbayani, S. Kullander and R. Froese (2010). Predicting the distributions of marine organisms at the global scale. Ecological Modelling 221(3): 467-478. PDF _Presents AM; assessing AquaMaps against other presence-only species models_

__Papers based on AquaMaps:__

Jones, M.C., S.R. Dyeb, J.K. Pinnegar and W.W.L. Cheung (2012). Modelling commercial fish distributions: Prediction and assessment using different approaches. Ecological Modelling 225(2012): 133-145. PDF _comparison of species distribution models including AquaMaps, Maxent and the Sea Around Us Project_

Coll, M., C. Piroddi, J. Steenbeek, K. Kaschner, F. Ben Rais Lasram et al. (2010). The biodiversity of the Mediterranean Sea: estimates, patterns, and threats. PLoS ONE 5(8): e11842. PDF _used AquaMaps to predict Med biodiversity.  Also: Threshold = 0._

Martin C.S., Fletcher R., Jones M.C., Kaschner K., Sullivan E., Tittensor D.P., Mcowen C., Geffert J.L., van Bochove J.W., Thomas H., Blyth S., Ravillious C., Tolley M., Stanwell-Smith D. (2014). Manual of marine and coastal datasets of biodiversity importance. May 2014 release. Cambridge (UK): UNEP World Conservation Monitoring Centre. 28 pp. (+ 4 annexes totalling 174 pp. and one e-supplement). PDF _review of marine data sets and data gaps etc, incl both IUCN and AM as well as many others_

-----

Hurlbert 2007 Species richness, hotspots, and the scale dependence of range maps in ecology and conservation. _mostly rasters of range maps?  "The scale dependence of range-map accuracy poses clear limitations on braod-scale ecological analyses and conservation assessments.  ... we provide guidance about the approriate scale of their use_

Jetz 2008 Ecological Correlates and Conservation Implications of Overestimating Species Geographic Ranges _EOO maps are usually highly interpolated and overestimate small-scale occurrence, which may bias research outcomes_

Pimm 2014 The biodiversity of species and their rates of extinction, distribtuion, and protection. _uses range maps to show biodiversity areas; may use IUCN range maps.  Also discusses gaps and possible things that can be done about them._

Rondinini 2006 Tradeoffs of different types of species occurrence data for use in systematic conservation planning _compares point locality, range maps, and distribution models in terms of omission and commission errors; also outlines Extent of Occurrence and Area of Occupancy distinctions._

García Molinos, Jorge, Benjamin S. Halpern, David S. Schoeman, Christopher J. Brown, Wolfgang Kiessling, Pippa J. Moore, John M. Pandolfi, Elvira S. Poloczanska, Anthony J. Richardson, and Michael T. Burrows. “Climate Velocity and the Future Global Redistribution of Marine Biodiversity.” Nature Climate Change advance online publication (August 31, 2015). doi:10.1038/nclimate2769.

Halpern, Benjamin S., Catherine Longo, Darren Hardy, Karen L. McLeod, Jameal F. Samhouri, Steven K. Katona, Kristin Kleisner, et al. “An Index to Assess the Health and Benefits of the Global Ocean.” Nature 488, no. 7413 (August 30, 2012): 615–20. doi:10.1038/nature11397.

Klein, Carissa J., Christopher J. Brown, Benjamin S. Halpern, Daniel B. Segan, Jennifer McGowan, Maria Beger, and James E.M. Watson. “Shortfalls in the Global Protected Area Network at Representing Marine Biodiversity.” Scientific Reports 5 (December 3, 2015): 17539. doi:10.1038/srep17539.

Selig, Elizabeth R., Catherine Longo, Benjamin S. Halpern, Benjamin D. Best, Darren Hardy, Cristiane T. Elfes, Courtney Scarborough, Kristin M. Kleisner, and Steven K. Katona. “Assessing Global Marine Biodiversity Status within a Coupled Socio-Ecological Perspective.” PLoS ONE 8, no. 4 (April 11, 2013): e60284. doi:10.1371/journal.pone.0060284.

* IUCN reference:
    * https://www.conservationtraining.org/course/view.php?id=217&lang=en
    * http://www.iucnredlist.org/technical-documents/red-list-training/iucnspatialresources
    * IUCN Red List - accessed 12/21/2015

* AquaMaps reference, and accessed date?

Predicting the distributions of marine organisms at the global scale
Ready, J., Kaschner, K., South, A. B., Eastwood, P. D., Rees, T., Rius, J., ... & Froese, R. (2010). Predicting the distributions of marine organisms at the global scale. Ecological Modelling, 221(3), 467-478.
Cited by 75

[PDF] Mapping world-wide distributions of marine mammal species using a relative environmental suitability (RES) model
K Kaschner, R Watson, AW Trites, D Pauly - Marine Ecology Progress …, 2006 - vliz.be
ABSTRACT: The lack of comprehensive sighting data sets precludes the application of 
standard habitat suitability modeling approaches to predict distributions of the majority of 
marine mammal species on very large scales. As an alternative, we developed an ...
Cited by 167 Related articles All 12 versions Web of Science: 86 Cite Save

[DOC] from aquamaps.org
[DOC] AquaMaps
K Kaschner, J Rius-Barile, K Kesner-Reyes… - Predicted range maps …, 2013 - aquamaps.org
AquaMaps is an approach to generating model-based, large-scale predictions of currently 
known natural occurrence of marine species. Models are constructed from estimates of the 
environmental tolerance of a given species with respect to depth, salinity, temperature, ...
Cited by 64 Related articles All 3 versions Cite Save More

[HTML] Current and future patterns of global marine mammal biodiversity
K Kaschner, DP Tittensor, J Ready, T Gerrodette… - PLoS …, 2011 - journals.plos.org
Abstract Quantifying the spatial distribution of taxa is an important prerequisite for the 
preservation of biodiversity, and can provide a baseline against which to measure the 
impacts of climate change. Here we analyse patterns of marine mammal species richness ...
Cited by 73 Related articles All 20 versions Web of Science: 34 Cite Save More