### 2016-03-21 data_explore/data_explore_global_maps.Rmd
incorporated Ben's comments; reframing a conclusion; preparing figures.  Moved the MPA analysis explanation of shifts into the SI

###Global representation of species

```{r setup, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

# Libraries and Paths

library(ggplot2)
library(maps)
library(readr)
library(data.table)
library(raster)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'data_explore') setwd('data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

By counting the number of unique species are found in each of the half degree cells we can then create a single global raster of number of species.

```{r read_am_and_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
iucn_spp_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           progress = TRUE,
                           col_types = 'cddddc')

am_spp_cells   <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/am_spp_cells_2015.csv'), 
                           progress = TRUE)

### LOICZID cell ID numbers - used in plotting rasters
loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'
```

```{r global_maps, echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE}

library(rasterVis)

myTheme <- rasterTheme(region = brewer.pal('Blues', n = 9))

message('Full AM dataset, species count\n')

all_am_rast_file   <- file.path(dir_data, 'rasters/am_all_rast.tif')

dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')

if(!file.exists(all_am_rast_file)) {
  am_spp_cells_big <- fread(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) %>%
    as.data.frame(stringsAsFactors = FALSE)

  all_am <- am_spp_cells_big %>%
    group_by(loiczid) %>%
    summarize(all_count = n()) %>%
    as.data.frame()

  all_am_map <- subs(loiczid_raster, all_am[ , c('loiczid', 'all_count')], 
                     by = 'loiczid', 
                     which = 'all_count', 
                     subsWithNA = TRUE)

  writeRaster(all_am_map, filename = all_am_rast_file, overwrite = TRUE)

} else {
  all_am_map   <- raster(all_am_rast_file)
}

### set up boundaries - from http://oscarperpinan.github.io/rastervis/FAQ.html#sec-7
ext <- as.vector(extent(all_am_map))

boundaries <- maps::map('world', fill=TRUE,
    xlim=ext[1:2], ylim=ext[3:4],
    plot=FALSE)

## read the map2SpatialPolygons help page for details
IDs <- sapply(strsplit(boundaries$names, ":"), function(x) x[1])
bPols <- maptools::map2SpatialPolygons(boundaries, IDs = IDs,
                              proj4string = CRS(projection(all_am_map)))

### create plot, add boundary polygons
am_levelplot1 <- levelplot(all_am_map, 
                           scales = list(at = seq(-180, 180, 45)),
                           xlab = NULL, ylab = NULL,
                           par.settings = myTheme) + 
  layer(sp.polygons(bPols, fill = 'grey90', col = 'grey30', lwd = .25))

png(file.path(dir_fig, 'all_aquamaps_map1.png'),
    width = 8.7, height = 6, units = 'in', res = 300)
print(am_levelplot1)
dev.off()

```

Now look at full IUCN species

```{r all_iucn, eval = FALSE, echo = FALSE}

message('Full IUCN dataset, species count\n')

all_iucn_rast_file   <- file.path(dir_data, 'rasters/iucn_all_rast.tif')
dir_data_iucn <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp')

if(!file.exists(all_iucn_rast_file)) {
  iucn_cells_file    <- file.path(dir_data_iucn, 'iucn_cells_2015.csv')
  iucn_spp_cells_all <- fread(iucn_cells_file) %>%
        as.data.frame(stringsAsFactors = FALSE)

  all_iucn <- iucn_spp_cells_all %>%
    group_by(loiczid) %>%
    summarize(all_count = n()) %>%
    as.data.frame()

  all_iucn_map <- subs(loiczid_raster, all_iucn[ , c('loiczid', 'all_count')], 
                     by = 'loiczid', 
                     which = 'all_count', 
                     subsWithNA = TRUE)

  writeRaster(all_iucn_map, filename = all_iucn_rast_file, overwrite = TRUE)

} else {
  all_iucn_map   <- raster(all_iucn_rast_file)
  values(all_iucn_map)[is.na(values(all_am_map))] <- NA
}

iucn_levelplot1 <- levelplot(all_iucn_map, 
                             scales = list(at = seq(-180, 180, 45)),
                             xlab = NULL, ylab = NULL,
                             par.settings = myTheme) + 
  layer(sp.polygons(bPols, fill = 'grey90', col = 'grey30', lwd = .25))

png(file.path(dir_fig, 'all_iucn_map1.png'),
    width = 8.7, height = 6, units = 'in', res = 300)
print(iucn_levelplot1)
dev.off()

```

``` {r variations_on_gl_dist_maps, echo = FALSE, eval = FALSE}
### rescale values relative to maximum cell value

am_map_max <- maxValue(all_am_map)
all_am_pct_map <- all_am_map/am_map_max

iucn_map_max <- maxValue(all_iucn_map)
all_iucn_pct_map <- all_iucn_map/iucn_map_max

diff_pct_map    <- all_am_pct_map - all_iucn_pct_map

### plot maps with distributions as percentages.  Should it be 0 - 1, or 
### 0 - max value?
levelplot(all_am_pct_map, par.settings = myTheme, 
          main = 'Aquamaps Dataset percent of max')
levelplot(all_iucn_pct_map, par.settings = myTheme, 
          main = 'IUCN Dataset percent of max')
### this is the difference one. Should it be min value to max value?
levelplot(diff_pct_map, par.settings = myTheme, 
          main = 'Difference: AquaMaps - IUCN by percent')

diff_ln_map <- all_am_ln_map - all_iucn_ln_map
levelplot(diff_ln_map, 
          par.settings = myTheme, 
          pretty = TRUE,
          margin = list(axis = TRUE),
#          main = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', 
          colorkey = list(at = seq(-1, 1, length = 15)))
maxValue(all_am_ln_map)
minValue(all_am_ln_map)
maxValue(all_iucn_ln_map)
minValue(all_iucn_ln_map)
maxValue(diff_ln_map)
minValue(diff_ln_map)
diff_ln_pts <- as.data.frame(rasterToPoints(diff_ln_map))

spp_richness_diff_plot <- ggplot(diff_ln_pts, aes(x = x, y = y)) +
  ggtheme_map + 
  geom_raster(aes(fill = layer)) +
  scale_fill_gradient2(low = 'red', mid = 'white', high = 'blue', midpoint = 0) +
  borders('world', color='gray40', fill='gray45', size = .1) +  # create a layer of borders
  scale_x_continuous(breaks = seq(-180, 180, by = 30), expand = c(0, 2)) +
  scale_y_continuous(breaks = seq( -90,  90, by = 30), expand = c(0, 2)) +
  labs(title = 'Difference in relative species richness (log-transformed): AquaMaps - IUCN', x = NULL, y = NULL)

spp_richness_diff_plot

```