### 2016-03-21 ms_2_intro_res_disc.Rmd

---
output: html_document
---

| Section                   |   Word #   | Char # (w/spc) |
| :------                   |    ------: |    ------: |
| Introduction              |  512 words |  3411 char |
| Results/Discussion        |  802 words |  5410 char |
| Implications (incomplete) |  532 words |  3334 char |
| Conclusions               |  390 words |  2813 char |
| Methods (incomplete)      |  612 words |  3889 char |
| total body                | 2848 words | 18857 char |

# Introduction

Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
 understanding species ranges and diversity,
 predicting species responses to human impacts and climate change,
 and managing and protecting species effectively.
A rich literature tackles the many dimensions of these questions.

One major outcome of this body of science is the various compiled databases of species distribution maps. Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps (Kaschner et al. ___2013???___) and International Union for Conservation of Nature (IUCN) (IUCN 2015).  These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status (Halpern et al. 2012, 2015; Selig et al. 2013), evaluating global biodiversity patterns (Coll et al. 2010; Martin et al. 2014, Pimm et al. 2014, Kaschner et al. 2011), predicting range shifts (Molinos et al. 2015), and setting conservation priorities (Klein et al. 2015).

<!--
Three foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; Ready et al. 2010) have been cited a total of 315 times.
-->

The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.  Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.  Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans (Rondinini et al. 2006, Jetz et al. 2008).

To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of species.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).  For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.

We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index (Halpern et al. 2012, 2015) and a global analysis of gaps in protection afforded by marine protected areas (MPAs) (Klein et al. 2015) - as case studies to explore the implications of prioritizing one data set over the other.  The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.

-----

# Results and Discussion

## How and why the datasets differ

The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (IUCN 2015) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.  In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.  The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of "probability of occurrence" for each species (Kaschner et al. 2006, Ready et al. 2010).  Studies frequently define "presence" for AquaMaps by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (Halpern et al. 2012, 2015).

The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).  Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.  Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (Rondinini et al. 2006).  By comparing maps resulting from IUCN and AquaMaps methodolog, we can identify and possibly address mechanistic causes for each type of error.

The two datasets have notably different taxonomic (Fig. 1) and regional (Fig. 2) coverage.  AquaMaps encompasses a broader range of taxa than IUCN, as IUCN spatial data files are only available for select taxonomic groups that have been comprehensively assessed. While species numbers in both datasets peak in tropical latitudes near the equator, species counts for IUCN maps drop quickly beyond 30°N and 30°S, while species counts for AquaMaps remain robust well into temperate latitudes.  The longitude frequency plots show a slight bias in the IUCN dataset away from the Atlantic and eastern Pacific compared to AquaMaps.  To achieve more comprehensive global coverage of species ranges these two datasets can be used together, but the underlying methodological differences complicate such direct comparisons.

``` {r load_spp_list_by_quad, echo = FALSE}
spp_list_quads <- read.csv('data/spp_list_quads.csv', stringsAsFactors = FALSE)
n_q1 <- nrow(spp_list_quads %>% filter(quad == 'q1'))
n_q2 <- nrow(spp_list_quads %>% filter(quad == 'q2'))
# n_q2_am_sm  <- nrow(spp_list_quads %>% filter(sm_range == 'AM' & quad == 'q2'))
n_q2_corals <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral') & quad == 'q2'))
n_corals    <- nrow(spp_list_quads %>% filter(str_detect(tolower(spp_group), 'coral')))
n_q3 <- nrow(spp_list_quads %>% filter(quad == 'q3'))
n_q4 <- nrow(spp_list_quads %>% filter(quad == 'q4'))

# n_q4_datapoor <- nrow(spp_list_quads %>% filter(quad == 'q4' & occurcells < 10))
# p_q4_datapoor <- round(n_q4_datapoor/n_q4 * 100, 1)
# n_q123_datapoor <- nrow(spp_list_quads %>% filter(quad != 'q4' & occurcells < 10))
# p_q123_datapoor <- round(n_q123_datapoor/n_spp_both * 100, 1)

# n_q4_rev <- nrow(spp_list_quads %>% filter(quad == 'q4' & !is.na(reviewed)))
# p_q4_rev <- round(n_q4_rev/n_q4 * 100, 1)
# n_q123_rev <- nrow(spp_list_quads %>% filter(quad != 'q4' &  !is.na(reviewed)))
# p_q123_rev <- round(n_q123_rev/n_spp_both * 100, 1)

# quad_count_means <- read_csv(file.path(dir_data, 'quads_data_poor.csv')) %>%
#   arrange(quad)
# dp_med_am_q4    <- quad_count_means$med_occurcells[4]
# dp_med_am_q123  <- quad_count_means$med_occurcells_total[1]
# dp_med_occ_q4   <- quad_count_means$med_occ[4]
# dp_med_occ_q123 <- quad_count_means$med_occ_total[1]

spp_quads_occ <- read_csv(file.path(dir_data, 'spp_quads_occ.csv'))
med_occ_q4   <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad == 'q4']))
med_occ_q123 <- round(median(spp_quads_occ$mean_occ[spp_quads_occ$quad %in% c('q1', 'q2', 'q3')]))

```

To explore differences in species distribution and range between the two datasets, we plotted the distribution alignment (how much of the smaller range falls within the larger range, i.e., where on the map) against the area alignment (ratio of smaller range area to larger range area, i.e., how much of the map) for each shared species (Fig. 3A). 
 This analysis revealed a general negative linear pattern, suggesting that increasing similarity in range area correlates with decreasing distribution alignment.
 AquaMaps tends to extrapolate species ranges into suitable areas beyond known occurrences, such that each additional unit of range predicted by AquaMaps will fall in different locations than an additional unit of range predicted using IUCN methodology. For species with dissimilar range areas, predicted distribution for the smaller range can more easily fall within the generous bounds of the larger range. For species with increasingly similar range areas, differences in methodology become more difficult to "hide," and the distribution alignment generally becomes poorer.

The mean distribution alignment for species included in both datasets was `r dist_align_mean`%; the mean area alignment was `r area_align_mean`%. By dividing the map-paired species into quadrants based on these means, we highlight categories of relationships that help further explain this general pattern. For representative maps from each category, please refer to the supplemental materials (SI Fig. XXX).

The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in both spatial distribution and extent. These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig. 3B). This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.  We expected to see significant concurrence between maps for IUCN Red List species of concern, assuming that increased attention would lead to improved understanding of species ranges; however, we found no compelling patterns to indicate that this was the case (SI Fig. XXX).

The area-mismatched ranges contained in the upper left quadrant (n = `r n_q2`) include many species whose spatial distribution is similar, but where the IUCN range is notably larger, often extending into deeper water.  For example, corals dominate this quadrant (n = `r n_q2_corals`; `r round(n_q2_corals/n_q2 * 100, 1)`% of all species in this quadrant), and IUCN range maps tend to extend corals into waters beyond their preferred depths, likely introducing errors of commission  (SI Figs. XXX, XXX). 
 Ocean depth is explicitly included in AquaMaps models, while depth is frequently overlooked as a factor in IUCN range maps.  Simply clipping IUCN range maps to known depth preferences would resolve many of these mismatches.

Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.  Consequently, area ratios are close to 100%, though similar areas are meaningless when the distributions are poorly aligned.

The most vexing cases are in the lower left quadrant (n = `r n_q4`), where neither distribution nor area match well. Data-poor species are more common in this quadrant; indeed, the median number of species occurrence records (averaging OBIS and GBIF occurrences) for this quadrant is `r med_occ_q4` records, compared to a median of `r med_occ_q123` records for species across the other three quadrants.  When extrapolating from limited observations, the AquaMaps model often predicts species presence well beyond known occurrences, introducing commission errors; at the same time, IUCN range maps generally target known occurrences, possibly introducing omission errors for data-limited species.

-----

# Implications

Method-driven differences in commission and omission errors drive clear and significant differences in species range descriptions between AquaMaps and IUCN datasets.   To examine the implications of these differences, we replicated two recent studies, varying only the prioritization of one data set over the other.

**Case Study: The Ocean Health Index**

The global Ocean Health Index (OHI) (Halpern et al. 2012, 2015), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones. To maximize the number of represented species, OHI gleans spatial distributions from both IUCN and AquaMaps datasets, prioritizing IUCN data for the `r format(n_spp_both, big.mark = ',')` species included in both sources.  OHI currently uses a probability of occurrence threshold of 40% to determine species presence for AquaMaps data.

We recalculated the OHI species status score under several scenarios to observe the impact of toggling the prioritized data set from IUCN to AquaMaps, and toggling the AquaMaps presence threshold from 40% to 0% (<a href = '#fig4'>Fig. 4</a>).  Reducing the threshold increases the apparent range of a species; the slight decrease in average score for scenario 1 suggests increased spatial representation of threatened species.  The slight increase in mean score for scenario 2 may indicate a small increase in spatial representation of low-risk species, a small decrease in spatial representation of high-risk species, or a combination of both.  The large decrease shown in scenario 3 indicates that a zero threshold greatly increases the spatial representation of high-risk species.

Given that only `r format(p_both_all, big.mark = ',')`% of species overlap, it is surprising that changing the priority for overlapping species from IUCN maps to AquaMaps would result in such large country-level score shifts as seen in scenarios 2 and 3.  While many country scores showed only small variation across all three scenarios, select countries gained up to 5 points while others dropped as many as 12 points.  Significantly, the global score decreased by two full points in scenario 3, indicating that an arbitrary change in how the two datasets are combined can result in a more dire assessment of species conservation status.

-----

**Case Study: MPA Gap Analysis**

Klein et al. (2015) compare the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect key species and identify which species fall through gaps in protection.  The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.  They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all ("gap" species).

We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using 2015 AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig5'>Fig. 5</a>).  We found a five-fold increase in the proportion of gap species (6.4% of species vs. 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs. 47.7%).  However, this comparison also indicates a larger proportion of well-protected species with greater than 10% of range protected (2.9% of species vs. 1.5%).  Conclusions drawn from each of these datasets would paint dramatically different pictures of the protection afforded by our current global MPA network.

-----

# Conclusions

AquaMaps and IUCN range maps show reasonable agreement for many well-studied species, but substantial differences illustrate uncertainty in our understanding of spatial distribution for many others.  Although many other appriaches exist for species distribution modeling, these two are the only ones applied globally to marine species.  Identifying and addressing differences in these datasets will increase their utility for research and conservation actions.  Several likely drivers of commission errors and omission errors between these datasets point to a few important ways to improve range alignment.

For IUCN range data, clipping ranges to known depth limits improves output for many species, most notably corals and other demersal species dependent on photosynthesis.  If species' depth limits are not known, simple rules of thumb will likely reduce commission errors without introducing substantial omission errors.  For example, for most corals, researchers could clip range maps to the photosynthetic limit of 200 meters.  For AquaMaps range data, dependent primarily on environmental and physical preferences and conditions, implementing area restrictions based on biogeographical criteria such as Marine Ecoregions of the World (Spalding et al. 2007) would likely decrease commission errors and improve predictive power, especially for data-poor species.

Many studies using AquaMaps data test the sensitivity of results by varying the probability of occurrence threshold to determine presence.  This decision ultimately represents a tradeoff between errors of commission (low threshold) and omission (high threshold).  Using AquaMaps in conjunction with IUCN can mitigate potential errors, while also increasing the taxonomic and spatial breadth of coverage, as long as the differences between the datasets can be reasonably minimized. In this case, we recommend a presence threshold of 0% as it most closely approximates the "limits of distribution" criterion defined by IUCN data providers.

<!-- summarize back to big picture: CBD, MPAs, climate velocity, etc... understanding species distributions, and therefore minimizing errors, is important to all these things -->
_AquaMaps and IUCN are two of the most important databases providing species distribution information across a wide range of marine species, taxa, and regions.  The different methodologies behind the creation of each database is likely to introduce different rates of commission and omission errors, which can have a profound impact on our understanding of species ranges.  By comparing overlapping species maps from these two databases, we have identified likely sources of error, allowing data users to address or account for them.  For users of AquaMaps, IUCN, and other datasets of species distribution, reducing the likelihood of these errors is critical to effective biodiversity research and conservation action._

Human impacts on oceans (pollution, ocean acidification, overfishing, extractive use) are increasingly pressuring the marine ecosystems that provide food, economic opportunity, cultural value, and life support.  Climate change and direct human impacts are driving shifts in species richness and abundance.  Managing and protecting the oceans to deliver these benefits sustainably requires much deeper understanding of the flora and fauna.  Managing and protecting marine species improves the resilience of critical ocean ecosystems to human-induced pressures. But effective management depends critically on accurate and robust understanding of the distribution of species we want to protect.