### 2016-03-27 data_explore/data_explore_coralmaps.Rmd
writing the raster and just gonna make the map in QGIS

---
title: 'Data Exploration coral maps: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(maps)
library(readr)
library(data.table)
library(raster)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'data_explore') setwd('data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

```{r read_iucn_data, echo = FALSE, warning = FALSE, message = FALSE, results = 'hide'}

spp_corals <- 'Oculina varicosa'

# Data files
#iucn spp cells WITHOUT extinct locations of IUCN subpopulations
iucn_coral_cells <- read_csv(file.path(dir_anx, 'explore_am_v_iucn/iucn_spp_cells_2015.csv'), 
                           col_types = 'cddddc', progress = TRUE) %>%
  as.data.frame() %>%
  filter(presence != 5) %>% ### removes extinct subpopulations
  filter(sciname %in% spp_corals)

### LOICZID cell ID numbers - used in plotting rasters
loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'

### this file created in data_prep
spp_list_corals <- read.csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'), stringsAsFactors = FALSE) %>%
  filter(sciname %in% spp_corals)

```

``` {r plot_coral_spp_at_depth, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
### define representative species from each quadrant, for over-plotting

  spp <- spp_list_corals

  map_iucn <- get_spp_map(spp$iucn_sid, iucn_spp_cells) %>%
    mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0)) %>%
    select(-area_iucn, -iucn_sid) %>%
    unique()

  r_iucn_spp <- subs(loiczid_raster, 
                     map_iucn[ , c('loiczid', 'iucn_pres')], 
                     by = 'loiczid', 
                     which = 'iucn_pres', 
                     subsWithNA = TRUE)
  writeRaster(r_iucn_spp, 'coralmaps/oculina_varicosa.tif')

#   spp_pts <- as.data.frame(rasterToPoints(r_iucn_spp)) %>%
#     mutate(presence = 'IUCN')
#   
#   rast_fine <- loiczid_raster %>%
#     crop(extent(x = c(-135, -45), y = c(0, 45)))
#   res(rast_fine) <- 0.05

#   depth_poly <- rgdal::readOGR(dsn = 'coralmaps', layer = 'bath_200m')
#   depth_pts <- rasterize(depth_poly, rast_fine) %>%
#     rasterToPoints() %>%
#     as.data.frame() %>%
#     rename(bath200m = layer)
#   
#   spp_plot <- ggplot(spp_pts, aes(x = x, y = y)) +
#     ggtheme_map + 
#     theme(panel.background = element_rect(fill = 'lightblue')) +
# #    geom_raster(aes(fill = presence), alpha = 1) +
#     geom_raster(data = depth_pts, aes(x = x, y = y, fill = bath200m),
#                 alpha = .5) +
# #    scale_fill_manual(values = c('IUCN' = '#d95f02')) +
#     borders('world', color='gray40', fill='gray45', size = .1) +
#     coord_cartesian(xlim = c(-105, -60), ylim = c(7.5, 37.5)) +
#     scale_x_continuous(breaks = seq(-105, -60, by = 15)) +
#     scale_y_continuous(breaks = seq( 0,  45, by = 15)) +
#     theme(legend.position = 'none',
#           axis.text = element_blank())
#   
#   print(spp_plot)
#   plotfile <- file.path(dir_fig, 'spp_maps', 
#                         sprintf('%s_%s.png', 
#                                 spp_quad, 
#                                 tolower(str_replace(spp_quads[i], ' ', '_'))))
#   ggsave(plotfile, height = 4.5, width = 8.7, units = 'cm')
  }

```

``` {r examine_spp_gp_quartiles, echo = FALSE, eval = FALSE}

spp_list1 <- spp_list %>%
  group_by(spp_group_text) %>%
  mutate(n_spp = n()) %>%
  ungroup()

### Create species lists by quadrant
spp_list_q1 <- spp_list1 %>%
  filter(area_ratio >= median(area_ratio) & sm_perc >= median(sm_perc))
spp_list_q2 <- spp_list1 %>%
  filter(area_ratio < median(area_ratio) & sm_perc >= median(sm_perc))
spp_list_q3 <- spp_list1 %>%
  filter(area_ratio >= median(area_ratio) & sm_perc < median(sm_perc))
spp_list_q4 <- spp_list1 %>%
  filter(area_ratio < median(area_ratio) & sm_perc < median(sm_perc))

spp_list_quads <- spp_list1 %>%
  mutate(quad = NA,
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
write.csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'), row.names = FALSE)

### Summarize quadrant species list by species group (e.g. CORALS1)
spp_gp_q1 <- spp_list_q1 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q1 = n())
spp_gp_q2 <- spp_list_q2 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q2 = n())
spp_gp_q3 <- spp_list_q3 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q3 = n())
spp_gp_q4 <- spp_list_q4 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q4 = n())

### Join species group by quadrant into a single dataframe
spp_gp_quadrants <- spp_gp_q1 %>%
  full_join(spp_gp_q2, by = c('spp_group_text', 'n_spp')) %>%
  full_join(spp_gp_q3, by = c('spp_group_text', 'n_spp')) %>%
  full_join(spp_gp_q4, by = c('spp_group_text', 'n_spp')) %>%
  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
  mutate(quad = str_replace(quad, 'n_spp_', ''),
         pct_quad = n_quad/n_spp,
         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))

spp_gp_quadrants <- spp_gp_quadrants %>%
  left_join(spp_gp_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(spp_group_text, pct_q1 = pct_quad),
            by = 'spp_group_text') %>%
  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
         quad = str_replace(quad, 'q3', 'q3 area-matched'),
         quad = str_replace(quad, 'q4', 'q4 poorly matched'))

```

<!-- # Multiple species maps per quadrant -->
``` {r child = 'data_explore_pt3.Rmd', eval = FALSE}
```