### 2016-05-13 testbib.Rmd

---
title: "Aligning marine species range data to better serve science and conservation"
author: Casey O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin
  S. Halpern^1,2,3^
output:
  html_document:
    highlight: haddock
    number_sections: no
    theme: spacelab
    toc: no
  pdf_document:
    number_sections: no
    toc: no
  word_document: default
bibliography: refs/zotero_lib.bib
csl: refs/ecology-letters.csl
---

All processing was completed using R statistical software [@r_core_team_r:_2016], and all code and intermediate data are available on GitHub at https://github.com/OHI-Science/IUCN-AquaMaps.

evaluating global biodiversity patterns [@kaschner_current_2011].

[@halpern_index_2012]

[@jones_modelling_2012]

[@selig_assessing_2013]

# REFERENCES