### 2016-05-18 case_study_mpa/3_mpa_spp_analysis.Rmd
updated to Mazu to rerun the figures.

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_M             <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
                       'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
                       'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')
dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv')

# if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

# MPA gap analysis

For each species:

* for each "present" LOICZID cell
    * determine total area: overall cell area (in km2) * marine area (cells/2500)
    * determine protected area: overall cell area (in km2) * protected area (cells/2500)
* sum total area and protected area across all "present" LOICZID cells
* report: 
    * ratio of (protected area:total area) (dimensionless, percentage)
    * total marine area and total protected area (km2)
    * species name, ID, taxonomic group, extinction risk

``` {r combine_mpa_areas_all}
yr_list <- c('y1990', 'y1995', 'y2000', 'y2005', 'y2010', 'y2014', 'y2015')

combine_mpa_areas <- function(mpa_file_base, yr_list) {

  hcaf_fn <- file.path(dir_am_data, 'hcaf_truncated.csv')
  message('Reading hcaf cells to get area per cell: ', hcaf_fn)
  ### get AquaMaps file of cells to translate cell IDs to area and ocean area
  hcaf <- read_csv(hcaf_fn) %>%
    select(loiczid, cellarea, oceanarea)

  ### attach the AquaMaps areas to the calculated percentages of marine area
  ### note: AquaMaps calculated areas differ from my calculated areas - see below
  cell_df <- read_csv(file.path(dir_git, 'case_study_mpa/celldata/loiczid_marine.csv'),
    col_types = 'dd') %>%
    rename(marine_ct = count) %>%
    left_join(hcaf, by = 'loiczid') %>%
    mutate(marine_pct  = marine_ct/2500,
           marine_area = marine_pct * cellarea) %>%
    select(-marine_ct)

  ### load each analysis year (in yr_list) and join them all together
  ### to note protection status year of each cell.

  for (yr in yr_list) { # yr <- yr_list[1]
    fn <- file.path(dir_git, sprintf(mpa_file_base, yr))
    message('Reading in mpa cell file: ', fn)
    df <- read_csv(fn, col_types = 'dd')
    names(df)[2] <- yr
    cell_df <- cell_df %>%
      left_join(df, by = 'loiczid')
    cell_df[ , paste(yr, '_pct', sep = '')]  <- cell_df[ , yr]/2500
    cell_df[ , paste(yr, '_area', sep = '')] <- cell_df[ , paste(yr, '_pct', sep = '')] * cell_df$cellarea

    ### remove intermediate columns
    cell_df[ , yr] <- NULL
    cell_df[ , paste(yr, '_pct', sep = '')] <- NULL
  }
  return(cell_df)
}

cell_df_klein <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_klein_i-iv_%s.csv', yr_list)
cell_df_all   <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_i-iv_%s.csv',       yr_list)
cell_df_des   <- combine_mpa_areas('case_study_mpa/celldata/loiczid_mpa_i-iv-des_%s.csv',   yr_list)

# ### check diffs in oceanarea vs marine_area
# x <- cell_df %>% 
#   filter(oceanarea != marine_area) %>%
#   select(loiczid, oceanarea, marine_area, cellarea)
# ### 18158 cells out of 259200 = 7%
# x <- x %>% mutate(pct_diff = (marine_area - oceanarea)/cellarea)
# hist(x$pct_diff)
# ### most differences cluster right around zero.  Not going to worry about it!

```

# recreate Klein results using aquamaps 2014 data

``` {r recreate Klein results 2014 AM data, echo = FALSE, eval = TRUE}

area_mpa <- cell_df_klein %>%
  summarize(marine_area  = sum(marine_area),
            mpa_tot_2014 = sum(y2014_area))

### Recreate Klein results, using 2014 AquaMaps
am_spp_cells2014 <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc_2014.csv'))
mpa_klein_am2014 <- am_spp_cells2014 %>%
  left_join(cell_df_klein, by = 'loiczid')

### filter to the appropriate threshold...
thresh <- 0.50

mpa_klein_am2014 <- mpa_klein_am2014 %>%
  filter(probability >= thresh)

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
mpa_klein_am2014 <- mpa_klein_am2014 %>%
  group_by(speciesid) %>%
  summarize(marine_range = sum(marine_area),
            mpa2014 = sum(y2014_area),
            pct2014 = mpa2014/marine_range)

mpa_klein_am2014 <- mpa_klein_am2014 %>%
  mutate(cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014))

mpa_klein_orig <- mpa_klein_am2014 %>%
  group_by(cat2014) %>%
  summarize(n_tot   = nrow(mpa_klein_am2014), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2014 = factor(cat2014, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2014)

knitr::kable(mpa_klein_orig)
# |cat2014 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 17344|   282|     1.6|
# |0-2%    | 17344|  7728|    44.6|
# |2-5%    | 17344|  7738|    44.6|
# |5-10%   | 17344|  1188|     6.8|
# |>10%    | 17344|   408|     2.4|

```

| study            | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :--------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015 | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, wdpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |

# recreate Klein analysis using 2015 aquamaps

``` {r load AM 2015 data, echo = FALSE}
am_spp_cells <- read_csv(file.path(dir_am_data, 'hcaf_sp_native_trunc.csv')) ### AM 2015

```

``` {r recreate Klein results 2015 AM data, echo = FALSE, eval = TRUE}

area_mpa <- cell_df_klein %>%
  summarize(marine_area  = sum(marine_area),
            mpa_tot_2015 = sum(y2015_area))
### Note: should be same as using 2014 data, since cell_df_klein is based
### on 2014 WDPA database

### Recreate Klein results, using 2015 AquaMaps
mpa_klein_am2015 <- am_spp_cells
mpa_klein_am2015 <- mpa_klein_am2015 %>%
  left_join(cell_df_klein, by = 'loiczid')

### filter to the appropriate threshold...
thresh <- 0.50

mpa_klein_am2015 <- mpa_klein_am2015 %>%
  filter(probability >= thresh)

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
mpa_klein_am2015 <- mpa_klein_am2015 %>%
  group_by(speciesid) %>%
  summarize(marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

mpa_klein_am2015 <- mpa_klein_am2015 %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

write_csv(mpa_klein_am2015, file.path(dir_git, 'case_study_mpa/celldata/mpa_by_spp_am2015.csv'))

mpa_klein_update <- mpa_klein_am2015 %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(mpa_klein_am2015), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(mpa_klein_update)
# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 22885|   379|     1.7|
# |0-2%    | 22885| 10454|    45.7|
# |2-5%    | 22885|  9777|    42.7|
# |5-10%   | 22885|  1654|     7.2|
# |>10%    | 22885|   621|     2.7|

```

| study            | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :--------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015 | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, wdpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| am2015, wdpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.7% | 45.7% | 42.7% |  7.2% | 2.7% |

# Switch to WDPA 2015 data set?

No.  Save that for time-series analysis.

``` {r iucn_only}

### Load IUCN cells data - 
### load spp_all from SPP_ICO v2016;
###

### compare results using only IUCN species maps
dir_iucn_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/iucn_spp')
iucn_cells_file <- file.path(dir_iucn_data, 'iucn_cells_v2016.csv')
iucn_spp_cells <- read_csv(iucn_cells_file, col_types = 'cddddc') %>%
  as.data.frame(stringsAsFactors = FALSE)

iucn_spp_cells <- iucn_spp_cells %>%
  left_join(cell_df_klein, by = 'loiczid')

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
iucn_spp_sum <- iucn_spp_cells %>%
  rename(iucn_sid = id_no) %>%
  group_by(iucn_sid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

iucn_spp_sum <- iucn_spp_sum %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015)) %>%
  mutate(speciesid = paste('iucn', iucn_sid, sep = ''))

iucn_spp_sum2 <- iucn_spp_sum %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(iucn_spp_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(iucn_spp_sum2)

# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     |  4024|   257|     6.4|
# |0-2%    |  4024|  2687|    66.8|
# |2-5%    |  4024|   810|    20.1|
# |5-10%   |  4024|   152|     3.8|
# |>10%    |  4024|   118|     2.9|

```

| study              | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :----------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015   | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, wdpa2014   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| am2015, wdpa2014   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.7% | 45.7% | 42.7% |  7.2% | 2.7% |
| iucn2015, wdpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 6.4% | 66.8% | 20.1% |  3.8% | 2.9% |

``` {r 2015 AM data w 0 threshold, echo = FALSE, eval = TRUE}

### Set up AquaMaps species with zero threshold
mpa_am2015_0 <- am_spp_cells
mpa_am2015_0 <- mpa_am2015_0 %>%
  left_join(cell_df_klein, by = 'loiczid')

### skip filtering step - threshold = 0

### group by speciesid, then summarize sum(marine area), sum(protected area), ratio
mpa_am2015_0 <- mpa_am2015_0 %>%
  group_by(speciesid) %>%
  summarize(marine_range = sum(marine_area),
            mpa2015 = sum(y2015_area),
            pct2015 = mpa2015/marine_range)

mpa_am2015_0 <- mpa_am2015_0 %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

mpa_am2015_0_sum <- mpa_am2015_0 %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(mpa_am2015_0), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(mpa_am2015_0_sum)
# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 22889|   268|     1.2|
# |0-2%    | 22889| 10643|    46.5|
# |2-5%    | 22889| 10187|    44.5|
# |5-10%   | 22889|  1455|     6.4|
# |>10%    | 22889|   336|     1.5|

```

``` {r combine IUCN and AM with 0% thresh AM on top}

# head(iucn_spp_sum)
# head(mpa_am2015_0)

### get list of overlap species and remove from IUCN species (they are
### included in the AM species list)
spp_overlap <- read_csv(file.path(dir_data, 'spp_am_v_iucn.csv'))

mpa_am2015_iucn <- mpa_am2015_0 %>%
  bind_rows(iucn_spp_sum %>%
              filter(!iucn_sid %in% spp_overlap$iucn_sid) %>%
              select(-iucn_sid, cellarea_tot, ocean_range))

mpa_am2015_iucn_sum <- mpa_am2015_iucn %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(mpa_am2015_iucn), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(mpa_am2015_iucn_sum)

# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 24634|   494|     2.0|
# |0-2%    | 24634| 11823|    48.0|
# |2-5%    | 24634| 10401|    42.2|
# |5-10%   | 24634|  1521|     6.2|
# |>10%    | 24634|   395|     1.6|

```

| study              | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :----------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015   | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014 50%   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| am2015 50%   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.7% | 45.7% | 42.7% |  7.2% | 2.7% |
| am2015 0%   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.2% | 46.5% | 44.5% |  6.4% | 1.5% |
| iucn2015 | IUCN I-IV, Sp. Mar |   2,915,745  | 6.4% | 66.8% | 20.1% |  3.8% | 2.9% |
| am2015 0% + iucn2015 | IUCN I-IV, Sp. Mar |   2,915,745  | 2.0% | 48.0% | 42.2% |  6.2% | 1.6% |

``` {r combine all runs into one df}
mpa_all_runs <- mpa_klein_am2014 %>%
  mutate(trial = 'AM 2014 (50%)') %>%
  rename(cat = cat2014, pct = pct2014) %>%
  bind_rows(mpa_klein_am2015 %>%
    mutate(trial = 'AM 2015 (50%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_0 %>%
    mutate(trial = 'AM 2015 (0%)') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(mpa_am2015_iucn %>%
    mutate(trial = 'AM 2015 (0%) +\nIUCN 2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  bind_rows(iucn_spp_sum %>%
    mutate(trial = 'IUCN 2015') %>%
    rename(cat = cat2015, pct = pct2015)) %>%
  select(speciesid, cat, pct, trial) %>%
  unique()

# names(mpa_all_runs)[3:7] <- paste('mpa', names(mpa_all_runs)[3:7], sep = '')

# mpa_all_runs <- mpa_all_runs %>%
#   select(trial, n_tot, 3:7) %>%
#   gather(mpa_cat, pct, 3:7) %>%
#   mutate(mpa_cat = factor(mpa_cat, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
#   arrange(mpa_cat)

write_csv(mpa_all_runs, file.path(dir_data, 'mpa_all_runs.csv'))

# mpa_runs <- read_csv(file.path(dir_data, 'mpa_all_runs.csv'))
```

``` {r plot this beast}
library(ggplot2)
source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

if(!exists('mpa_all_runs'))
  mpa_all_runs <- read_csv(file.path(dir_data, 'mpa_all_runs.csv'))

mpa_all_runs1 <- mpa_all_runs %>%
  group_by(trial) %>%
  mutate(n_trial = n()) %>%
  group_by(trial, cat, n_trial) %>%
  summarize(p_cat = n()/first(n_trial)) %>%
  ungroup() %>%
  mutate(cat = factor(cat, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  mutate(trial = factor(trial, levels = c('IUCN 2015', 'AM 2015 (0%) +\nIUCN 2015', 'AM 2015 (0%)', 'AM 2015 (50%)', 'AM 2014 (50%)'))) %>%
  mutate(p_cat = p_cat * 100) %>%
  arrange(cat)

### Plot the bar chart
barchart_mpa_gap <- ggplot(mpa_all_runs1, 
                           aes(x = trial, fill = cat, weight = p_cat)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = .8) +
  scale_fill_manual(values = c('>10%'  = '#4dac26', 
                               '5-10%' = '#b8e186', 
                               '2-5%'  = '#fbe5c5', 
                               '0-2%'  = '#f1b6da',
                               'gap'   = '#d01c8b')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 125), breaks = c(seq(0, 100, 20))) +
  geom_text(aes(label = sprintf('n = %s', n_trial), y = 101), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  scale_x_discrete(labels = c('5. IUCN 2015',
                              '4. AM 2015 (0%) +\nIUCN 2015', 
                              '3. AM 2015 (0%)', 
                              '2. AM 2015 (50%)', 
                              '1. AM 2014 (50%)')) +
  coord_flip() +
  labs(x = 'Scenario', 
       y = 'Percent of species by MPA coverage', 
       fill = 'MPA coverage')

print(barchart_mpa_gap)
ggsave(plot = barchart_mpa_gap,
       filename = file.path(dir_fig, 'fig4_barchart_mpa_gap.png'),
       width = 11.4, height = 5, units = 'cm', dpi = 600)
ggsave(plot = barchart_mpa_gap,
       filename = file.path(dir_fig, 'fig4_barchart_mpa_gap.tif'), device = 'tiff',
       width = 11.4, height = 5, units = 'cm', dpi = 600)

mpa_all_runs_wide <- mpa_all_runs1 %>%
  mutate(p_cat = paste(round(p_cat, 1), '%', sep = '')) %>%
  spread(cat, p_cat) %>%
  rename(Scenario = trial, n = n_trial)
names(mpa_all_runs_wide)[4:7] <- paste('covered', names(mpa_all_runs_wide)[4:7], sep = ' ')
mpa_all_runs_wide <- mpa_all_runs_wide[rev(mpa_all_runs_wide$Scenario), ]

knitr::kable(mpa_all_runs_wide)

write_csv(mpa_all_runs_wide, file.path(dir_data, 'mpa_all_runs_summary.csv'))
```

|Scenario               |     n|gap  |covered 0-2% |covered 2-5% |covered 5-10% |covered >10% |
|:----------------------|-----:|----:|------------:|------------:|-------------:|------------:|
|am2014 (50%)           | 17344|1.6% |44.6%        |44.6%        |6.8%          |2.4%         |
|am2015 (50%)           | 22885|1.7% |45.7%        |42.7%        |7.2%          |2.7%         |
|am2015 (0%)            | 22889|1.2% |46.5%        |44.5%        |6.4%          |1.5%         |
|am2015 (0%) & iucn2015 | 24634|2%   |48%          |42.2%        |6.2%          |1.6%         |
|iucn2015               |  4024|6.4% |66.8%        |20.1%        |3.8%          |2.9%         |