### 2016-08-07 coralmaps/coral_maps.Rmd
cleaning up the old drafts; cleaning up the data_explores

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(maps)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'data_explore') setwd('data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

```{r get_bathy_raster, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

### get shapefile
### rasterize to half-degree grid; or raster::extract

if(!file.exists(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))) {

  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'coralmaps/ne_10m_bathymetry_K_200'), 
                       layer = 'ne_10m_bathymetry_K_200', 
                       stringsAsFactors = TRUE)

  ### LOICZID cell ID numbers - used in plotting rasters
  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

  bath_list <- raster::extract(loiczid_raster, poly_bath, 
                             weights = TRUE, normalizeWeights = FALSE,
                             progress = 'text')
  bath_df <- lapply(bath_list, function(x) {as.data.frame(x)}) %>%
    bind_rows() %>%
    rename(loiczid = value) %>%
    group_by(loiczid) %>%
    summarize(weight = sum(weight))

  ### generate raster?
  bath_raster <- subs(loiczid_raster, bath_df, by = 'loiczid', which = 'weight')
  values(bath_raster)[values(bath_raster) > 5] <- NA

  write_csv(bath_df, file.path(dir_git, 'coralmaps', 'bathy_cells.csv'))
  ### probably want to filter to just 100% 200m+
}

```

``` {r}
### find spp depth limits
### focus just on corals?

if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_depth_raw.csv'))) {

  library(jsonlite)
  api_key <- 'fb71ae836f415f04f41176f6d30c4a9e4cea620d46b9e5021bf2fb142ea51bf5'

  get_from_api <- function(url, param) {
      api_info <- fromJSON(sprintf(url, param, api_key)) %>%
        data.frame(stringsAsFactors = FALSE)
  }

  mc_get_from_api <- function(url, param_vec) {
    numcores = ifelse(Sys.info()[['nodename']] == 'mazu', 12, 1)
    out_df <- parallel::mclapply(param_vec, 
                            function(x) get_from_api(url, x),
                            mc.cores = numcores) %>% 
      bind_rows() 
    out_df <- out_df %>%
      setNames(names(.) %>%
                 str_replace('result.', ''))
  }

  ### get paired species list
  paired_maps_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
  spp_map_pairs <- read_csv(paired_maps_file) %>%
    filter(!is.na(popn_category)) %>% ### filter out unassessed species
    filter(!is.na(sm_perc))           ### filter out NA maps

  ### get full species list
  spp_all_file <- file.path(dir_data, 'spp_all.csv')
  spp_all <- read_csv(spp_all_file)

  coral_spp_sid <- spp_all %>%
    filter(str_detect(tolower(spp_group), 'coral')) %>%
    .$iucn_sid %>%
    unique()

  spp_narr_url <- 'http://apiv3.iucnredlist.org/api/v3/species/narrative/id/%s?token=%s'
  spp_narr     <- mc_get_from_api(spp_narr_url, coral_spp_sid)

  write_csv(spp_narr, file.path(dir_git, 'coralmaps', 'coral_spp_narratives.csv'))

  spp_depth <- spp_narr %>%
    select(iucn_sid = species_id, habitat) %>%
    mutate(paired_map = (iucn_sid %in% spp_map_pairs$iucn_sid))

spp_depth1 <- spp_depth %>%
  mutate(depth = str_extract_all(habitat, '\\d++(?:\\.\\d++)?(?=\\s*+(?:m))')) %>%
  unnest(depth) %>%
  group_by(iucn_sid) %>%
  summarize(max_depth = max(as.integer(depth)))

spp_depth <- spp_depth %>%
  left_join(spp_depth1, by = 'iucn_sid') %>%
  mutate(shallow_str = str_detect(tolower(habitat), 'shallow'),
         shallow = max_depth <= 200 | shallow_str)

write_csv(spp_depth, file.path(dir_git, 'coralmaps', 'coral_spp_depth.csv'))

summary(spp_depth)
 #    iucn_sid        habitat          paired_map        max_depth      shallow_str     shallow       
 # Min.   :132859   Length:562         Mode :logical   Min.   :  1.00   Mode :logical   Mode:logical  
 # 1st Qu.:133081   Class :character   FALSE:174       1st Qu.: 20.00   FALSE:194       TRUE:544      
 # Median :133284   Mode  :character   TRUE :388       Median : 20.00   TRUE :367       NA's:18       
 # Mean   :133283                      NA's :0         Mean   : 27.55   NA's :1                       
 # 3rd Qu.:133486                                      3rd Qu.: 30.00                                 
 # Max.   :133703                                      Max.   :200.00                                 
 #                                                     NA's   :46

nrow(spp_depth %>% filter(max_depth > 100))
# [1] 5
### of 562 species, only 5 are listed as occurring below 100 m; none are below 200 m

spp_depth_na <- spp_depth %>%
  filter(is.na(shallow))

### 17 species (of 396) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
```

``` {r extract_IUCN_cells_for_coral_pairs}

bath_df <-  read_csv(file.path(dir_git, 'coralmaps', 'bathy_cells.csv'), col_types = 'dd') %>%
  filter(weight >= 1) %>%
  mutate(deep = 1) %>%
  select(-weight)

### load iucn spp_cells
coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
                          pattern = 'CORAL*.*',
                          full.names = TRUE)
coral_cells_iucn <- lapply(coral_files, read_csv) %>%
  bind_rows() %>%
  select(iucn_sid, loiczid = LOICZID) %>%
  filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
  left_join(bath_df, by = 'loiczid') %>%
  mutate(deep = ifelse(is.na(deep), 0, deep))

write_csv(coral_cells_iucn, file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))

```

``` {r calc_areas_for_iucn_corals}
### calc areas
### read hcaf to get areas; left join to coral_cells_iucn
### group_by iucn_sid and depth, summarize area (deep and shallow)
### group_by iucn_sid, summarize total area
coral_cells_iucn <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv'))

hcaf <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
  select(loiczid, cellarea, oceanarea)

iucn_coral_areas <- coral_cells_iucn %>%
  left_join(hcaf, by = 'loiczid') %>%
  group_by(iucn_sid, deep) %>%
  summarize(deeparea = sum(oceanarea)) %>%
  group_by(iucn_sid) %>%
  mutate(totarea = sum(deeparea),
         shallow_pct = (totarea - deeparea)/totarea) %>%
  arrange(desc(deep)) %>%
  .[!duplicated(.$iucn_sid), ]

### get species list
data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(popn_category)) %>%   ### filter out unassessed species
  filter(!is.na(sm_perc))             ### filter out NA maps

iucn_coral_areas <- iucn_coral_areas %>%
  left_join(spp_depth,     by = 'iucn_sid') %>%
  left_join(spp_map_pairs, by = 'iucn_sid')

write_csv(iucn_coral_areas, file.path(dir_git, 'coralmaps', 'iucn_coral_areas.csv'))

```

``` {r calc_overlaps}
if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))) {

  ### get species list
  coral_cells_iucn_clipped <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0) %>%
    mutate(prop_area = 1)

  spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
    filter(iucn_sid %in% coral_cells_iucn_clipped$iucn_sid) %>%
    select(am_sid, sciname, iucn_sid)

  ### copy aquamaps cells to local file, keeping only corals in the paired maps
  if(!file.exists(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))) {
    coral_cells_am <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
      rename(am_sid = speciesid, prob = probability) %>%
      filter(am_sid %in% spp_coralmaps$am_sid) 
    write_csv(coral_cells_am, file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))
  }

  coral_cells_am <- read_csv(file.path(dir_git, 'coralmaps', 'coral_cells_am.csv'))

  loiczid_raster_file  <- file.path(dir_git, 'shiny/data/loiczid_raster.grd')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

  spp_coralmaps <- spp_coralmaps %>%
    mutate(area_total = NA,  area_am = NA,  area_iucn = NA,
           area_overlap = NA,
           sm_perc    = NA, #filled with % of smaller range that is within the larger range
           sm_range   = NA) #what range is smaller

  i         <- 0 ### initialize the counting index
  am_thresh <- 0 ### set the aquamaps presence threshold
  sid_vector <- unique(spp_coralmaps$iucn_sid)
  message(sprintf('There are a total of %s distinct species, by IUCN SID', 
                  length(sid_vector)))

  for (sid in sid_vector) { 
    # sid <- spp_coralmaps$iucn_sid[1]
    i <- i + 1
    sids <- spp_coralmaps %>% 
      select(am_sid, iucn_sid, sciname) %>%
      filter(iucn_sid == sid)
    rowindex <- (spp_coralmaps$iucn_sid == sid)

    message(sprintf('%s. Processing species iucn_sid: %s \n   am_sid(s):  %s \n   sciname(s): %s', 
                    i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
    map_iucn <- get_spp_map(sid, coral_cells_iucn_clipped) %>%
      mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
    map_am   <- get_spp_map(sids$am_sid, coral_cells_am) %>%
      mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))

    ### rasterize aquamaps map
    if(any(duplicated(map_am$loiczid))) message('Duplicated cells in am spp map')
    map_am <- map_am %>%
      dplyr::select(loiczid, am_pres) %>% 
      unique()

    r_am_spp <- subs(x = loiczid_raster, y = map_am %>% select(loiczid, am_pres), 
                     by = 'loiczid', which = 'am_pres', 
                     subsWithNA = TRUE)

    ### rasterize IUCN map
    if(TRUE %in% duplicated(map_iucn$loiczid)) stop('Duplicated cells in IUCN map')
    r_iucn_spp <- subs(loiczid_raster, map_iucn %>% select(loiczid, iucn_pres), 
                       by = 'loiczid', which = 'iucn_pres', 
                       subsWithNA = TRUE)

    ### Start calculating areas etc.
    area_iucn <- area(r_iucn_spp, na.rm = TRUE) %>%
      cellStats(., stat = 'sum')
    area_am   <- area(r_am_spp, na.rm = TRUE) %>%
      cellStats(., stat = 'sum')
    spp_coralmaps[rowindex, ]$area_iucn <- area_iucn
    spp_coralmaps[rowindex, ]$area_am   <- area_am

    ### Get total area (IUCN and AM combined)
    all_cells <- data.frame(loiczid = unique(c(map_iucn$loiczid, map_am$loiczid)),
                           spp_pres = 1)
    r_all_cells <- subs(loiczid_raster, all_cells, 
                       by = 'loiczid', which = 'spp_pres', 
                       subsWithNA = TRUE)
    area_total  <- area(r_all_cells, na.rm = TRUE) %>% cellStats(., stat = 'sum')
    spp_coralmaps[rowindex, ]$area_total <- area_total

    ### Get total overlap (IUCN and AM together). If area_iucn = 0, just assign 0
    if(area_iucn != 0) {
      overlap_cells <- unique(map_iucn$loiczid[map_iucn$loiczid %in% map_am$loiczid])
      if(length(overlap_cells) > 0) {
        overlap_cells <- data.frame(loiczid = overlap_cells,
                                  spp_pres = 1)
        r_overlap_cells <- subs(loiczid_raster, overlap_cells, 
                                by = 'loiczid', which = 'spp_pres', 
                                subsWithNA = TRUE)
        area_overlap <- area(r_overlap_cells, na.rm = TRUE) %>% cellStats(., stat = 'sum')
      } else area_overlap <- 0 ### if no overlapping cells, assign zero
    } else area_overlap <- 0 ### if no IUCN map cells, assign zero
    spp_coralmaps[rowindex, ]$area_overlap <- area_overlap

    message(sprintf('   Species iucn_sid %s: smaller range is %s. Smaller-in-larger is %.2f%%, A_sm:A_lg is %.2f%%.',
                    sid, ifelse(area_am < area_iucn, 'AquaMaps', 'IUCN'), 
                    ifelse(area_am < area_iucn, area_overlap/area_am*100, area_overlap/area_iucn*100),
                    ifelse(area_am < area_iucn, area_am/area_iucn*100, area_iucn/area_am*100)))
  }

  spp_coralmaps <- spp_coralmaps %>%
    mutate(sm_range   = ifelse(area_am < area_iucn, 'AM', 'IUCN'),
           sm_perc    = ifelse(area_am < area_iucn, (area_overlap / area_am) * 100, (area_overlap / area_iucn) * 100),
           area_ratio = area_am / area_iucn,
           area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
           area_ratio = area_ratio * 100)

  spp_coralmaps <- spp_coralmaps %>%
    mutate(lg_area = ifelse(area_am > area_iucn, area_am, area_iucn),
           sm_area = ifelse(area_am < area_iucn, area_am, area_iucn),
           lg_area_pct = 100 * lg_area / max(area_total, na.rm = TRUE)) %>% 
             ### use largest data set as approx for total ocean range
    select(-lg_area, -sm_area)

  spp_coralmaps <- spp_coralmaps %>%
    mutate(area_total  = round(area_total, 1),
           area_am     = round(area_am, 1),
           area_iucn   = round(area_iucn, 1),
           sm_perc     = round(sm_perc, 3),
           area_ratio  = round(area_ratio, 3),
           lg_area_pct = round(lg_area_pct, 3))

  print(head(spp_coralmaps))

  ### combine old and new area dataframes
  spp_coralmaps <- spp_coralmaps %>%
    select(iucn_sid, am_sid, sciname, area_am, 
           area_iucn_clipped = area_iucn, area_overlap_clipped = area_overlap, 
           sm_perc_clipped = sm_perc, sm_range_clipped = sm_range, area_ratio_clipped = area_ratio) %>%
    left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
                select(am_sid, iucn_sid, 
                       area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
                       sm_perc_raw = sm_perc, sm_range_raw = sm_range, area_ratio_raw = area_ratio),
              by = c('iucn_sid', 'am_sid'))

  write_csv(spp_coralmaps, file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))
}
```

``` {r}

spp_coralmaps <- read_csv(file.path(dir_git, 'coralmaps', 'coral_spp_areas.csv'))

### check raw iucn_area vs clipped - looks proportional; scaling factor due to difference in area calcs?
scatterplot1 <- ggplot(spp_coralmaps, aes(x = area_iucn_raw, y = area_iucn_clipped)) +
  geom_point() +
  geom_abline(color = 'red')
scatterplot1

### rough recalcs
### assume that by removing the deep cells, we will not remove cells that are
### included in AquaMaps, since AquaMaps already filters by depth.  This means
### the overlap area will not be affected.
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, perc, c(sm_perc_raw, sm_perc_clipped)) %>%
  gather(area, ratio, c(area_ratio_raw, area_ratio_clipped)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'clip') & !(str_detect(area, 'clip')))) %>%
  mutate(method = ifelse(str_detect(dist, 'clip'), '200m clip', 'all depth'))

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
clipped_quads <- ggplot(iucn_coral_fixed %>%
                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
                                               filter(perc > 100) %>% 
                                               .$iucn_sid)),
                    aes(x = ratio, y = perc, group = iucn_sid)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  # annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100, 
  #          alpha = .3, fill= "#4dac26")  + 
  # annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100, 
  #          alpha = .3, fill= "#b8e186") + 
  # annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0, 
  #          alpha = .3, fill= "#f1b6da") + 
  # annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0, 
  #          alpha = .3, fill= "#d01c8b") +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 102),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 103),
                     breaks = c(seq(0, 100, 25)),
                     labels = c('0%', '25%', '50%', '75%', '100%')) +

  geom_line(color = 'grey50', size = .5, alpha = .3) +
  geom_point(aes(color = method), size = 2, alpha = .8) +
  labs(x = 'Area ratio (smaller/larger)',
       y = 'Distribtion (%smaller in larger)')

print(clipped_quads)

ggsave(plot = clipped_quads,
       filename = file.path(dir_fig, 'scatter_quadmap_corals.png'), device = 'png',
       width = 17.8, height = 9, units = 'cm', dpi = 600)

```