### 2016-08-07 drafts/initial_version/ms_2_intro.Rmd

---
output: pdf_document
---

# Introduction

Knowing where species exist and thrive is fundamental to the sciences of ecology, biogeography, and conservation, among many others.  This knowledge provides foundational information for 
 understanding species ranges and diversity,
 predicting species responses to human impacts and climate change,
 and managing and protecting species effectively.
A rich literature tackles the many dimensions of these questions.

One major outcome of this body of science is the various compiled databases of species distribution maps. Two global-scale repositories predict marine species ranges throughout the world's oceans – AquaMaps [@kaschner_aquamaps:_2015] and International Union for Conservation of Nature (IUCN) [@iucn_iucn_2015].  These two spatial datasets have been used in hundreds of studies and applications for a wide range of purposes, including assessing marine species status [@halpern_index_2012; @halpern_patterns_2015; @selig_assessing_2013-1], evaluating global biodiversity patterns [@coll_biodiversity_2010; @martin_manual_2014; @pimm_biodiversity_2014; @kaschner_current_2011], predicting species range shifts [@garciamolinos_climate_2015], and setting conservation priorities [@klein_shortfalls_2015].

<!--
Three foundational papers relating to AquaMaps (Kaschner et al. 2006, 2011; Ready et al. 2010) have been cited a total of 315 times.
-->

The two datasets ostensibly describe the same information, but significant differences in methodology and intent could lead to dramatically different understandings of our marine ecosystems, with significant implications for policy and conservation recommendations.  Importantly, biases in taxonomic or spatial coverage within a dataset could shift management and conservation actions away from places or species that are most in need.  Inaccurate indications of presence or absence could lead to ineffective marine reserve systems and management plans [@rondinini_tradeoffs_2006; @jetz_ecological_2008].

To understand the implications of differences between the AquaMaps and IUCN datasets, we compare how each data source represents the global spatial and taxonomic distribution of marine species.  Most notably, AquaMaps includes range maps for many more species (currently `r format(n_spp_am, big.mark = ',')` species; `r p_spp_am`% of total), such that most global analyses related to marine biodiversity to date have used AquaMaps (IUCN range map data exist for only `r format(n_spp_iucn, big.mark = ',')` unique marine species).  For the `r format(n_spp_both, big.mark = ',')` species (`r p_both_all`% of total) mapped in both datasets, we examine how well the maps align, determine several issues that lead to misalignment between predicted species distributions, and outline possible improvements.

We then reexamine two recent marine biodiversity studies - an assessment of the status of global biodiversity within the Ocean Health Index [@halpern_index_2012; @halpern_patterns_2015] and a global analysis of gaps in protection afforded by marine protected areas (MPAs) [@klein_shortfalls_2015] - as case studies to explore the implications of prioritizing one data set over the other.  The results highlight possible consequences of different data use decisions on our understanding of marine biodiversity status and protection.