### 2016-08-07 drafts/initial_version/ms_4_refs.Rmd

# Acknowledgments

<!-- Acknowledgments. List acknowledgments and funding sources. PNAS collaborates with FundRef to use a standard taxonomy of funder names. Dedications are rarely allowed. -->

We are thankful to Melanie Frazier and Julia Stewart Lowndes for their insightful comments on earlier versions of this manuscript, and to the National Center for Ecological Analysis and Synthesis for computation support.  We acknowledge financial support from the Gordon and Betty Moore Foundation.

-----

# References

<!--
References. References must be in PNAS style. Only published or in-press papers and books may be cited in the reference list. For references to in-press articles, please confirm with the cited journal that the article is in press and include a DOI number and online publication date. Unpublished abstracts of papers presented at meetings or references to "data not shown" are not permitted. References should be cited in numerical order as they appear in text, and all references, including for the SI, should be included in the main manuscript file. Because tables and figures will be inserted in the text where first cited, references in these sections should be numbered accordingly. Include the full title for each cited article. All authors (unless there are more than five) should be named in the citation. If there are more than five, list the first author's name followed by et al. Provide inclusive page ranges for journal articles and book chapters. Cite databases in the text or as footnotes.

Journal articles are cited as follows:

10. Neuhaus J-M, Sitcher L, Meins F, Jr, Boller T (1991) A short C-terminal sequence is necessary and sufficient for the targeting of chitinases to the plant vacuole. Proc Natl Acad Sci USA 88(22):10362–10366.
-->