### 2016-08-17 data_explore/data_explore.Rmd
incorporating Jamie's comments

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(maps)
library(readr)
library(data.table)
library(raster)
library(tidyr)
library(dplyr)
library(stringr)

dir_M             <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
                       'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
                       'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

if(basename(getwd()) != 'data_explore') setwd('data_explore')
source('data_explore_fxns.R')
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

```{r get_spp_list, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}
### this file created in data_prep
spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))

```

# Figures 1 a b c

``` {r load data, echo = FALSE, warning = FALSE, message = FALSE}
### use data_prep.Rmd to create this file, if it does not yet exist.

data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
spp_map_pairs <- read.csv(data_file, stringsAsFactors = FALSE) %>%
  filter(!is.na(popn_category)) %>% ### filter out unassessed species
  filter(!is.na(sm_perc))           ### filter out NA maps

### generate other lists

spp_all_file <- file.path(dir_data, 'spp_all.csv')
spp_all <- read_csv(spp_all_file)

iucn_file <- file.path(dir_data, 'spp_iucn.csv')
spp_iucn <- read_csv(iucn_file)

message('Getting IUCN maps list')
iucn_maps_file <- file.path(dir_data, 'spp_iucn_maps.csv')
spp_iucn_maps <- read_csv(iucn_maps_file) %>%
  filter(spp_group != 'hagfishes') %>%           ### not from IUCN...
  filter(sciname %in% spp_iucn$sciname)   ### ditch freshwater spp

```

``` {r breakdown by taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}

sciname_lookup <- read_csv(file.path(dir_data, 'sciname_lookups/spp_iucn_scinames.csv'))

spp_am_file <- file.path(dir_data, 'spp_am.csv')
spp_am <- read_csv(spp_am_file)

spp_am <- spp_am %>%
  left_join(read_csv(file.path(dir_M, 'git-annex/globalprep/_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>% 
              dplyr::select(am_sid = speciesid, class) %>%
              unique(), 
            by = 'am_sid')

iucn_maps_class <- spp_iucn_maps %>% 
  left_join(sciname_lookup %>% 
              dplyr::select(sciname, class) %>%
              unique(), 
            by = 'sciname')

taxa_iucn_all <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
  filter(!is.na(class)) %>% ### don't count the unknowns
  mutate(class = str_to_title(class),
         n_iucn_tot = n()) %>%        ### length of whole maps list
  group_by(class) %>%
  summarize(n_iucn_class_all  = n(),
            p_iucn_class_all  = round((n_iucn_class_all/first(n_iucn_tot) * 100), 2))

taxa_iucn_solo <- iucn_maps_class[!duplicated(iucn_maps_class$sciname), ] %>%
  filter(!is.na(class)) %>%                 ### don't count the unknowns
  filter(!sciname %in% spp_am$sciname) %>%  ### filter list to IUCN only
  mutate(class = str_to_title(class),
         n_iucn_solo = n()) %>%             ### length of IUCN only list
  group_by(class) %>%
  summarize(n_iucn_class_solo = n(),
            p_iucn_class_solo = round((n_iucn_class_solo/first(n_iucn_solo) * 100), 2))

taxa_am_all <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  mutate(class = str_to_title(class),
         n_am_tot = n()) %>%        ### length of whole AM maps list
  group_by(class) %>%
  summarize(n_am_class_all  = n(),
            p_am_class_all  = round((n_am_class_all/first(n_am_tot) * 100), 2))

taxa_am_solo <- spp_am %>%
  filter(!class %in% c('Not assigned', 'null') & !is.na(class)) %>%
  filter(!sciname %in% iucn_maps_class$sciname) %>%  ### filter list to AM only
  mutate(class = str_to_title(class),
         n_am_solo = n()) %>%       ### length of AM only list
  group_by(class) %>%
  summarize(n_am_class_solo = n(),
            p_am_class_solo = round((n_am_class_solo/first(n_am_solo) * 100), 2))

taxa_both <- spp_am %>%
  filter(sciname %in% spp_iucn_maps$sciname) %>%
  filter(!class %in% c('Not assigned', 'null')) %>%
  mutate(class = str_to_title(class),
         n_both_tot = n()) %>%
  group_by(class) %>%
  summarize(n_class_both = n(),
            p_class_both = round(n_class_both/first(n_both_tot) * 100, 2))

taxa_df <- taxa_am_all %>%
  full_join(taxa_am_solo,   by = 'class') %>%
  full_join(taxa_iucn_all,  by = 'class') %>%
  full_join(taxa_iucn_solo, by = 'class') %>%
  full_join(taxa_both,      by = 'class') %>%
  filter(!(is.na(n_iucn_class_all) & p_am_class_all < 0.5))

# knitr::kable(taxa_df %>%
#                dplyr::select(class, 
#                              AquaMaps = n_am_class_all, 
#                              IUCN = n_iucn_class_all, 
#                              Both = n_class_both))
```

``` {r plot_barline_charts_by_taxa, echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE}
### bar charts comparing number of species in each class for each data set

taxa_df1 <- taxa_df %>%
  dplyr::select(class, AquaMaps = n_am_class_solo, IUCN = n_iucn_class_solo, Paired = n_class_both) %>%
  mutate(AquaMaps = ifelse(is.na(AquaMaps), 0, AquaMaps),
         IUCN     = ifelse(is.na(IUCN),     0, IUCN),
         Paired     = ifelse(is.na(Paired),     0, Paired),
         n_total  = AquaMaps + IUCN + Paired,
         IUCN_end = 100 * (IUCN + Paired)/n_total,
         IUCN_start = 0,
         AquaMaps_end   = 100,
         AquaMaps_start = 100 * (n_total - AquaMaps - Paired)/n_total,
         Paired_end = AquaMaps_start,
         Paired_start = IUCN_end) %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2)

taxa_all_df <- taxa_df1 %>%
  group_by(dataset) %>%
  summarize(class = 'Total',
            AquaMaps_total = sum(AquaMaps) + sum(Paired),
            IUCN_total = sum(IUCN) + sum(Paired),
            n_total  = sum(n_total),
            IUCN_end = 100 * IUCN_total/n_total,
            IUCN_start = 0,
            AquaMaps_end   = 100,
            AquaMaps_start = 100 * (n_total - AquaMaps_total)/n_total,
            Paired_end = AquaMaps_start,
            Paired_start = IUCN_end) %>%
  select(-dataset) %>% distinct() %>%
  gather(key = dataset,  value = start, ends_with('start')) %>%
  gather(key = dataset2, value = end, ends_with('end')) %>%
  mutate(dataset = str_replace(dataset, '_start', ''),
         dataset2 = str_replace(dataset2, '_end', '')) %>%
  filter(dataset == dataset2) %>%
  select(-dataset2, -AquaMaps_total, -IUCN_total)

taxa_df2 <- taxa_all_df %>%
  bind_rows(taxa_all_df %>%
              mutate(class = ' ',
                     start = 0,
                     end   = 0,
                     n_total = NA)) %>%
  bind_rows(taxa_df1 %>%
              select(-IUCN, -AquaMaps, -Paired)) %>%
  mutate(class = factor(class, levels = unique(.$class), ordered = TRUE),
         n_label = ifelse(!is.na(n_total), paste0('n = ', n_total), ''),
         dataset = factor(dataset, levels = c('AquaMaps', 'Paired', 'IUCN'), ordered = TRUE))

break_nums <- seq(0, 100, 20)

dodge <- position_dodge(width = 0.1)

bar_taxa <- ggplot(taxa_df2, aes(x = class, color = dataset)) +
  ggtheme_plot + 
  theme(panel.grid.major.x = element_blank()) +
  geom_linerange(aes(ymin = start, ymax = end, size = dataset),
                 position = position_dodge(width = 0.5)) +
  scale_size_manual(values = c('IUCN' = .25, 'AquaMaps' = .25, 'Paired' = 2.5)) +
  scale_color_manual(values = c('IUCN' = 'black', 'AquaMaps' = 'grey60', 'Paired' = 'black')) +
  geom_text(aes(label = n_label, y = 102),
            size = 2,
            angle = 90, hjust = 0, color = 'grey30') +
  theme(axis.text.x = element_text(angle = 75, hjust = 1)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 130),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +
  labs(x = 'Class',
       y = 'Percent of mapped species',
       fill = 'Dataset')

print(bar_taxa)
ggsave(filename = file.path(dir_fig, 'fig1a_bar_by_taxa_all.png'), 
       width = 17.8, height = 7, units = 'cm', dpi = 600)

```

![](../figures/bar2_by_taxa_all.png)

``` {r, child = 'data_explore_global_maps.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
```

# Figure 2a b

## Quadrant map

``` {r pct_overlap_vs_ratio_of_area, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

### define 25%, 50%, and 75% quartiles for both x and y axes
area_align_mean <- mean(spp_list$area_ratio, na.rm = TRUE)
dist_align_mean <- mean(spp_list$sm_perc, na.rm = TRUE)

### define windows for labels
q1 <- c('x1' = 84, 'x2' = 98, 'y1' = 92.5, 'y2' = 97.5)
q2 <- c('x1' =  2, 'x2' = 22, 'y1' = 92.5, 'y2' = 97.5)
q3 <- c('x1' = 84, 'x2' = 98, 'y1' =  2.5, 'y2' = 7.5)
q4 <- c('x1' =  2, 'x2' = 18, 'y1' =  2.5, 'y2' = 7.5)

spp_list_labs <- spp_list %>%
  rename(x = area_ratio, y = sm_perc) %>%
  mutate(fade = FALSE,
         fade = ifelse((x > q1[1] & x < q1[2] & y > q1[3] & y < q1[4]), TRUE, fade),
         fade = ifelse((x > q2[1] & x < q2[2] & y > q2[3] & y < q2[4]), TRUE, fade),
         fade = ifelse((x > q3[1] & x < q3[2] & y > q3[3] & y < q3[4]), TRUE, fade),
         fade = ifelse((x > q4[1] & x < q4[2] & y > q4[3] & y < q4[4]), TRUE, fade))

break_nums <- seq(0, 100, 20)

### mongo plot time
scatter_quadplot <- ggplot(spp_list_labs, aes(x = x, y = y)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = area_align_mean, xmax = 100, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#4dac26")  + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymin = dist_align_mean, ymax = 100, 
           alpha = .3, fill= "#b8e186") + 
  annotate("rect", xmin = area_align_mean, xmax = 100, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#f1b6da") + 
  annotate("rect", xmax = area_align_mean, xmin =   0, ymax = dist_align_mean, ymin =   0, 
           alpha = .3, fill= "#d01c8b") +
  ### add grid lines
  geom_vline(xintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5)

### plot the data points!
scatter_quadplot <- scatter_quadplot +
  # geom_point(color = '#4d4dac', aes(alpha = fade), show.legend = FALSE) +
  # geom_point(data = spp_list_labs %>%
  #              filter(str_detect(spp_group_text, 'Corals')),
  #            color = 'red', aes(alpha = fade), 
  #            shape = 4, stroke = .5, 
  #            size = 1, show.legend = FALSE) +
  # scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1)) +
  geom_point(color = 'black', aes(alpha = fade), size = .1, show.legend = FALSE) +
  scale_alpha_manual(values = c('TRUE' = .15, 'FALSE' = 1))

### label the quadrants:
scatter_quadplot <- scatter_quadplot +
  annotate("text", x = 97, y = 95, hjust = 1, vjust = .5, size = 3, color = 'grey20', 
           fontface = 'bold.italic', label = "Well-aligned") + 
  annotate("text", x =  3, y = 95, hjust = 0, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Distribution-aligned") + 
  annotate("text", x = 97, y =  5, hjust = 1, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Area-aligned") + 
  annotate("text", x =  3, y =  5, hjust = 0, vjust = .5, size = 3, color = 'grey20',
           fontface = 'bold.italic', label = "Poorly aligned")

### Manage axis scales 
scatter_quadplot <- scatter_quadplot +
  scale_x_continuous(expand = c(0, 0), limits = c(-1, 101),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums))

### here are mean labels:
scatter_quadplot <- scatter_quadplot +
  annotate(geom = 'text', x = area_align_mean, y = 5,
           hjust = 0, vjust = 0, color = 'grey30', size = 2,
           fontface = 'bold.italic', angle = 90,
           label = sprintf('Mean = %s%%', round(area_align_mean, 1))) +
  annotate(geom = 'text', x = 5, y = dist_align_mean,
           hjust = 0, vjust = 0, color = 'grey30', size = 2,
           fontface = 'bold.italic', angle = 0,
           label = sprintf('Mean = %s%%', round(dist_align_mean, 1)))

scatter_quadplot <- scatter_quadplot +
  labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
       y = bquote('Distribution alignment '*alpha['dist']*' (%)'))

# print(scatter_quadplot)

ggsave(file.path(dir_fig, 'fig2a_scatter_quadmap.png'),
       width = 17.8, height = 9, units = 'cm', dpi = 600)

```

![](../figures/scatter_quadmap.png)

## Break down quads by species group

``` {r examine_spp_gp_quartiles, echo = FALSE, eval = TRUE}

spp_list1 <- spp_list %>%
  group_by(spp_group_text) %>%
  mutate(n_spp = n()) %>%
  ungroup()

### Create species lists by quadrant
spp_list_q1 <- spp_list1 %>%
  filter(area_ratio >= mean(area_ratio) & sm_perc >= mean(sm_perc))
spp_list_q2 <- spp_list1 %>%
  filter(area_ratio < mean(area_ratio) & sm_perc >= mean(sm_perc))
spp_list_q3 <- spp_list1 %>%
  filter(area_ratio >= mean(area_ratio) & sm_perc < mean(sm_perc))
spp_list_q4 <- spp_list1 %>%
  filter(area_ratio < mean(area_ratio) & sm_perc < mean(sm_perc))

spp_list_quads <- spp_list1 %>%
  mutate(quad = NA,
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q1$sciname, 'q1', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q2$sciname, 'q2', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q3$sciname, 'q3', quad),
         quad = ifelse(is.na(quad) & sciname %in% spp_list_q4$sciname, 'q4', quad))
write_csv(spp_list_quads, file.path(dir_data, 'spp_list_quads.csv'))

### Summarize quadrant species list by species group (e.g. CORALS1)
spp_gp_q1 <- spp_list_q1 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q1 = n())
spp_gp_q2 <- spp_list_q2 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q2 = n())
spp_gp_q3 <- spp_list_q3 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q3 = n())
spp_gp_q4 <- spp_list_q4 %>%
  group_by(spp_group_text) %>%
  summarize(n_spp = first(n_spp), n_spp_q4 = n())

### Join species group by quadrant into a single dataframe
spp_gp_quadrants <- spp_gp_q1 %>%
  full_join(spp_gp_q2, by = c('spp_group_text', 'n_spp')) %>%
  full_join(spp_gp_q3, by = c('spp_group_text', 'n_spp')) %>%
  full_join(spp_gp_q4, by = c('spp_group_text', 'n_spp')) %>%
  gather(quad, n_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
  mutate(quad = str_replace(quad, 'n_spp_', ''),
         pct_quad = n_quad/n_spp,
         n_quad   = ifelse(is.na(n_quad),   0, n_quad),
         pct_quad = ifelse(is.na(pct_quad), 0, pct_quad))

quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
                         'quad_name' = factor(c('well-matched', 'dist-matched', 
                                                'area-matched', 'poorly matched'),
                                             ordered = TRUE))

spp_gp_quadrants <- spp_gp_quadrants %>%
  left_join(spp_gp_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(spp_group_text, pct_q1 = pct_quad),
            by = 'spp_group_text') %>%
  left_join(quad_names, by = 'quad')

break_nums <- seq(0, 100, 20)

### Plot the bar chart
barchart_spp_gp_quads <- ggplot(spp_gp_quadrants %>% 
                                  transform(spp_group_text = reorder(spp_group_text, pct_q1)), 
                                aes(x = spp_group_text, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = 1) +
  scale_fill_manual(values = c('q4' = '#d01c8b', 
                               'q3' = '#f1b6da', 
                               'q2' = '#b8e186',
                               'q1' = '#4dac26'),
                    labels = quad_names$quad_name) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = break_nums/100,
                     labels = sprintf('%s%%', break_nums)) + 
  ### add grid lines; horizontal but then get flipped
  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
  ### add text
  geom_text(# data = spp_gp_quadrants %>%
              # select(spp_group_text, n_spp) %>%
              # distinct(),
            aes(label = sprintf('n = %s', n_spp), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'Taxonomic Group', 
       y = 'Percent of species by quadrant', 
       fill = 'Alignment')

print(barchart_spp_gp_quads)

ggsave(plot = barchart_spp_gp_quads,
       filename = file.path(dir_fig, 'fig2b_barchart_spp_gp_quads.png'),
       width = 17.8, height = 7, units = 'cm', dpi = 600)

write_csv(spp_gp_quadrants, file.path(dir_data, 'spp_gp_quads.csv'))

```

![](../figures/barchart_spp_gp_quads.png)

# Plots for SI

## Break down quadrants by risk category

``` {r examine_risk_quartiles, echo = FALSE, eval = TRUE}

### Summarize quadrant species list by species group (e.g. CORALS1)
spp_risk_q1 <- spp_list_q1 %>%
  group_by(popn_category) %>%
  summarize(n_spp_q1 = n())
spp_risk_q2 <- spp_list_q2 %>%
  group_by(popn_category) %>%
  summarize(n_spp_q2 = n())
spp_risk_q3 <- spp_list_q3 %>%
  group_by(popn_category) %>%
  summarize(n_spp_q3 = n())
spp_risk_q4 <- spp_list_q4 %>%
  group_by(popn_category) %>%
  summarize(n_spp_q4 = n())

### Join species group by quadrant into a single dataframe
spp_risk_quadrants <- spp_risk_q1 %>%
  full_join(spp_risk_q2, by = c('popn_category')) %>%
  full_join(spp_risk_q3, by = c('popn_category')) %>%
  full_join(spp_risk_q4, by = c('popn_category')) %>%
  gather(quad, n_spp_quad, n_spp_q1, n_spp_q2, n_spp_q3, n_spp_q4) %>% 
  group_by(popn_category) %>%
  mutate(n_tot_cat = sum(n_spp_quad)) %>%
  ungroup() %>%
  mutate(quad = str_replace(quad, 'n_spp_', ''),
         pct_quad   = n_spp_quad/n_tot_cat,
         pct_quad   = ifelse(is.na(pct_quad),   0, pct_quad))

spp_risk_quadrants1 <- spp_risk_quadrants %>%
  left_join(spp_risk_quadrants %>% 
              filter(quad == 'q1') %>%
              dplyr::select(popn_category, pct_q1 = pct_quad),
            by = 'popn_category') %>%
  mutate(quad = str_replace(quad, 'q1', 'q1 well-matched'),
         quad = str_replace(quad, 'q2', 'q2 dist-matched'),
         quad = str_replace(quad, 'q3', 'q3 area-matched'),
         quad = str_replace(quad, 'q4', 'q4 poorly matched')) %>%
  left_join(spp_list %>%
              dplyr::select(popn_category, cat_rank = category_score) %>%
              unique() %>%
              mutate(cat_rank = ifelse(popn_category %in% c('DD', 'NE'), -1, cat_rank)),
            by = 'popn_category')

### Plot the bar chart
barchart_spp_risk_quads <- ggplot(spp_risk_quadrants1 %>% 
                                  transform(popn_category = reorder(popn_category, cat_rank)), 
                                aes(x = popn_category, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = 1) +
  scale_fill_manual(values = c('q1 well-matched'   = '#4dac26', 
                               'q2 dist-matched'   = '#b8e186', 
                               'q3 area-matched'   = '#f1b6da', 
                               'q4 poorly matched' = '#d01c8b')) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1.1),
                     breaks = break_nums/100,
                     labels = sprintf('%s%%', break_nums)) + 
  ### add grid lines
  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
  ### add text
  geom_text(aes(label = sprintf('n = %s', n_tot_cat), y = 1.01), hjust = 0, 
            size = 2, 
            color = 'grey30') +
  coord_flip() +
  labs(x = 'Risk category', 
       y = 'Percent of species by quadrant', 
#       title = 'Risk category by quadrant',
       fill = 'Alignment')

# print(barchart_spp_risk_quads)
ggsave(plot = barchart_spp_risk_quads,
       filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.tif'), device = 'tiff',
       width = 17.8, height = 4.5, units = 'cm', dpi = 600)
ggsave(plot = barchart_spp_risk_quads,
       filename = file.path(dir_fig, 'si_barchart_spp_risk_quads.png'),
       width = 17.8, height = 4.5, units = 'cm', dpi = 600)

```

![](../figures/si_barchart_spp_risk_quads.png)

``` {r risk quads linear model}

spp_risk_df <- spp_list_quads %>%
  select(iucn_sid, spp_group = spp_group_text, popn_category, category_score, quad) %>%
  mutate(quad_good = ifelse(quad == 'q1', 1, 0))

risk_mdl1 <- glm(as.factor(quad_good) ~ category_score + spp_group, 
                 family = binomial,
                 data = spp_risk_df)

risk_mdl2 <- glm(as.factor(quad_good) ~ category_score, 
                 family = binomial,
                 data = spp_risk_df)

risk_mdl3 <- glm(as.factor(quad_good) ~ spp_group, 
                 family = binomial,
                 data = spp_risk_df)

BIC(risk_mdl1, risk_mdl2, risk_mdl3)
summary(risk_mdl1)
summary(risk_mdl2)
summary(risk_mdl1)

binomial logistic regression model

# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q1'])
# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q2'])
# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q3'])
# hist(spp_risk_df$category_score[spp_risk_df$quad == 'q4'])

```

-----

## AquaMaps threshold exploration

``` {r, child = 'data_explore_am_thresh.Rmd', echo = FALSE, message = FALSE, eval = TRUE}
```

-----

## Count species occurrences as proxy for data poverty

``` {r count_occurrences_by_quad, echo = FALSE, warning = FALSE}

gbif_occur <- read_csv(file.path(dir_data, 'gbif_occurrence.csv'))
obis_occur <- read_csv(file.path(dir_data, 'obis_occurrence.csv'))

totes_occur <- obis_occur %>%
  left_join(gbif_occur) %>%
  rowwise() %>%
  mutate(mean_occ = mean(c(gbif_occ, obis_occ), na.rm = TRUE)) %>%
  group_by(iucn_sid) %>%
  summarize(mean_occ = mean(mean_occ, na.rm = TRUE),
            tots_occ = sum(mean_occ, na.rm = TRUE))

spp_quads_occ <- spp_list_quads %>%
  left_join(totes_occur %>%
              dplyr::select(iucn_sid, mean_occ),
            by = 'iucn_sid')

write_csv(spp_quads_occ, file.path(dir_data, 'spp_quads_occ.csv'))

```

-----