### 2016-08-29 clip_fao/fao_clip_atl.Rmd

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(tmap)
data(World)

library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

- get aquamaps cells for all quadrant 3 species (and quadrant 4?)
- focus on any particular group? filter as appropriate
    - maybe by `occurcells` field? for low-data spp
- add lat/long from the hcaf file (or at least long)
- for each species:
    - identify east/west limits of range (between -180 and +180)
        - what about species who cross over in pacific? maybe a flag for pacific species?
        - maybe just select Pacific species and avoid Atlantic?
    - rescale 0-1?
        - here we'd be looking for what happens at the edges - is there a sudden jump from no presence to a bunch just at a single longitude?
        - with this, we could stack all the species and see the cumulative effects of the ends
    - plot end points, east/west bounds on a horizontal line/histogram/density plot, and put verticals on the longitudes of the FAO region boundaries
        - not even worrying about N/S? or maybe include just a big swath across the equator or whatever?

``` {r get_am_spp}

### get quadrant species list; filter to q3 and q4
spp_quad_list <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
  select(sciname, am_sid, iucn_sid, reviewed, occurcells, quad) %>%
  filter(quad %in% c('q3', 'q4'))

hist(spp_quad_list$occurcells)
nrow(spp_quad_list %>% filter(occurcells > 100))
### 110 out of 1078 (~10%) have more than 100 occurrence cells

### load full AM list and filter to those in q3 and q4
am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
                                   'v2016/int/am_cells_spp_prob0.csv')) %>%
  filter(am_sid %in% spp_quad_list$am_sid) %>% 
  distinct()

### join with hcaf to get longitudes (and latitudes) for each cell
hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015',
                           'csv/hcaf_truncated.csv')) %>%
  select(-csquarecode)

am_spp_latlong <- am_spp_cells %>%
  left_join(hcaf, by = 'loiczid')

### filter out species from Pacific... rough pass:
### - filter out everything west of 100 W and east of 70 E
### - this leaves southeast Pacific - coast of Chile etc

am_spp_latlong <- am_spp_latlong %>%
  filter(centerlong <= 50 & centerlong >= -65)

```

Major fishing area boundaries
- Band 1: 42 W from 35 N to 59 N (North Atlantic)
- Band 2: 40 W from 5 N to 36 N (temperate Atlantic)
- Band 3: 20 W from equator to 50 S (tropical and southern Atlantic); 30 E from 25 S on down (below Africa)

``` {r band_1}
### 35 north - look for spike at 42 W.  Not big! dammit.
am_spp_band1 <- am_spp_latlong %>%
  filter(centerlat >= 35)

am_spp_band1_lims <- am_spp_band1 %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimit),
            e_lim = max(elimit),
            n_w_lim = sum(wlimit == w_lim),
            n_e_lim = sum(elimit == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(-65, 50), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_band1_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(-42), color = 'red')

```

``` {r band_2}
### 5-36 north - look for spike at 40 W.  NONE! dammit
am_spp_band2 <- am_spp_latlong %>%
  filter(centerlat <= 36 & centerlat >= 5)

am_spp_band2_lims <- am_spp_band2 %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimit),
            e_lim = max(elimit),
            n_w_lim = sum(wlimit == w_lim),
            n_e_lim = sum(elimit == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(-65, 50), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_band2_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(-40), color = 'red')

```

``` {r band_3}
### 50 S to equator - look for spike at 20 W and 30 E
am_spp_band3 <- am_spp_latlong %>%
  filter(centerlat <= 0 & centerlat >= -70)

am_spp_band3_lims <- am_spp_band3 %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimit),
            e_lim = max(elimit),
            n_w_lim = sum(wlimit == w_lim),
            n_e_lim = sum(elimit == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  mutate(long = ifelse(long %in% c(-65, 50), NA, long)) %>%
  filter(!is.na(long))
    ### delete artificially imposed limits

tmp <- am_spp_band3_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(-20, 30), color = 'red')

```

``` {r what_species_are_these}

### which species are clipped at bottom of Africa
clipped_spp <- am_spp_band3_lims %>%
  filter(long == 30) %>%
  select(am_sid, lim) %>%
  distinct() %>%
  left_join(spp_quad_list, by = 'am_sid')

write_csv(clipped_spp, file.path(dir_git, 'clip_fao', 'clipped_s_afr.csv'))
```

``` {r plot_w_lim_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')

spp_examples <- sample_n(clipped_spp %>% filter(lim == 'w_lim'), 20)

assemble_map <- function(map_rast, spp) {
  message('...mapping ', spp)
  map_obj <- tm_shape(map_rast) +
    tm_raster(palette = 'Reds',
              colorNA = NULL,
              title = spp,
              alpha = 1) +
    tm_shape(World) +
      tm_polygons() + 
    tm_layout(basemaps = "Esri.WorldTopoMap", 
              # title.position = 'TOP', 
              legend.outside = TRUE, attr.outside = TRUE)

  return(map_obj)
}

rast_list <- vector('list', length = 20)

for (i in 1:nrow(spp_examples)) {  # i <- 1

  spp <- spp_examples[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

  r_am_spp  <-  subs(loiczid_raster, 
                     map_cells[ , c('loiczid', 'am_pres')], 
                     by = 'loiczid', 
                     which = 'am_pres', 
                     subsWithNA = TRUE)

  new_ext <- extent(c(0, 70, -70, 0))
  r_am_spp <- crop(r_am_spp, new_ext)

  rast_list[i] <- r_am_spp
  names(rast_list)[i] <- spp

  spp_plot <- assemble_map(r_am_spp, spp$sciname)

  print(spp_plot)

}

mapstack <- stack(rast_list) %>%
  calc(sum, na.rm = TRUE)

values(mapstack)[values(mapstack) == 0] <- NA

fao_polys <- rgdal::readOGR(dsn = file.path(dir_anx, '_raw_data/FAO/FAO_AREAS'),
                            layer = 'FAO_AREAS')

stack_plot <- assemble_map(mapstack, 'multi') +
  tm_shape(fao_polys) +
  tm_borders()
print(stack_plot)

```

``` {r plot_e_lim_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')

spp_examples_e <- clipped_spp %>% 
  filter(lim == 'e_lim')

assemble_map <- function(map_rast, spp) {
  message('...mapping ', spp)
  map_obj <- tm_shape(map_rast) +
    tm_raster(palette = 'Reds',
              colorNA = NULL,
              title = spp,
              alpha = 1) +
    tm_shape(World) +
      tm_polygons() + 
    tm_layout(basemaps = "Esri.WorldTopoMap", 
              # title.position = 'TOP', 
              legend.outside = TRUE, attr.outside = TRUE)

  return(map_obj)
}

rast_list <- vector('list', length = length(spp_examples_e))

for (i in 1:nrow(spp_examples_e)) {  # i <- 1

  spp <- spp_examples_e[i, ]
  map_cells <- am_spp_cells %>%
    filter(am_sid == spp$am_sid) %>%
    mutate(am_pres = 1)

  r_am_spp  <-  subs(loiczid_raster, 
                     map_cells[ , c('loiczid', 'am_pres')], 
                     by = 'loiczid', 
                     which = 'am_pres', 
                     subsWithNA = TRUE)

  new_ext <- extent(c(0, 70, -70, 0))
  r_am_spp <- crop(r_am_spp, new_ext)

  rast_list[i] <- r_am_spp
  names(rast_list)[i] <- spp

  spp_plot <- assemble_map(r_am_spp, spp$sciname)

  print(spp_plot)

}

mapstack_e <- stack(rast_list) %>%
  calc(sum, na.rm = TRUE)

values(mapstack_e)[values(mapstack_e) == 0] <- NA

fao_polys <- rgdal::readOGR(dsn = file.path(dir_anx, '_raw_data/FAO/FAO_AREAS'),
                            layer = 'FAO_AREAS')

stack_plot_e <- assemble_map(mapstack_e, 'multi') +
  tm_shape(fao_polys) +
  tm_borders()
print(stack_plot_e)

```