### 2016-08-29 manuscript-main.Rmd
new draft sent off to Courtney (and Halley and Claire)

---
title: "Aligning marine species range data to better serve science and conservation"
author: Casey C. O'Hara^1^, Jamie C. Afflerbach^1^, Courtney Scarborough^1^, Benjamin
  S. Halpern^1,2,3^
csl: refs/plos-one.csl
output:
  word_document: default
  pdf_document:
    number_sections: no
    toc: no
  html_document:
    highlight: haddock
    number_sections: no
    theme: spacelab
    toc: no
bibliography: refs/zotero_lib.bib
---

<!-- Full title	250 characters	Specific, descriptive, concise, and comprehensible to readers outside the field	 -->
<!--   Impact of Cigarette Smoke Exposure on Innate Immunity: A Caenorhabditis elegans Model -->
<!--   Solar Drinking Water Disinfection (SODIS) to Reduce Childhood Diarrhoea in Rural Bolivia: A Cluster-Randomized, Controlled Trial -->
<!-- Short title	70 characters	State the topic of the study	 -->
<!--   Cigarette Smoke Exposure and Innate Immunity -->
<!--   SODIS and Childhood Diarrhoea -->

_Author affiliation:_ <!--Include department, institution, and complete address, with the ZIP/postal code, for each author. Use superscripts to match authors with institutions. Authors are encouraged to supply their ORCID ID at submission-->

1. National Center for Ecological Analysis and Synthesis, University of California, 735 State Street Suite 300, Santa Barbara CA 93101
2. Bren School of Environmental Science and Management, University of California, Santa Barbara CA 93106
3. Department of Life Sciences, Imperial College London, Silwood Park Campus, Buckhurst Rd, Ascot, West Berkshire SL5 7PY, United Kingdom

_Corresponding author:_ <!--Corresponding author: The name, complete address, telephone number, and email address of the author to whom correspondence and proofs should be sent. Email addresses will appear in the article footnotes.-->
Casey O'Hara, National Center for Ecological Analysis and Synthesis, University of California, 735 State Street, Suite 300, Santa Barbara CA 93101; (805) 892-2500; ohara@nceas.ucsb.edu

<!--
_Keywords:_  Marine spatial planning, species distribution model, extent of occurrence, spatial ecology
-->

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width = 6, fig.height = 4, fig.path = 'Figs/',
                      echo = FALSE, message = FALSE, warning = FALSE)

# library(ggplot2)
# library(maps)

# library(raster)
# library(data.table)
library(tidyr)
library(dplyr)
library(stringr)
library(readr)

# dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
#            'Darwin'  = '/Volumes/data_edit',
#            'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

# dir_anx  <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')

if(basename(getwd()) != 'IUCN-AquaMaps') setwd('~/github/IUCN-AquaMaps')

dir_data <- 'data'
dir_fig  <- 'figures'

```

``` {r load data}
### use data_explore.Rmd to create this file, if it does not yet exist.  This
### filters out presence = Extinct from IUCN cells.

data_file <- 'data/spp_list_w_area_trimmed.csv'
spp_map_pairs <- read_csv(data_file)
### NOTE: as of Jan 31, includes all map pairs, even ones where IUCN area == 0

### generate other lists
spp_all_file <- 'data/spp_all.csv'
spp_all <- read_csv(spp_all_file) ### NOTE: contains duped scinames

# iucn_marine_file <- 'data/spp_iucn_marine.csv'
# spp_iucn_marine <- read_csv(iucn_marine_file)

iucn_maps_file <- 'data/spp_iucn_maps.csv'
spp_iucn_maps <- read_csv(iucn_maps_file) %>%
  filter(!duplicated(iucn_sid))

spp_am_file <- 'data/spp_am.csv'
spp_am <- read_csv(spp_am_file)

```

``` {r calc species numbers}
n_spp_am     <- nrow(spp_am)
n_spp_am_rev <- nrow(spp_am %>% filter(reviewed == 1))
n_spp_iucn   <- length(unique(spp_iucn_maps$iucn_sid))
n_spp_both   <- length(unique(spp_map_pairs$iucn_sid))
n_spp_all    <- n_spp_am + n_spp_iucn - n_spp_both

p_spp_am     <- round(n_spp_am/n_spp_all    * 100, 1)
p_spp_iucn   <- round(n_spp_iucn/n_spp_all  * 100, 1)
p_both_am    <- round(n_spp_both/n_spp_am   * 100, 1)
p_both_iucn  <- round(n_spp_both/n_spp_iucn * 100, 1)
p_both_all   <- round(n_spp_both/n_spp_all  * 100, 1)

spp_list <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv'))
area_align_mean <- round(mean(spp_list$area_ratio, na.rm = TRUE), 1)
dist_align_mean <- round(mean(spp_list$sm_perc, na.rm = TRUE), 1)

```

``` {r captions_for_barchart_by_taxa, echo = FALSE}
n_spp <- spp_map_pairs %>%
  mutate(spp_group = ifelse(str_detect(spp_group, 'CORAL'), 'CORAL', spp_group)) %>%
  group_by(spp_group) %>%
  summarize(n_spp = n()) %>%
  arrange(desc(n_spp))

n_corals   <- nrow(spp_map_pairs %>% 
                     filter(str_detect(spp_group, 'CORAL')))
n_bonyfish <- nrow(spp_map_pairs %>% 
                     filter(spp_group %in% c('WRASSES', 'BUTTERFLYFISH', 'PARROTFISH',
                                             'COMBTOOTHBLENNIES', 'SEABREAMS_PORGIES', 
                                             'PUFFERFISH', 'GROUPERS', 'TUNAS_BILLFISHES', 
                                             'SURGEONFISH_TANGS_UNICORNFISH', 'ANGELFISH', 
                                             'DAMSELFISH', 'BONEFISH_TARPONS')))
```

``` {r captions_for_tally_quad_counts, echo = FALSE}
dist_median <- round(median(spp_map_pairs$sm_perc,    na.rm = TRUE), 1)
area_median <- round(median(spp_map_pairs$area_ratio, na.rm = TRUE), 1)
q1_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio >= area_median)
q2_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio >= area_median)
q3_count <- sum(spp_map_pairs$sm_perc >= dist_median & spp_map_pairs$area_ratio <  area_median)
q4_count <- sum(spp_map_pairs$sm_perc <  dist_median & spp_map_pairs$area_ratio <  area_median)
```

-----

``` {r abstract, child = 'ms_0_abst.Rmd'}
```

<!--
| Section        |   Word #   | Char/spc # |
| :------        |    ------: |    ------: |
| Abstract       |  169 words |     - char |
-->

-----

``` {r intro, child = 'ms_1_intro.Rmd'}
```

``` {r methods, child = 'ms_2_methods.Rmd'}
```

``` {r results_discussion, child = 'ms_3_res_disc_concl.Rmd'}
```

<!--
| Section                   |   Word #   | Char/spc # |
| :------                   |    ------: |    ------: |
| Introduction              |  653 words |     - char |
| Methods                   |  878 words |     - char |
| Results/Discussion        | 2150 words |     - char |
| Conclusions               |  499 words |     - char |
| total body                | 4187 words |     - char |
-->

-----

``` {r references, child = 'ms_4_refs.Rmd'}
```

<!--
| Figure  |  w x h (px) | w x h (cm)  | char # |
| :------ |     ------: |     ------: |------: |
| 1 A - C | 2102 x 1536 | 17.8 x 13.0 |   4680 |
| caption |             |             |    367 | 
| 2 A - B | 2102 x 1940 | 17.8 x 16.5 |   5940 |
| caption |             |             |    934 | 
|       3 | 1027 x  531 |  8.7 x  4.5 |    810 |
| caption |             |             |    744 | 
|       4 | 1346 x  590 | 11.4 x  5.0 |   1800 |
| caption |             |             |    870 | 
|  total  |             |             |  16145 |
-->