### 2016-08-29 ms_0_abst.Rmd

## Abstract

<!-- 
The Abstract comes after the title page in the manuscript file. The abstract text is also entered in a separate field in the submission system.

The Abstract should:
  Describe the main objective(s) of the study
  Explain how the study was done, including any model organisms used, without methodological detail
  Summarize the most important results and their significance
  Not exceed 300 words
Abstracts should not include:
  Citations
  Abbreviations, if possible
-->

Species distribution data provide the foundation for a wide range of ecological research studies and conservation management decisions, yet most species ranges remain unknown, and existing range maps often suffer from data limitations and inconsistencies.  AquaMaps and the International Union for Conservation of Nature (IUCN) are two distinct efforts to map marine species distributions at a global scale.  Together these databases represent `r format(n_spp_all, big.mark = ',')` species (`r format(p_spp_am, big.mark = ',')`% within AquaMaps, `r format(p_spp_iucn, big.mark = ',')`% within IUCN), with only `r format(n_spp_both, big.mark = ',')` shared species.  Here we examine differences in predicted species ranges between the two datasets and find that these misalignments mainly result from divergent methodologies that introduce differing frequencies of commission and omission errors.  We illustrate the scientific and management implications of these differences by repeating a global analysis of gaps in coverage of marine protected areas, and find significantly different results depending on how the two datasets are used.  While we suggest methods to address specific errors in each spatial dataset, it remains essential to understand the implications of dataset differences for conservation planning and decision-making.