### 2016-09-02 clip_fao/fao_clip_pac.Rmd

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(tmap)
data(World)
library(maptools)
data(wrld_simpl)

library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

- get aquamaps cells for all quadrant 3 species (and quadrant 4?)
- focus on any particular group? filter as appropriate
    - maybe by `occurcells` field? for low-data spp
- add lat/long from the hcaf file (or at least long)
- for each species:
    - identify east/west limits of range (between -180 and +180)
        - what about species who cross over in pacific? maybe a flag for pacific species?
        - maybe just select Pacific species and avoid Atlantic?
    - rescale 0-1?
        - here we'd be looking for what happens at the edges - is there a sudden jump from no presence to a bunch just at a single longitude?
        - with this, we could stack all the species and see the cumulative effects of the ends
    - plot end points, east/west bounds on a horizontal line/histogram/density plot, and put verticals on the longitudes of the FAO region boundaries
        - not even worrying about N/S? or maybe include just a big swath across the equator or whatever?

``` {r get_am_spp}

### get quadrant species list; filter to q3 and q4
spp_quad_list <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
  select(sciname, am_sid, iucn_sid, reviewed, occurcells, quad) %>%
  filter(quad %in% c('q3', 'q4'))

nrow(spp_quad_list %>% filter(occurcells > 100))
### 110 out of 1078 (~10%) have more than 100 occurrence cells

### load full AM list and filter to those in q3 and q4
am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
                                   'v2016/int/am_cells_spp_prob0.csv')) # %>%
  # filter(am_sid %in% spp_quad_list$am_sid) %>% 
  # distinct()

am_spp_list <- read_csv(file.path(dir_git, 'data/spp_all.csv')) %>%
  filter(!is.na(am_sid)) %>%
  arrange(reviewed) %>%
  filter(!duplicated(am_sid)) %>%
  select(am_sid, sciname, reviewed) %>%
  distinct()

message('loading hcaf and adjusting to se asia')
### join with hcaf to get longitudes (and latitudes) for each cell
hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015',
                           'csv/hcaf_truncated.csv')) %>%
  select(-csquarecode, -cellarea, -oceanarea) %>%
  mutate(clongadj  = ifelse(centerlong < 0, centerlong + 360, centerlong),
         wlimitadj = ifelse(wlimit < 0, wlimit + 360, wlimit),
         elimitadj = ifelse(elimit < 0, elimit + 360, elimit)) %>%
  filter(clongadj >= 60 & clongadj <= 240)

### filter out species from Atlantic... rough pass:
### - filter out everything west of 30 E and east of 70 W
message('binding hcaf to am_spp_latlong')

am_spp_latlong <- hcaf %>%
  left_join(am_spp_cells, by = 'loiczid')

```

Major fishing area boundaries
- 175W north from 25S
- 120W from 5N to 60 S
- 150E south from 25S (south of Australia)
- 80E from 0N to 55S
- 30 E from Africa on down

``` {r northmost_band, eval = FALSE}
### 5N north     - look for spike at 175 W (185E)
am_spp_north <- am_spp_latlong %>%
  filter(centerlat >= 5)

am_spp_north_lims <- am_spp_north %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n()) %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  filter(!(long <= 70 | long >= 200))
    ### delete artificially imposed limits

tmp <- am_spp_north_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(77, 185), color = 'red')

```

``` {r central_band}
### 25 S to 5 N  - look for spikes at 120W (240E), 175W (185E), 80E (and 77E)
am_spp_central <- am_spp_latlong %>%
  filter(centerlat <= 5 & centerlat >= -25)

message('summarizing...')
am_spp_central_lims <- am_spp_central %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n())

message('gathering...')
am_spp_central_lims <- am_spp_central_lims %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  select(-lim_n) %>%
  filter(!(long <= 70 | long >= 200))
    ### delete artificially imposed limits

write_csv(am_spp_central_lims, file.path(dir_git, 'clip_fao/am_spp_central_lims.csv'))

tmp <- am_spp_central_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(77, 185), color = 'red')

```

``` {r what_species_are_these}

clipped_spp <- am_spp_central_lims %>%
  # bind_rows(am_spp_north_lims) %>%
  # filter(long %in% c(77, 185)) %>%
  filter(long == 185) %>%
  select(am_sid, lim) %>%
  distinct() %>%
  left_join(am_spp_list, by = 'am_sid')

write_csv(clipped_spp, file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))

```

``` {r plot_setup, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

loiczid_raster <- raster('data/rasters/loiczid_raster.tif')
fao_polys <- rgdal::readOGR(dsn = file.path(path.expand(dir_git), 'clip_fao/fao_rgns'),
                            layer = 'fao_rgns2')

rast_ext  <- extent(c(-30, 330, -75, 75))
fao_polys <- crop(fao_polys, rast_ext)

# wrld_simpl_shift <- wrld_simpl %>%
#   raster::shift(360)
# 
# World1 <- rgeos::gUnion(wrld_simpl, wrld_simpl_shift, byid = FALSE) %>%
#   crop(rast_ext) %>%
#   SpatialPolygonsDataFrame(data.frame(x = 1))
# 
# maptools::writePolyShape(World1, 'clip_fao/fao_rgns/world1')

World1 <- maptools::readShapePoly('clip_fao/fao_rgns/world1')

rast_spp_shift <- rast_spp %>%
  raster::shift(360)

rast_spp_e <- merge(rast_spp, rast_spp_shift)

assemble_map <- function(map_rast, spp) {
  message('...mapping ', spp)
  map_obj <- tm_shape(map_rast) +
    tm_raster(palette = 'Reds',
              colorNA = NULL,
              title = spp,
              alpha = 1) +
    tm_shape(World1) +
      tm_polygons() + 
    tm_layout(basemaps = "Esri.WorldTopoMap", 
              # title.position = 'TOP', 
              legend.outside = TRUE, attr.outside = TRUE)

  return(map_obj)
}
```

``` {r plot_spp_e, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))

spp_examples_e <- clipped_spp %>%
  filter(lim == 'w_lim')

message('grouping affected species and summing to loiczid')
### group affected spp by loiczid and sum

# spp_cells_clipped <- am_spp_cells %>%
#   filter(am_sid %in% spp_examples_e$am_sid) %>%
#   group_by(loiczid) %>%
#   summarize(n_spp = n())
# write_csv(spp_cells_clipped, file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))

spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))

### create raster, then shift raster

rast_spp  <-  subs(loiczid_raster, 
                   spp_cells_clipped[ , c('loiczid', 'n_spp')], 
                   by = 'loiczid', 
                   which = 'n_spp', 
                   subsWithNA = TRUE)

rast_spp_shift <- rast_spp %>%
  raster::shift(360)

rast_spp_e <- merge(rast_spp, rast_spp_shift)

values(rast_spp_e)[values(rast_spp_e) == 0] <- NA
# values(rast_spp_e)[is.na(values(rast_spp_e))] <- 0
rast_spp_e  <- crop(rast_spp_e, rast_ext)

rast_spp_map_e <- 
  tm_shape(rast_spp_e) +
    tm_raster(palette = 'Reds',
              breaks = c(1, 15, 100, 200, 300, Inf),
              colorNA = 'lightblue1',
              textNA = 'Zero',
              title = 'Species count',
              alpha = 1) +
  tm_shape(World1) +
    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
  tm_shape(fao_polys) +
    tm_borders(col = 'grey40', lwd = 0.25) +
    tm_text('F_AREA_1',
            col = 'orangered4',
            fontface = 'bold.italic',
            fontfamily = 'Helvetica',
            size = .35) +
  tm_layout(# basemaps = "Esri.WorldTopoMap", 
            frame = TRUE,
            legend.text.size = .35,
            legend.title.size = .45,
            # title.position = 'TOP', 
            legend.outside = FALSE,
            legend.position = c('right', 'top'),
            legend.bg.color = 'white',
            legend.bg.alpha = .8,
            attr.outside = TRUE)

print(rast_spp_map_e)

save_tmap(rast_spp_map_e, 
          file.path(dir_fig, 'map_am_edge_map.png'), 
          width = 19, 
          # height = 6, 
          units = 'cm', dpi = 600)

```

``` {r plot_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))

spp_examples_w <- clipped_spp %>%
  filter(lim == 'e_lim')

message('grouping affected species and summing to loiczid')
### group affected spp by loiczid and sum
spp_cells_clipped <- am_spp_cells %>%
  filter(am_sid %in% spp_examples_w$am_sid) %>%
  group_by(loiczid) %>%
  summarize(n_spp = n())

### create raster, then shift raster

rast_spp  <-  subs(loiczid_raster, 
                   spp_cells_clipped[ , c('loiczid', 'n_spp')], 
                   by = 'loiczid', 
                   which = 'n_spp', 
                   subsWithNA = TRUE)

rast_spp_shift <- rast_spp %>%
  raster::shift(360)

rast_spp_w <- merge(rast_spp, rast_spp_shift)

values(rast_spp_w)[values(rast_spp_w) == 0] <- NA
# values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
rast_spp_w <- crop(rast_spp_w, extent(c(85, 220, -59, 50)))

rast_spp_map_w <- 
  tm_shape(rast_spp_w) +
    tm_raster(palette = 'Reds',
              breaks = c(1, 50, 100, 500, 1000, 2000, Inf),
              colorNA = 'lightblue1',
              textNA = 'Zero',
              title = 'Species count',
              alpha = 1) +
  tm_shape(World1) +
    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
  tm_shape(fao_polys) +
    tm_borders(col = 'grey40', lwd = 0.25) +
    tm_text('F_CODE_1',
            col = 'orangered4',
            fontface = 'bold.italic',
            fontfamily = 'Helvetica',
            size = .35) +
  tm_layout(basemaps = "Esri.WorldTopoMap", 
            frame = FALSE,
            legend.text.size = .35,
            legend.title.size = .45,
            # title.position = 'TOP', 
            legend.outside = TRUE,
            # legend.position = c('right', 'bottom'),
            # legend.bg.color = 'white',
            # legend.bg.alpha = .5,
            attr.outside = TRUE)

print(rast_spp_map_w)

save_tmap(rast_spp_map_w, file.path(dir_fig, 'map_am_edge_w.png'), 
          width = 9, height = 6, units = 'cm', dpi = 600)

```

``` {r combine_rasts}

stack_plot_both <- 
  tm_shape(rast_spp_w) +
    tm_raster(palette = 'Greens',
              colorNA = NULL,
              title = 'n_spp',
              alpha = 1) +
  tm_shape(rast_spp_e) +
    tm_raster(palette = 'Reds',
              colorNA = NULL,
              title = 'n_spp') +
  tm_shape(World1) +
    tm_polygons() + 
  tm_shape(fao_polys_e) +
    tm_borders() +
  tm_layout(basemaps = "Esri.WorldTopoMap", 
            frame = FALSE,
            # title.position = 'TOP', 
            legend.outside = TRUE, attr.outside = TRUE)

print(stack_plot_both)
```