### 2016-09-23 clip_depth/coral_maps.Rmd
wrapping up edits from JA, BH, CS updating fig captions

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

```{r get_bathy_raster, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

### get shapefile
### rasterize to half-degree grid; or raster::extract

if(!file.exists(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'))) {

  poly_bath <- readOGR(dsn = file.path(path.expand(dir_git), 'clip_depth/ne_10m_bathymetry_K_200'), 
                       layer = 'ne_10m_bathymetry_K_200', 
                       stringsAsFactors = TRUE)

  ### LOICZID cell ID numbers - used in plotting rasters
  loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

  bath_list <- raster::extract(loiczid_raster, poly_bath, 
                             weights = TRUE, normalizeWeights = FALSE,
                             progress = 'text')
  bath_df <- lapply(bath_list, function(x) {as.data.frame(x)}) %>%
    bind_rows() %>%
    rename(loiczid = value) %>%
    group_by(loiczid) %>%
    summarize(weight = sum(weight))

  ### generate raster?
  bath_raster <- subs(loiczid_raster, bath_df, by = 'loiczid', which = 'weight')
  values(bath_raster)[values(bath_raster) > 5] <- NA

  writeRaster(bath_raster, file.path(dir_git, 'clip_depth', 'bathy_cells.tif'))

  write_csv(bath_df, file.path(dir_git, 'clip_depth', 'bathy_cells.csv'))
  # bath_df <- read_csv(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'), col_types = 'dd')
  ### probably want to filter to just 100% 200m+
}

```

``` {r extract_IUCN_cells_for_coral_pairs}

bath_df <-  read_csv(file.path(dir_git, 'clip_depth', 'bathy_cells.csv'), col_types = 'dd') %>%
  filter(weight >= 1) %>% ### cell must be 100% within the deep polygon to count as deep
  mutate(deep = 1) %>%
  dplyr::select(-weight)

spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))

### load iucn spp_cells
coral_files <- list.files(file.path(dir_anx, 'v2016', 'iucn_intersections'), 
                          pattern = 'CORAL*.*',
                          full.names = TRUE)
coral_cells_iucn <- lapply(coral_files, read_csv) %>%
  bind_rows() %>%
  dplyr::select(iucn_sid, loiczid = LOICZID) %>%
  # filter(iucn_sid %in% spp_depth$iucn_sid[spp_depth$paired_map]) %>%
  left_join(bath_df, by = 'loiczid') %>%
  mutate(deep = ifelse(is.na(deep), 0, deep))

write_csv(coral_cells_iucn, file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv'))
length(unique(coral_cells_iucn$iucn_sid))
### NOTE: only 473 coral species with IUCN maps
```

``` {r}
### find spp depth limits
### focus just on corals?

if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))) {

  library(jsonlite)
  api_key <- 'fb71ae836f415f04f41176f6d30c4a9e4cea620d46b9e5021bf2fb142ea51bf5'

  get_from_api <- function(url, param) {
      api_info <- fromJSON(sprintf(url, param, api_key)) %>%
        data.frame(stringsAsFactors = FALSE)
  }

  mc_get_from_api <- function(url, param_vec) {
    numcores = ifelse(Sys.info()[['nodename']] == 'mazu', 12, 1)
    out_df <- parallel::mclapply(param_vec, 
                            function(x) get_from_api(url, x),
                            mc.cores = numcores) %>% 
      bind_rows() 
    out_df <- out_df %>%
      setNames(names(.) %>%
                 str_replace('result.', ''))
  }

  ### get paired species list
  paired_maps_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
  spp_map_pairs <- read_csv(paired_maps_file) %>%
    filter(!is.na(popn_category)) %>% ### filter out unassessed species
    filter(!is.na(sm_perc))           ### filter out NA maps

  ### get full species list
  spp_all_file <- file.path(dir_data, 'spp_all.csv')
  spp_all <- read_csv(spp_all_file)

  coral_spp_sid <- spp_all %>%
    filter(str_detect(tolower(spp_group), 'coral')) %>%
    .$iucn_sid %>%
    unique()

  spp_narr_url <- 'http://apiv3.iucnredlist.org/api/v3/species/narrative/id/%s?token=%s'
  spp_narr     <- mc_get_from_api(spp_narr_url, coral_spp_sid)

  write_csv(spp_narr, file.path(dir_git, 'clip_depth', 'coral_spp_narratives.csv'))

  spp_depth <- spp_narr %>%
    dplyr::select(iucn_sid = species_id, habitat) %>%
    mutate(paired_map = (iucn_sid %in% spp_map_pairs$iucn_sid))

  spp_depth1 <- spp_depth %>%
    mutate(depth = str_extract_all(habitat, '\\d++(?:\\.\\d++)?(?=\\s*+(?:m))')) %>%
    unnest(depth) %>%
    group_by(iucn_sid) %>%
    summarize(max_depth = max(as.integer(depth)))

  spp_depth <- spp_depth %>%
    left_join(spp_depth1, by = 'iucn_sid') %>%
    mutate(shallow_str = str_detect(tolower(habitat), 'shallow'),
           shallow = max_depth <= 200 | shallow_str)

  write_csv(spp_depth, file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))

} else 
  spp_depth <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_depth.csv'))

summary(spp_depth)
 #    iucn_sid        habitat          paired_map        max_depth      shallow_str     shallow       
 # Min.   :132859   Length:562         Mode :logical   Min.   :  1.00   Mode :logical   Mode:logical  
 # 1st Qu.:133081   Class :character   FALSE:174       1st Qu.: 20.00   FALSE:194       TRUE:544      
 # Median :133284   Mode  :character   TRUE :388       Median : 20.00   TRUE :367       NA's:18       
 # Mean   :133283                      NA's :0         Mean   : 27.55   NA's :1                       
 # 3rd Qu.:133486                                      3rd Qu.: 30.00                                 
 # Max.   :133703                                      Max.   :200.00                                 
 #                                                     NA's   :46

nrow(spp_depth %>% filter(max_depth > 100))
# [1] 5
### of 562 species, only 5 are listed as occurring below 100 m; none are below 200 m.
### NOTE: Many of these 562 species have no maps in IUCN; they are mapped in AquaMaps though.

spp_depth_na <- spp_depth %>%
  filter(is.na(shallow))

### 18 species (of 562) have no depth information; one is listed as being dredged from 13 fathoms (~ 25 m)
quantile(spp_depth$max_depth, seq(.9, 1, .01), na.rm = TRUE)
 #   90%    91%    92%    93%    94%    95%    96%    97%    98%    99%   100% 
 # 40.00  44.65  50.00  50.00  50.00  61.25  70.80  78.20  84.10 100.00 200.00

spp_d1 <- spp_depth %>% filter(iucn_sid %in% coral_cells_iucn$iucn_sid) ### mapped spp only
quantile(spp_d1$max_depth, seq(.9, 1, .01), na.rm = TRUE)
 #   90%    91%    92%    93%    94%    95%    96%    97%    98%    99%   100% 
 # 40.00  45.00  50.00  50.00  50.00  54.60  70.00  75.56  82.12 100.00 152.00

```

``` {r plot_coral_composite_map}

library(tmap)
data(World)

loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
loiczid_raster       <- raster(loiczid_raster_file)
names(loiczid_raster) <- 'loiczid'

coral_cells_iucn <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv'))

coral_comp <- coral_cells_iucn %>%
  dplyr::select(-iucn_sid) %>%
  mutate(deep = deep * 200) %>%
  distinct() %>%
  as.data.frame()

rast_coral_comp <- raster::subs(x = loiczid_raster, y = coral_comp %>% dplyr::select(loiczid, deep), 
                     by = 'loiczid', which = 'deep', 
                     subsWithNA = TRUE) %>%
  crop(extent(c(-180, 180, -63, 85)))

map_coral_comp <- 
  tm_shape(rast_coral_comp) +
    tm_raster(breaks = c(0, 200, Inf),
              palette = c('lightblue', 'cyan4'),
              colorNA = NULL,
              title = 'depth (meters)',
              alpha = 1) +
  tm_shape(World) +
    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
  tm_layout(basemaps = "Esri.WorldTopoMap", 
            legend.text.size = .8,
            legend.title.size = .9,
            # title.position = 'TOP', 
            legend.outside = FALSE, 
            legend.position = c('left', 'bottom'),
            legend.bg.color = 'white',
            legend.bg.alpha = .5,
            attr.outside = TRUE)

print(map_coral_comp)

save_tmap(map_coral_comp,
          filename = file.path(dir_fig, 'coral_map.png'), 
          width = 19, height = 8, units = 'cm', dpi = 600)

```

``` {r calc_areas_for_iucn_corals}
### calc areas
### read hcaf to get areas; left join to coral_cells_iucn
### group_by iucn_sid and depth, summarize area (deep and shallow)
### group_by iucn_sid, summarize total area
coral_cells_iucn <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
  mutate(deep = ifelse(deep == 1, 'deep', 'shallow'))

hcaf <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
  dplyr::select(loiczid, cellarea, oceanarea)

iucn_coral_areas <- coral_cells_iucn %>%
  left_join(hcaf, by = 'loiczid') %>%
  group_by(iucn_sid, deep) %>%
  summarize(area = sum(oceanarea)) %>%
  group_by(iucn_sid) %>%
  mutate(totarea = sum(area),
         pct_area = area/totarea) %>%
  filter(deep == 'shallow') %>%
  rename(shallow_pct = pct_area) %>%
  mutate(deep_pct = 1 - shallow_pct)

### get species list
data_file <- file.path(dir_data, 'spp_list_w_area_trimmed.csv')
spp_map_pairs <- read_csv(data_file) %>%
  filter(!is.na(popn_category)) %>%   ### filter out unassessed species
  filter(!is.na(sm_perc))             ### filter out NA maps

iucn_coral_areas <- iucn_coral_areas %>%
  left_join(spp_depth,     by = 'iucn_sid') %>%
  left_join(spp_map_pairs, by = 'iucn_sid')

write_csv(iucn_coral_areas, file.path(dir_git, 'clip_depth', 'iucn_coral_areas.csv'))

```

``` {r calc_overlaps}
if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))) {

  ### get species list
  coral_cells_iucn_shallow <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
    filter(deep == 0)

  spp_coralmaps <- read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
    filter(iucn_sid %in% coral_cells_iucn_shallow$iucn_sid) %>%
    dplyr::select(am_sid, sciname, iucn_sid)

  ### copy aquamaps cells to local file, keeping only corals in the paired maps
  if(!file.exists(file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))) {
    coral_cells_am <- read_csv(file.path(dir_anx, '../_raw_data/aquamaps/d2015/csv', 'hcaf_sp_native_trunc.csv')) %>%
      rename(am_sid = speciesid, prob = probability) %>%
      filter(am_sid %in% spp_coralmaps$am_sid) 
    write_csv(coral_cells_am, file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))
  }

  coral_cells_am <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_am.csv'))

  loiczid_raster_file  <- file.path(dir_git, 'shiny_am_iucn/data/loiczid_raster.tif')
  loiczid_raster       <- raster(loiczid_raster_file)
  names(loiczid_raster) <- 'loiczid'

  spp_coralmaps <- spp_coralmaps %>%
    mutate(area_total = NA,  area_am = NA,  area_iucn = NA,
           area_overlap = NA,
           sm_perc    = NA, #filled with % of smaller range that is within the larger range
           sm_range   = NA) #what range is smaller

  i         <- 0 ### initialize the counting index
  am_thresh <- 0 ### set the aquamaps presence threshold
  sid_vector <- unique(spp_coralmaps$iucn_sid)
  message(sprintf('There are a total of %s distinct species, by IUCN SID', 
                  length(sid_vector)))

  for (sid in sid_vector) { 
    # sid <- spp_coralmaps$iucn_sid[1]
    i <- i + 1
    sids <- spp_coralmaps %>% 
      dplyr::select(am_sid, iucn_sid, sciname) %>%
      filter(iucn_sid == sid)
    rowindex <- (spp_coralmaps$iucn_sid == sid)

    message(sprintf('%s. Processing species iucn_sid: %s \n   am_sid(s):  %s \n   sciname(s): %s', 
                    i, sid, paste(sids$am_sid, collapse = ', '), paste(sids$sciname, collapse = ' ')))
    map_iucn <- get_spp_map(sid, coral_cells_iucn_shallow) %>%
      mutate(iucn_pres = ifelse(area_iucn > 0, 1, 0))
    map_am   <- get_spp_map(sids$am_sid, coral_cells_am) %>%
      mutate(am_pres = ifelse(am_prob >= am_thresh, 1, 0))

    ### rasterize aquamaps map
    if(any(duplicated(map_am$loiczid))) message('Duplicated cells in am spp map')
    map_am <- map_am %>%
      dplyr::select(loiczid, am_pres) %>% 
      unique()

    r_am_spp <- subs(x = loiczid_raster, y = map_am %>% dplyr::select(loiczid, am_pres), 
                     by = 'loiczid', which = 'am_pres', 
                     subsWithNA = TRUE)

    ### rasterize IUCN map
    if(TRUE %in% duplicated(map_iucn$loiczid)) stop('Duplicated cells in IUCN map')
    r_iucn_spp <- subs(loiczid_raster, map_iucn %>% dplyr::select(loiczid, iucn_pres), 
                       by = 'loiczid', which = 'iucn_pres', 
                       subsWithNA = TRUE)

    ### Start calculating areas etc.
    area_iucn <- area(r_iucn_spp, na.rm = TRUE) %>%
      cellStats(., stat = 'sum')
    area_am   <- area(r_am_spp, na.rm = TRUE) %>%
      cellStats(., stat = 'sum')
    spp_coralmaps[rowindex, ]$area_iucn <- area_iucn
    spp_coralmaps[rowindex, ]$area_am   <- area_am

    ### Get total area (IUCN and AM combined)
    all_cells <- data.frame(loiczid = unique(c(map_iucn$loiczid, map_am$loiczid)),
                           spp_pres = 1)
    r_all_cells <- subs(loiczid_raster, all_cells, 
                       by = 'loiczid', which = 'spp_pres', 
                       subsWithNA = TRUE)
    area_total  <- area(r_all_cells, na.rm = TRUE) %>% cellStats(., stat = 'sum')
    spp_coralmaps[rowindex, ]$area_total <- area_total

    ### Get total overlap (IUCN and AM together). If area_iucn = 0, just assign 0
    if(area_iucn != 0) {
      overlap_cells <- unique(map_iucn$loiczid[map_iucn$loiczid %in% map_am$loiczid])
      if(length(overlap_cells) > 0) {
        overlap_cells <- data.frame(loiczid = overlap_cells,
                                  spp_pres = 1)
        r_overlap_cells <- subs(loiczid_raster, overlap_cells, 
                                by = 'loiczid', which = 'spp_pres', 
                                subsWithNA = TRUE)
        area_overlap <- area(r_overlap_cells, na.rm = TRUE) %>% cellStats(., stat = 'sum')
      } else area_overlap <- 0 ### if no overlapping cells, assign zero
    } else area_overlap <- 0 ### if no IUCN map cells, assign zero
    spp_coralmaps[rowindex, ]$area_overlap <- area_overlap

    message(sprintf('   Species iucn_sid %s: smaller range is %s. Smaller-in-larger is %.2f%%, A_sm:A_lg is %.2f%%.',
                    sid, ifelse(area_am < area_iucn, 'AquaMaps', 'IUCN'), 
                    ifelse(area_am < area_iucn, area_overlap/area_am*100, area_overlap/area_iucn*100),
                    ifelse(area_am < area_iucn, area_am/area_iucn*100, area_iucn/area_am*100)))
  }

  spp_coralmaps <- spp_coralmaps %>%
    mutate(sm_range   = ifelse(area_am < area_iucn, 'AM', 'IUCN'),
           sm_perc    = ifelse(area_am < area_iucn, (area_overlap / area_am) * 100, (area_overlap / area_iucn) * 100),
           area_ratio = area_am / area_iucn,
           area_ratio = ifelse(area_ratio > 1, 1 / area_ratio, area_ratio),
           area_ratio = area_ratio * 100)

  spp_coralmaps <- spp_coralmaps %>%
    mutate(lg_area = ifelse(area_am > area_iucn, area_am, area_iucn),
           sm_area = ifelse(area_am < area_iucn, area_am, area_iucn),
           lg_area_pct = 100 * lg_area / max(area_total, na.rm = TRUE)) %>% 
             ### use largest data set as approx for total ocean range
    dplyr::select(-lg_area, -sm_area)

  spp_coralmaps <- spp_coralmaps %>%
    mutate(area_total  = round(area_total, 1),
           area_am     = round(area_am, 1),
           area_iucn   = round(area_iucn, 1),
           sm_perc     = round(sm_perc, 3),
           area_ratio  = round(area_ratio, 3),
           lg_area_pct = round(lg_area_pct, 3))

  print(head(spp_coralmaps))

  ### combine old and new area dataframes
  spp_coralmaps <- spp_coralmaps %>%
    dplyr::select(iucn_sid, am_sid, sciname, area_am, 
           area_iucn_shallow = area_iucn, area_overlap_shallow = area_overlap, 
           sm_perc_shallow = sm_perc, sm_range_shallow = sm_range, area_ratio_shallow = area_ratio) %>%
    left_join(read_csv(file.path(dir_data, 'spp_list_w_area_trimmed.csv')) %>%
                dplyr::select(am_sid, iucn_sid, 
                       area_iucn_raw = area_iucn, area_overlap_raw = area_overlap, 
                       sm_perc_raw = sm_perc, sm_range_raw = sm_range, area_ratio_raw = area_ratio),
              by = c('iucn_sid', 'am_sid'))

  write_csv(spp_coralmaps, file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))
}

```

``` {r generate_area_plots}

spp_coral_clip <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_areas.csv')) %>%
  dplyr::select(iucn_sid, shallow_pct, deep_pct, totarea, shallow_area = area) %>%
  mutate(shallow_pct = 100 * shallow_pct,
         deep_pct    = 100 * deep_pct)

spp_small_effect <- spp_coral_clip %>%
  filter(shallow_pct > 80) 
### only four out of 481 with such small effects; one of those did not change at all.

summary(spp_coral_clip$shallow_pct)
  # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 22.40   48.20   51.71   52.37   56.68  100.00

scatterplot1 <- ggplot(spp_coral_clip, aes(x = totarea, y = shallow_area)) +
  geom_point(size = 0.25, color = 'black') +
  geom_abline(color = 'red')
scatterplot1

### for area plot, arrange from highest shallow pct to lowest (best to worst)
spp_coral_clip <- spp_coral_clip %>%
  arrange(desc(shallow_pct))

spp_coral_clip$l_to_h <- c(1:nrow(spp_coral_clip))

spp_coral_clip2 <- spp_coral_clip %>%
  gather(depth, pct_value, c(shallow_pct, deep_pct)) %>%
  mutate(depth = ifelse(depth == 'shallow_pct', '0 - 200 m', '200 m +'))

break_nums <- seq(0, 100, 20)
num_spp <- max(spp_coral_clip2$l_to_h)
coral_area_plot <- ggplot(spp_coral_clip2, aes(x = l_to_h, y = pct_value, fill = depth)) +
  ggtheme_plot +
  geom_area(show.legend = FALSE) +
  scale_fill_manual(values = c('0 - 200 m' = 'grey90', '200 m +' = 'grey10')) +
  geom_hline(yintercept = break_nums, color = 'white', size = .25) +
  annotate("text", x = (num_spp - 10), y = 95, hjust = 1, vjust = .5, size = 3, color = 'white', 
           fontface = 'bold.italic', label = "200 m and deeper") + 
  annotate("text", x = 10, y = 5, hjust = 0, vjust = .5, size = 3, color = 'grey10', 
           fontface = 'bold.italic', label = "0 - 200 m") + 
  scale_y_continuous(# limits = c(0, 100), ### why is this causing a vertical line?
                     breaks = break_nums,
                     labels = paste0(break_nums, '%')) +
  scale_x_continuous(limits = c(0, num_spp), ### why is this causing a vertical line?
                     breaks = c(seq(0, 400, 100), num_spp)) +
  labs(y = 'IUCN % of coral range',
       x = 'Count of IUCN coral species maps')
print(coral_area_plot)

ggsave(plot = coral_area_plot,
       filename = file.path(dir_fig, 'fig3a_coral_area_plot.png'), device = 'png',
       width = 9, height = 5, units = 'cm', dpi = 600)

```

``` {r generate_coral_quadmap}

spp_coralmaps <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, d_perc, c(sm_perc_raw, sm_perc_shallow)) %>%
  gather(area, a_ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'shallow') & !(str_detect(area, 'shallow')))) %>%
  mutate(method = ifelse(str_detect(dist, 'shallow'), '200 m clip', 'original'))

write_csv(iucn_coral_fixed, file.path(dir_git, 'clip_depth', 'iucn_coral_fixed.csv'))

break_nums <- seq(0, 100, 20)

### Plot the new quad graph
clipped_quads <- ggplot(iucn_coral_fixed %>%
                      filter(!iucn_sid %in% (iucn_coral_fixed %>% 
                                               filter(d_perc > 100) %>% 
                                               .$iucn_sid)),
                    aes(x = a_ratio, y = d_perc, group = iucn_sid)) +
  ggtheme_plot + 
  ### color the quadrant backgrounds:
  annotate("rect", xmin = 54.5, xmax = 100, ymin = 63, ymax = 100,
           alpha = .3, fill= "#4dac26")  +
  annotate("rect", xmax = 54.5, xmin =   0, ymin = 63, ymax = 100,
           alpha = .3, fill= "#b8e186") +
  annotate("rect", xmin = 54.5, xmax = 100, ymax = 63, ymin =   0,
           alpha = .3, fill= "#f1b6da") +
  annotate("rect", xmax = 54.5, xmin =   0, ymax = 63, ymin =   0,
           alpha = .3, fill= "#d01c8b") +
  ### add grid lines
  geom_vline(xintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +
  geom_hline(yintercept = break_nums, size = 0.25, color = 'white', alpha = .5) +

  scale_x_continuous(expand = c(0, 0), limits = c(-1, 101),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 101),
                     breaks = break_nums,
                     labels = sprintf('%s%%', break_nums)) +

  geom_line(color = 'grey50', size = .25, alpha = .3) +
  geom_point(aes(color = method, size = method), alpha = 1) +
  scale_color_manual(values = c('original' = 'grey60', '200 m clip' = 'black')) +
  scale_size_manual(values = c('original' = .25, '200 m clip' = .25)) +
    labs(x = bquote('Area ratio '*alpha['ext']*' (%)'), 
         y = bquote('Distribution alignment '*alpha['dist']*' (%)'))

print(clipped_quads)

ggsave(plot = clipped_quads,
       filename = file.path(dir_fig, 'si_scatter_quadmap_corals.png'), device = 'png',
       width = 19, height = 9, units = 'cm', dpi = 600)

```

``` {r examine_coral_quartiles, echo = FALSE, eval = TRUE}

iucn_coral_fixed <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_fixed.csv'))

### identify quads based on original breaks from data_explore.Rmd

area_align_mean <- 54.53334
dist_align_mean <- 62.98212

# iucn_sid  am_sid  sciname  area_am  area_iucn_shallow  area_overlap_shallow  sm_range_shallow  area_iucn_raw
# area_overlap_raw  sm_range_raw  dist  perc  area  ratio  method

iucn_coral_quads <- iucn_coral_fixed %>%
  dplyr::select(iucn_sid, sciname, d_perc, a_ratio, method) %>% ### ditch am_sid, b/c duplicates
  distinct() %>%
  mutate(quad = 'q1',
         quad = ifelse(a_ratio < area_align_mean  & d_perc >= dist_align_mean, 'q2', quad),
         quad = ifelse(a_ratio >= area_align_mean & d_perc < dist_align_mean,  'q3', quad),
         quad = ifelse(a_ratio < area_align_mean  & d_perc < dist_align_mean,  'q4', quad)) %>%
  group_by(method) %>%
  mutate(n_total = n()) %>%
  group_by(method, quad) %>%
  mutate(n_quad = n(),
         pct_quad = n_quad/n_total) %>%
  ungroup()

quad_names <- data.frame('quad' = c('q1', 'q2', 'q3', 'q4'),
                         'quad_name' = factor(c('well-aligned', 'dist-aligned', 
                                                'area-aligned', 'poorly aligned'),
                                             ordered = TRUE))

iucn_coral_quads <- iucn_coral_quads %>%
  left_join(quad_names, by = 'quad')

write_csv(iucn_coral_quads, file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv'))

break_nums <- seq(0, 100, 20)

iucn_coral_quad_sum <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_quads.csv')) %>%
  dplyr::select(method, quad, quad_name, pct_quad) %>%
  distinct()
### Plot the bar chart
barchart_coral_quads <- ggplot(iucn_coral_quad_sum, 
                                aes(x = method, fill = quad, weight = pct_quad)) +
  ggtheme_plot + 
  theme(panel.grid.major.y = element_blank()) +
  geom_bar(stat = 'count', alpha = 1) +
  scale_fill_manual(values = c('q4' = '#d01c8b', 
                               'q3' = '#f1b6da', 
                               'q2' = '#b8e186',
                               'q1' = '#4dac26'),
                    labels = quad_names$quad_name) +
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1),
                     breaks = break_nums/100,
                     labels = sprintf('%s%%', break_nums)) + 
  ### add grid lines; horizontal but then get flipped
  geom_hline(yintercept = break_nums/100, size = 0.25, color = 'white', alpha = .5) +
  coord_flip() +
  labs(x = 'Depth limit', 
       y = 'Percent of corals by quadrant', 
       fill = 'Alignment')

print(barchart_coral_quads)

# ggsave(plot = barchart_coral_quads,
#        filename = file.path(dir_fig, 'barchart_coral_quads.png'),
#        width = 19, height = 4, units = 'cm', dpi = 600)

write_csv(iucn_coral_quad_sum, file.path(dir_git, 'clip_depth', 'iucn_coral_quad_sum.csv'))

```