### 2016-09-24 clip_fao/fao_clip_pac.Rmd
added t-test for mean area-per-cells

---
title: 'Data Exploration: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(ggplot2)
library(tmap)
data(World)
library(maptools)
data(wrld_simpl)

library(readr)
library(raster)
library(rgdal)
library(tidyr)
library(dplyr)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_M, 'git-annex/globalprep')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')

source(file.path(dir_git, 'data_explore/data_explore_fxns.R'))

```

- get aquamaps cells for all quadrant 3 species (and quadrant 4?)
- focus on any particular group? filter as appropriate
    - maybe by `occurcells` field? for low-data spp
- add lat/long from the hcaf file (or at least long)
- for each species:
    - identify east/west limits of range (between -180 and +180)
        - what about species who cross over in pacific? maybe a flag for pacific species?
        - maybe just select Pacific species and avoid Atlantic?
    - rescale 0-1?
        - here we'd be looking for what happens at the edges - is there a sudden jump from no presence to a bunch just at a single longitude?
        - with this, we could stack all the species and see the cumulative effects of the ends
    - plot end points, east/west bounds on a horizontal line/histogram/density plot, and put verticals on the longitudes of the FAO region boundaries
        - not even worrying about N/S? or maybe include just a big swath across the equator or whatever?

``` {r get_am_spp}

### get quadrant species list; filter to q3 and q4
spp_quad_list <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
  dplyr::select(sciname, am_sid, iucn_sid, reviewed, occurcells, quad) %>%
  filter(quad %in% c('q3', 'q4'))

nrow(spp_quad_list %>% filter(occurcells > 100))
### 110 out of 1078 (~10%) have more than 100 occurrence cells

### load full AM list and filter to those in q3 and q4
am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
                                   'v2016/int/am_cells_spp_prob0.csv')) # %>%
  # filter(am_sid %in% spp_quad_list$am_sid) %>% 
  # distinct()

am_spp_list <- read_csv(file.path(dir_git, 'data/spp_all.csv')) %>%
  filter(!is.na(am_sid)) %>%
  arrange(reviewed) %>%
  filter(!duplicated(am_sid)) %>%
  dplyr::select(am_sid, sciname, reviewed) %>%
  distinct()

message('loading hcaf and adjusting to se asia')
### join with hcaf to get longitudes (and latitudes) for each cell
hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015',
                           'csv/hcaf_truncated.csv')) %>%
  dplyr::select(-csquarecode, -cellarea, -oceanarea) %>%
  mutate(clongadj  = ifelse(centerlong < 0, centerlong + 360, centerlong),
         wlimitadj = ifelse(wlimit < 0, wlimit + 360, wlimit),
         elimitadj = ifelse(elimit < 0, elimit + 360, elimit))

message('binding hcaf to am_spp_latlong')

am_spp_latlong <- hcaf %>%
  left_join(am_spp_cells, by = 'loiczid')

```

Major fishing area boundaries
- 175W north from 25S
- 120W from 5N to 60 S
- 150E south from 25S (south of Australia)
- 80E from 0N to 55S
- 30 E from Africa on down

``` {r atlantic_spp} 
### identify (to cut) species that appear in the mid-atlantic - focus just on pacific/indopacific spp only

hcaf_atl <- hcaf %>%
  filter(centerlong >= -60 & centerlong <= 10)

am_spp_atl <- hcaf_atl %>%
  left_join(am_spp_cells, by = 'loiczid')

spp_atl <- am_spp_atl %>%
  .$am_sid %>%
  unique()
```

``` {r central_band}
### 25 S to 20 N  - look for spikes at 120W (240E), 80E (and 77E), and especially 175W (185E)
am_spp_central <- am_spp_latlong %>%
  filter(centerlat <= 20 & centerlat >= -25)

message('summarizing...')
am_spp_central_lims <- am_spp_central %>%
  group_by(am_sid) %>%
  summarize(w_lim = min(wlimitadj),
            e_lim = max(elimitadj),
            n_w_lim = sum(wlimitadj == w_lim),
            n_e_lim = sum(elimitadj == e_lim),
            n_tot = n())

message('gathering...')
am_spp_central_lims <- am_spp_central_lims %>%
  gather(lim, long, w_lim:e_lim) %>%
  gather(lim_n, ncells, n_w_lim:n_e_lim) %>%
  filter(str_detect(lim_n, lim)) %>%
  dplyr::select(-lim_n) %>%
  filter(!(long <= 70 | long >= 200))
    ### delete artificially imposed limits

write_csv(am_spp_central_lims, file.path(dir_git, 'clip_fao/am_spp_central_lims.csv'))
### NOTE: these still include spp that appear in Atlantic, so global -
### "east limit" and "west limit" are not exactly accurate
tmp <- am_spp_central_lims %>%
  group_by(long) %>%
  summarize(n_cells = sum(ncells),
            n_spp   = n())

ggplot(tmp, aes(x = long, y = n_spp)) +
  geom_point() +
  geom_vline(xintercept = c(77, 185), color = 'red')

```

``` {r what_species_are_these}
### here, the `filter(!am_sid %in% spp_atl)` line eliminates species with a border
### on 185E who also appear in the Atlantic... global species have no eastern/
### western border...
clipped_spp <- am_spp_central_lims %>%
  filter(!am_sid %in% spp_atl) %>%
  filter(long == 185) %>%
  dplyr::select(am_sid, lim) %>%
  distinct() %>%
  left_join(am_spp_list, by = 'am_sid')

write_csv(clipped_spp, file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))

```

``` {r plot_setup, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

loiczid_raster <- raster(file.path(dir_git, 'data/rasters/loiczid_raster.tif'))

fao_polys <- rgdal::readOGR(dsn = file.path(path.expand(dir_git), 'clip_fao/fao_rgns'),
                            layer = 'fao_rgns2')

rast_ext  <- extent(c(-30, 330, -63, 75)) ### was -75 but changed to -63 to cut more from bottom
fao_polys <- crop(fao_polys, rast_ext)

# wrld_simpl_shift <- wrld_simpl %>%
#   raster::shift(360)
# 
# World1 <- rgeos::gUnion(wrld_simpl, wrld_simpl_shift, byid = FALSE) %>%
#   crop(rast_ext) %>%
#   SpatialPolygonsDataFrame(data.frame(x = 1))
# 
# maptools::writePolyShape(World1, 'clip_fao/fao_rgns/world1')

World1 <- maptools::readShapePoly('clip_fao/fao_rgns/world1')

```

``` {r plot_spp_e, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))

spp_examples_e <- clipped_spp %>%
  filter(lim == 'w_lim')

message('grouping affected species and summing to loiczid')
### group affected spp by loiczid and sum

spp_cells_clipped <- am_spp_cells %>%
  filter(am_sid %in% spp_examples_e$am_sid) %>%
  group_by(loiczid) %>%
  summarize(n_spp = n())
write_csv(spp_cells_clipped, file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv'))

# spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_e.csv')) %>%
#   as.data.frame()

### create raster, then shift raster

rast_spp  <-  subs(loiczid_raster, 
                   spp_cells_clipped[ , c('loiczid', 'n_spp')], 
                   by = 'loiczid', 
                   which = 'n_spp', 
                   subsWithNA = TRUE)

rast_spp_shift <- rast_spp %>%
  raster::shift(360)

rast_spp_e <- merge(rast_spp, rast_spp_shift)

values(rast_spp_e)[values(rast_spp_e) == 0] <- NA
# values(rast_spp_e)[is.na(values(rast_spp_e))] <- 0
rast_spp_e  <- crop(rast_spp_e, rast_ext)

### boundary line
fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
  Lines(ID = 'a') %>%
  list() %>%
  SpatialLines()

rast_spp_map_e <- 
  tm_shape(rast_spp_e) +
    tm_raster(palette = 'Reds',
              breaks = c(1, 15, 100, 150, 200, Inf),
              colorNA = 'lightblue1',
              textNA = 'Zero',
              title = 'Species count',
              alpha = 1) +
  tm_shape(World1) +
    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
  tm_shape(fao_polys) +
    tm_borders(col = 'grey40', lwd = 0.25) +
    tm_text('F_AREA_1',
            col = 'royalblue4',
            fontface = 'bold.italic',
            fontfamily = 'Helvetica',
            size = .5) +
  tm_shape(fao_boundary) +
    tm_lines(col = 'blue', lwd = 1.5) +
  tm_layout(# basemaps = "Esri.WorldTopoMap", 
            frame = TRUE,
            legend.text.size = .6,
            legend.title.size = .7,
            # title.position = 'TOP', 
            legend.outside = FALSE,
            legend.position = c('right', 'top'),
            legend.bg.color = 'white',
            legend.bg.alpha = 1,
            attr.outside = TRUE)

print(rast_spp_map_e)

save_tmap(rast_spp_map_e, 
          file.path(dir_fig, 'map_am_edge.png'), 
          width = 19, 
          # height = 6, 
          units = 'cm', dpi = 600)

```

``` {r plot_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))

spp_examples_w <- clipped_spp %>%
  filter(lim == 'e_lim')

message('grouping affected species and summing to loiczid')
### group affected spp by loiczid and sum

spp_cells_clipped <- am_spp_cells %>%
  filter(am_sid %in% spp_examples_w$am_sid) %>%
  group_by(loiczid) %>%
  summarize(n_spp = n())
write_csv(spp_cells_clipped, file.path(dir_git, 'clip_fao', 'spp_cells_clipped_w.csv'))

# spp_cells_clipped <- read_csv(file.path(dir_git, 'clip_fao', 'spp_cells_clipped_w.csv')) %>%
#   as.data.frame()

### create raster, then shift raster

rast_spp  <-  subs(loiczid_raster, 
                   spp_cells_clipped[ , c('loiczid', 'n_spp')], 
                   by = 'loiczid', 
                   which = 'n_spp', 
                   subsWithNA = TRUE)

rast_spp_shift <- rast_spp %>%
  raster::shift(360)

rast_spp_w <- merge(rast_spp, rast_spp_shift)

values(rast_spp_w)[values(rast_spp_w) == 0] <- NA
# values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
rast_spp_w  <- crop(rast_spp_w, rast_ext)

### boundary line
fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
  Lines(ID = 'a') %>%
  list() %>%
  SpatialLines()

rast_spp_map_w <- 
  tm_shape(rast_spp_w) +
    tm_raster(palette = 'Reds',
              breaks = c(1, 15, 100, 1000, 2000, Inf),
              colorNA = '#ddeeff',
              textNA = 'Zero',
              title = 'Species count',
              alpha = 1) +
  tm_shape(World1) +
    tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
  tm_shape(fao_polys) +
    tm_borders(col = 'grey40', lwd = 0.25) +
    tm_text('F_AREA_1',
            col = 'royalblue4',
            fontface = 'bold.italic',
            fontfamily = 'Helvetica',
            size = .5) +
  tm_shape(fao_boundary) +
    tm_lines(col = 'blue', lwd = 1.5) +
  tm_layout(# basemaps = "Esri.WorldTopoMap", 
            frame = TRUE,
            legend.text.size = .6,
            legend.title.size = .7,
            # title.position = 'TOP', 
            legend.outside = FALSE,
            legend.position = c('right', 'top'),
            legend.bg.color = 'white',
            legend.bg.alpha = .95,
            attr.outside = TRUE)

print(rast_spp_map_w)

# save_tmap(rast_spp_map_w, 
#           file.path(dir_fig, 'map_am_edge_w.png'), 
#           width = 19, 
#           units = 'cm', dpi = 600)

```

``` {r plot_several_spp_w, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv'))

spp_examples_w <- clipped_spp %>%
  filter(lim == 'e_lim')

spp_w_cells <- am_spp_cells %>%
  filter(am_sid %in% spp_examples_w$am_sid)

spp_w_all <- spp_w_cells %>%
  group_by(am_sid) %>%
  summarize(n_cells = n())

spp_w_samples <- spp_w_all %>%
  filter(n_cells > 9000) %>% ### focus on large-range species
  sample_n(20) %>%
  .$am_sid

print(spp_w_all %>% filter(am_sid %in% spp_w_samples))
### some good candidates:
#     am_sid n_cells
# Fis-143854   23422
#  Fis-34036   13796
# URM-2-3539   16664
#  Fis-33989   15715
spp_plotted <- read_csv('data/spp_all.csv') %>%
  filter(am_sid %in% spp_w_samples) %>%
  left_join(read_csv('data/speciesoccursum.csv') %>%
              dplyr::select(am_sid = speciesid, occurcells, fbname),
            by = 'am_sid')

for(spp in spp_w_samples) {
  # spp <- spp_w_samples[1]
  message('Processing map for ', spp)
  message('... filtering cells for ', spp)
  spp_cells <- spp_w_cells %>%
    filter(am_sid == spp) %>%
    mutate(presence = 1)
  rast_spp  <-  subs(loiczid_raster, 
                     spp_cells[ , c('loiczid', 'presence')], 
                     by = 'loiczid', 
                     which = 'presence', 
                     subsWithNA = TRUE)

  message('... shifting raster and setting 0 to NA for ', spp)
  rast_spp_shift <- rast_spp %>%
    raster::shift(360)

  rast_spp_w <- merge(rast_spp, rast_spp_shift)

  values(rast_spp_w)[values(rast_spp_w) == 0] <- NA
  # values(rast_spp_w)[is.na(values(rast_spp_w))] <- 0
  rast_spp_w  <- crop(rast_spp_w, rast_ext)

  ### boundary line
  fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
    Lines(ID = 'a') %>%
    list() %>%
    SpatialLines()

  message('... generating map for ', spp)

  rast_spp_map_w <- 
    tm_shape(rast_spp_w) +
      tm_raster(palette = 'Reds',
                colorNA = '#ddeeff',
                textNA = 'not present',
                title = spp,
                alpha = 1) +
    tm_shape(World1) +
      tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
    tm_shape(fao_polys) +
      tm_borders(col = 'grey40', lwd = 0.25) +
      tm_text('F_AREA_1',
              col = 'royalblue4',
              fontface = 'bold.italic',
              fontfamily = 'Helvetica',
              size = .5) +
    tm_shape(fao_boundary) +
      tm_lines(col = 'blue', lwd = 1.5) +
    tm_layout(# basemaps = "Esri.WorldTopoMap", 
              frame = TRUE,
              legend.text.size = .6,
              legend.title.size = .7,
              # title.position = 'TOP', 
              legend.outside = FALSE,
              legend.position = c('right', 'top'),
              legend.bg.color = 'white',
              legend.bg.alpha = .95,
              attr.outside = TRUE)

  print(rast_spp_map_w)
}

# save_tmap(rast_spp_map_w, 
#           file.path(dir_fig, 'map_am_edge_w.png'), 
#           width = 19, 
#           units = 'cm', dpi = 600)

```

``` {r plot_selected_spp, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

### Select four example species
spp_ex <- c('Hoplichthys regani' = 'Fis-34036', 
            'Ipnops meadi'       = 'Fis-143854', 
            'Culeolus herdmani'  = 'URM-2-3539', 
            'Chrionema chlorotaenia' = 'Fis-33989')[1]

spp_ex_cells <- am_spp_cells %>%
  filter(am_sid %in% spp_ex)

spp_ex_info <- read_csv('data/spp_all.csv') %>%
  filter(am_sid %in% spp_ex)

for(i in 1:length(spp_ex)) {
  # i <- 1
  spp <- spp_ex[i]
  spp_name <- names(spp_ex)[i]

  message('Processing map for ', spp)

  message('... loading occurrences for ', spp)
  spp_occur <- read_csv(paste0('clip_fao/obs_for_', tolower(spp) %>% 
                                 str_replace_all('-', '_'), '.csv')) %>%
    mutate(long = ifelse(long < 0, long + 360, long),
           obs = toupper(obs),
           obs = ifelse(obs == 'BOTH', 'Both', obs),
           obs = factor(obs, levels = c('OBIS', 'GBIF', 'Both')))
  spp_occur_pts <- SpatialPointsDataFrame(spp_occur[ , 1:2], data = spp_occur[ , 3])

  message('... creating raster for ', spp)
  spp_cells <- spp_ex_cells %>%
    filter(am_sid == spp) %>%
    mutate(presence = 1)
  rast_spp  <-  subs(loiczid_raster, 
                     spp_cells[ , c('loiczid', 'presence')], 
                     by = 'loiczid', 
                     which = 'presence', 
                     subsWithNA = TRUE)

  message('... shifting raster and setting 0 to NA for ', spp)
  rast_spp_shift <- rast_spp %>%
    raster::shift(360)

  rast_spp <- merge(rast_spp, rast_spp_shift)

  values(rast_spp)[values(rast_spp) == 0] <- NA
  rast_spp  <- crop(rast_spp, rast_ext)

  ### boundary line
  fao_boundary <- Line(matrix(c(185, 185, -25, 20), ncol = 2)) %>%
    Lines(ID = 'a') %>%
    list() %>%
    SpatialLines()

  message('... generating map for ', spp)

  rast_spp_map <- 
    tm_shape(rast_spp) +
      tm_raster(palette = 'Reds',
                colorNA = '#ddeeff',
                textNA  = 'not present',
                # title   = spp_name,
                legend.show = FALSE,
                showNA  = FALSE,
                alpha   = 1) +
    tm_shape(World1) +
      tm_polygons(border.col = 'grey45', col = 'grey40', lwd = 0.25) + 
    tm_shape(fao_polys) +
      tm_borders(col = 'grey40', lwd = 0.25) +
      tm_text('F_AREA_1',
              col = 'royalblue4',
              fontface = 'bold.italic',
              fontfamily = 'Helvetica',
              size = .5) +
    tm_shape(fao_boundary) +
      tm_lines(col = 'blue', lwd = 1.5)

  rast_map <- rast_spp_map +
    tm_shape(spp_occur_pts) + 
      tm_bubbles(col = 'obs',
                 title.col = 'Occurrences ',
                 palette = c('red', 'blue', 'green'),
                 border.col = 'red',
                 border.lwd = 0.5,
                 border.alpha = 1,
                 size = 1,
                 scale = .4,
                 alpha = .3) +
      # tm_dots(col = 'red') +
    tm_layout(frame = TRUE,
              legend.text.size = .6,
              legend.title.size = .7,
              legend.outside = FALSE,
              legend.position = c('right', 'top'),
              legend.bg.color = 'white',
              legend.bg.alpha = .95,
              attr.outside = TRUE)

  print(rast_map)

  message('... saving map')
  save_tmap(rast_map,
            paste0('clip_fao/map_fao_clip_', tolower(spp_name) %>% 
                     str_replace_all(' ', '_'), '.png'),
            width = 19,
            units = 'cm', dpi = 600)

}

```

``` {r area_to_data_ratio, echo = FALSE, warning = FALSE, message = FALSE, eval = TRUE}

# am_spp_cells <- read_csv(file.path(dir_anx, 'spp_ico',
#                                    'v2016/int/am_cells_spp_prob0.csv')) 
# 
# am_spp_list <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/speciesoccursum.csv')) %>%
#   dplyr::select(am_sid = speciesid, genus, species, reviewed, occurcells) %>%
#   mutate(sciname = paste(str_trim(genus), str_trim(species), sep = ' ')) %>%
#   dplyr::select(-genus, -species) %>%
#   distinct()
# 
# hcaf <- read_csv(file.path(dir_anx, '_raw_data/aquamaps/d2015/csv/hcaf_truncated.csv')) %>%
#   dplyr::select(loiczid, cellarea, oceanarea)
# 
# am_spp_cells2 <- hcaf %>%
#   left_join(am_spp_cells, by = 'loiczid')
# 
# 
# spp_area_df <- am_spp_cells2 %>%
#   group_by(am_sid) %>%
#   summarize(cellarea = sum(cellarea),
#             oceanarea = sum(oceanarea))
# 
# spp_area_df <- spp_area_df %>%
#   inner_join(am_spp_list, by = 'am_sid') %>%
#   mutate(area_per_cells = oceanarea/occurcells)
# 
# clipped_spp <- read_csv(file.path(dir_git, 'clip_fao', 'clipped_mid_pac.csv')) %>%
#   filter(lim == 'e_lim') %>%
#   dplyr::select(am_sid)
# 
# spp_area_df <- spp_area_df %>%
#   mutate(log_area_cells = log(area_per_cells + 1),
#          clipped = am_sid %in% clipped_spp$am_sid)

# write_csv(spp_area_df, 'clip_fao/spp_area_per_cell.csv')

spp_area_df <- read_csv('clip_fao/spp_area_per_cell.csv')

spp_area_clipped <- spp_area_df %>%
  filter(clipped)

### compare ocean areas
summary(spp_area_df$oceanarea)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   #   1806    946600   2921000  10210000   7875000 501300000 
summary(spp_area_clipped$oceanarea)
   #   Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
   # 134900   4646000   7032000   8960000   9690000 111300000

### compare occurcells
summary(spp_area_df$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 0.00    6.00   16.00   57.13   49.00 7663.00 
summary(spp_area_clipped$occurcells)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 3.00    6.00   15.00   26.86   33.00  494.00

### compare area_per_cells
summary(spp_area_df %>% filter(occurcells != 0) %>% .$area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    #  190    50540   158000   485300   439100 44700000 
summary(spp_area_clipped$area_per_cells)
    # Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
    # 6578   202900   452400   827200   923300 26090000

### compare reviewed species
nrow(spp_area_df %>% filter(!str_detect(reviewed, 'null')))
   # 1296 out of 22889 = 5.66%
nrow(spp_area_clipped %>% filter(!str_detect(reviewed, 'null')))
   # 22 out of 3208 = .686%

hist(spp_area_clipped$area_per_cells)
hist(spp_area_clipped$log_area_cells)
hist(spp_area_df$log_area_cells)

t.test(x = spp_area_clipped$log_area_cells,
       y = spp_area_df %>% filter(occurcells != 0) %>% .$log_area_cells,
       alternative = 'greater') ### clipped sample mean greater than overall
# t = 49.082, df = 5244.6, p-value < 2.2e-16
# alternative hypothesis: true difference in means is greater than 0
# 95 percent confidence interval:
#  1.058579      Inf
# sample estimates:
# mean of x mean of y 
#  12.97963  11.88434

```