### 2016-09-28 ms_4_refs.Rmd
the submitted version

# Acknowledgments

We are thankful to Melanie Frazier, Julia Stewart Lowndes, Halley Froelich, and Claire Runge for their insightful comments on this manuscript, and to the National Center for Ecological Analysis and Synthesis for computation support.  We acknowledge financial support from the Gordon and Betty Moore Foundation.

-----

# References

<!-- references will be automatically added from Zotero bib when knitting the document from manuscript-main.Rmd -->