### 2017-04-11 shiny_am_iucn/data_prep.Rmd
setting up to get a better sample of species for maps

---
title: 'Data prep for Shiny app'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
  pdf_document:
    toc: true
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(maps)
library(data.table)
library(raster)
library(tidyverse)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git   <- '~/github/IUCN-AquaMaps'
dir_shiny <- file.path(dir_git, 'shiny_am_iucn')

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico/v2016')

# if(basename(getwd()) != 'data_explore') setwd('data_explore')
source(file.path(dir_git, 'data_explore', 'data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

# Set up IUCN and AquaMaps species lists and cell lists

Things to consider when choosing species to include in the app:

* The pull down menus get unwieldy when long.
* Maps with more area (i.e. more cells) look better.
* We want a good cross section of species from each quadrant.
* We want a good cross section of species that are expert reviewed?

``` {r choose_species_to_map}

### copy and trim the list of overlapping species
spp_list_quads <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
  select(sciname, iucn_sid, am_sid, 
         area_iucn, area_am,
         dist_align = sm_perc, area_ratio, 
         spp_group_text, 
         reviewed, occurcells, 
         n_spp, quad) %>%
  mutate(reviewed = ifelse(!is.na(reviewed), TRUE, FALSE))
write_csv(spp_list_quads, file.path(dir_shiny, 'data', 'spp_list_quads_app.csv'))

### copy the barchart files too
file.copy(file.path(dir_git, 'data', 'spp_gp_quads.csv'), file.path(dir_shiny, 'data', 'spp_gp_quads_app.csv'))
file.copy(file.path(dir_git, 'data', 'spp_gp_quads_ex.csv'), file.path(dir_shiny, 'data', 'spp_gp_quads_ex_app.csv'))

### Because we're using newer cell data (2016), and the quads list is based
### on 2015 data, cut any spp whose IDs have changed or otherwise gone awol.
if(!exists('iucn_cells')) {
  iucn_cells <- read_csv(file.path(dir_anx, 'int', 'iucn_cells_spp.csv'), col_types = 'ciid__')
}

if(!exists('am_cells')) {
  am_cells <- read_csv(file.path(dir_anx, 'int', 'am_cells_spp_prob0.csv'), col_types = 'cd')
}

spp_list_quads2016 <- spp_list_quads %>%
  filter(iucn_sid %in% iucn_cells$iucn_sid & am_sid %in% am_cells$am_sid)

### determine species list from IUCN area and AquaMaps area, as well as quadrant.
### Knock out anything below first quartile for both; also knock out the really huge ones?
# summary(spp_list_quads$area_iucn)
     # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
     # 8383   1308000   4427000  13050000  16260000 355400000 
# summary(spp_list_quads$area_am)
     # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
     # 2940   1425000   4495000  10960000  10100000 364700000
spp_list_med_areas <- spp_list_quads2016 %>%
  filter(area_iucn > quantile(area_iucn, .25)) %>%
  filter(area_am > quantile(area_am, .25)) %>%
  filter(area_am < quantile(area_am, .75) | area_iucn < quantile(area_iucn, .75))

### Start with a sample of 15% of the medium-area species; this is about 170 spp.
### Then group by species group and quadrant, and for each, ensure a minimum
### representation by selecting the first two (or one, if group is only 1)
spp_list_samples <- spp_list_med_areas %>%
  mutate(keep_spp = ifelse(runif(n()) <= .15, TRUE, FALSE)) %>%
  group_by(spp_group_text, quad) %>%
  mutate(n_spp_gp = n(),
         keep_spp = ifelse(iucn_sid == first(iucn_sid), TRUE, keep_spp),
         keep_spp = ifelse(iucn_sid == nth(iucn_sid, 2) & n_spp_gp >= 2, TRUE, keep_spp),
         kept     = sum(keep_spp)) %>%
  ungroup() %>%
  filter(keep_spp) %>%
  select(iucn_sid, am_sid, sciname, spp_group_text) %>%
  distinct()

write_csv(spp_list_samples, file.path(dir_shiny, 'data', 'spp_list_app.csv'))

```

``` {r prep_cells_files}

spp_list <- read_csv(file.path(dir_shiny, 'data', 'spp_list_app_stable.csv'))

### read IUCN cells from spp_ico/v2016, trim to just overlapping species cells
if(!exists('iucn_cells')) {
  iucn_cells <- read_csv(file.path(dir_anx, 'int', 'iucn_cells_spp.csv'), col_types = 'ciid__')
}

iucn_cells_map <- iucn_cells %>%
  filter(presence != 5) %>%
  filter(iucn_sid %in% spp_list$iucn_sid) %>%
  select(-sciname, -presence) %>%
  distinct()

write_csv(iucn_cells_map, file.path(dir_shiny, 'data', 'iucn_cells_app.csv'))

if(!exists('am_cells')) {
  am_cells <- read_csv(file.path(dir_anx, 'int', 'am_cells_spp_prob0.csv'), col_types = 'cd')
}

am_cells_map <- am_cells %>%
  filter(am_sid %in% spp_list$am_sid) %>%
  distinct()

write_csv(am_cells_map, file.path(dir_shiny, 'data', 'am_cells_app.csv'))

```

## Coral species

Get coral species file and bring it over here; clean it up a bit for easier processing.

``` {r get_coral_spp_info}
spp_coralmaps <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, dist_align, c(sm_perc_raw, sm_perc_shallow)) %>%
  gather(area, area_ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'shallow') & !(str_detect(area, 'shallow')))) %>%
  mutate(method = ifelse(str_detect(dist, 'shallow'), '200 m clip', 'all depth')) %>%
  select(iucn_sid, am_sid, sciname, dist_align, area_ratio, method) %>%
  distinct()

write_csv(iucn_coral_fixed, file.path(dir_shiny, 'data', 'coral_spp_areas_app.csv'))

```

***

``` {r prov_footer, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```