### 2017-04-14 shiny_am_iucn/data_prep.Rmd
added coral maps functionality and lots of explanatory text

---
title: 'Data prep for Shiny app'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
  pdf_document:
    toc: true
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(maps)
library(data.table)
library(raster)
library(tidyverse)
library(stringr)

dir_M <- c('Windows' = '//mazu.nceas.ucsb.edu/ohi',
           'Darwin'  = '/Volumes/ohi',    ### connect (cmd-K) to smb://mazu/ohi
           'Linux'   = '/home/shares/ohi')[[ Sys.info()[['sysname']] ]]

dir_git   <- '~/github/IUCN-AquaMaps' %>% path.expand()
dir_shiny <- file.path(dir_git, 'shiny_am_iucn')

dir_anx <- file.path(dir_M, 'git-annex/globalprep/spp_ico/v2016')

# if(basename(getwd()) != 'data_explore') setwd('data_explore')
source(file.path(dir_git, 'data_explore', 'data_explore_fxns.R'))
### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

# Set up IUCN and AquaMaps species lists and cell lists

Things to consider when choosing species to include in the app:

* The pull down menus get unwieldy when long.
* Maps with more area (i.e. more cells) look better.
* We want a good cross section of species from each quadrant.
* We want a good cross section of species that are expert reviewed?

``` {r choose_species_to_map}

### copy and trim the list of overlapping species
spp_list_quads <- read_csv(file.path(dir_git, 'data', 'spp_list_quads.csv')) %>%
  select(sciname, iucn_sid, am_sid, 
         area_iucn, area_am,
         dist_align = sm_perc, area_ratio, 
         spp_group_text, 
         reviewed, occurcells, 
         n_spp, quad) %>%
  mutate(reviewed = ifelse(!is.na(reviewed), TRUE, FALSE))
write_csv(spp_list_quads, file.path(dir_shiny, 'data', 'spp_list_quads_app.csv'))

### copy the barchart files too
file.copy(file.path(dir_git, 'data', 'spp_gp_quads.csv'), file.path(dir_shiny, 'data', 'spp_gp_quads_app.csv'))
file.copy(file.path(dir_git, 'data', 'spp_gp_quads_ex.csv'), file.path(dir_shiny, 'data', 'spp_gp_quads_ex_app.csv'))

### Because we're using newer cell data (2016), and the quads list is based
### on 2015 data, cut any spp whose IDs have changed or otherwise gone awol.
if(!exists('iucn_cells')) {
  iucn_cells <- read_csv(file.path(dir_anx, 'int', 'iucn_cells_spp.csv'), col_types = 'ciid__')
}

if(!exists('am_cells')) {
  am_cells <- read_csv(file.path(dir_anx, 'int', 'am_cells_spp_prob0.csv'), col_types = 'cd')
}

spp_list_quads2016 <- spp_list_quads %>%
  filter(iucn_sid %in% iucn_cells$iucn_sid & am_sid %in% am_cells$am_sid)

### determine species list from IUCN area and AquaMaps area, as well as quadrant.
### Knock out anything below first quartile for both; also knock out the really huge ones?
# summary(spp_list_quads$area_iucn)
     # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
     # 8383   1308000   4427000  13050000  16260000 355400000 
# summary(spp_list_quads$area_am)
     # Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
     # 2940   1425000   4495000  10960000  10100000 364700000
spp_list_med_areas <- spp_list_quads2016 %>%
  filter(area_iucn > quantile(area_iucn, .25)) %>%
  filter(area_am > quantile(area_am, .25)) %>%
  filter(area_am < quantile(area_am, .75) | area_iucn < quantile(area_iucn, .75))

### Start with a sample of 15% of the medium-area species; this is about 170 spp.
### Then group by species group and quadrant, and for each, ensure a minimum
### representation by selecting the first two (or one, if group is only 1)
spp_list_samples <- spp_list_med_areas %>%
  mutate(keep_spp = ifelse(runif(n()) <= .15, TRUE, FALSE)) %>%
  group_by(spp_group_text, quad) %>%
  mutate(n_spp_gp = n(),
         keep_spp = ifelse(iucn_sid == first(iucn_sid), TRUE, keep_spp),
         keep_spp = ifelse(iucn_sid == nth(iucn_sid, 2) & n_spp_gp >= 2, TRUE, keep_spp),
         kept     = sum(keep_spp)) %>%
  ungroup() %>%
  filter(keep_spp) %>%
  select(iucn_sid, am_sid, sciname, spp_group_text) %>%
  distinct()

write_csv(spp_list_samples, file.path(dir_shiny, 'data', 'spp_list_app.csv'))

```

``` {r prep_cells_files}

spp_list <- read_csv(file.path(dir_shiny, 'data', 'spp_list_app_stable.csv'))

### read IUCN cells from spp_ico/v2016, trim to just overlapping species cells
if(!exists('iucn_cells')) {
  iucn_cells <- read_csv(file.path(dir_anx, 'int', 'iucn_cells_spp.csv'), col_types = 'ciid__')
}

iucn_cells_map <- iucn_cells %>%
  filter(presence != 5) %>%
  filter(iucn_sid %in% spp_list$iucn_sid) %>%
  select(-sciname, -presence) %>%
  distinct()

write_csv(iucn_cells_map, file.path(dir_shiny, 'data', 'iucn_cells_app.csv'))

if(!exists('am_cells')) {
  am_cells <- read_csv(file.path(dir_anx, 'int', 'am_cells_spp_prob0.csv'), col_types = 'cd')
}

am_cells_map <- am_cells %>%
  filter(am_sid %in% spp_list$am_sid) %>%
  distinct()

write_csv(am_cells_map, file.path(dir_shiny, 'data', 'am_cells_app.csv'))

```

## Get FAO regions

``` {r copy_fao_rgns}

# fao_files <- list.files(file.path(dir_git, 'clip_fao/fao_rgns'), pattern = 'fao_rgns')
# 
# file.copy(from = file.path(dir_git, 'clip_fao/fao_rgns', fao_files),
#           to   = file.path(dir_shiny, 'data', fao_files %>%
#                              str_replace('\\.', '_app.')))
# fao_rgn <- rgdal::readOGR(dsn = 'data', layer = "fao_rgns2_app") %>%
#   crop(extent(-180, 180, -90, 90))
# rgdal::writeOGR(fao_rgn, dsn = 'data', layer = "fao_rgns2_app", driver = 'ESRI Shapefile', overwrite.layer = TRUE)

```

## Coral species

Get coral species alignment file and bring it over here; clean it up a bit for easier processing.

``` {r get_coral_spp_info}
spp_coralmaps <- read_csv(file.path(dir_git, 'clip_depth', 'coral_spp_areas.csv'))

### recalc areas and overlaps
iucn_coral_fixed <- spp_coralmaps %>%
  gather(dist, dist_align, c(sm_perc_raw, sm_perc_shallow)) %>%
  gather(area, area_ratio, c(area_ratio_raw, area_ratio_shallow)) %>%
  filter(!(str_detect(dist, 'raw') & !(str_detect(area, 'raw')))) %>%
  filter(!(str_detect(dist, 'shallow') & !(str_detect(area, 'shallow')))) %>%
  mutate(method = ifelse(str_detect(dist, 'shallow'), '200 m clip', 'all depth')) %>%
  select(iucn_sid, am_sid, sciname, dist_align, area_ratio, method) %>%
  distinct()

write_csv(iucn_coral_fixed, file.path(dir_shiny, 'data', 'coral_align_app.csv'))

```

### Get coral cells and info for mapping

``` {r get_coral_info}

### load iucn spp_cells from main analysis; clip to just those in app
spp_list <- read_csv(file.path(dir_shiny, 'data', 'spp_list_app_stable.csv'))

coral_cells_app <- read_csv(file.path(dir_git, 'clip_depth', 'coral_cells_iucn.csv')) %>%
  filter(iucn_sid %in% spp_list$iucn_sid)
# length(unique(coral_cells_app$iucn_sid))
write_csv(coral_cells_app, file.path(dir_shiny, 'data', 'coral_cells_app.csv'))

coral_spp_list  <- spp_list %>%
  select(iucn_sid, sciname) %>%
  filter(iucn_sid %in% coral_cells_app$iucn_sid) %>%
  distinct()
write_csv(spp_coral_maps, file.path(dir_shiny, 'data', 'coral_spp_list.csv'))

coral_quad_sum <- read_csv(file.path(dir_git, 'clip_depth', 'iucn_coral_quad_sum.csv'))
write_csv(coral_quad_sum, file.path(dir_shiny, 'data', 'coral_quads_app.csv'))

```

``` {r simplify_bathymetry}
library(maptools)
library(tmap)

bathy_shp_orig <- rgdal::readOGR(dsn = file.path(dir_git, 'clip_depth/ne_10m_bathymetry_K_200'),
                            layer = 'ne_10m_bathymetry_K_200') %>%
  rgeos::gUnaryUnion()

keep_df <- data.frame(
  id   = 1:length(bathy_shp_orig@polygons[[1]]@Polygons),
  area = lapply(bathy_shp_orig@polygons, function(x) sapply(x@Polygons, function(y) y@area)) %>%
    unlist(),
  hole = lapply(bathy_shp_orig@polygons, function(x) sapply(x@Polygons, function(y) y@hole)) %>%
    unlist()
)
quantile(keep_df$area)
#           0%          25%          50%          75%         100% 
# 2.319330e-10 2.026033e-02 3.717621e-02 9.453268e-02 4.108813e+04

### ditch the small polys.  Tolerance? 6e-2 (.06) is .25° x .25°; can go bigger,
### perhaps up to a single .5° cell, or .25?
hole_thresh <- .0625
feat_thresh <- 2

keep_df1 <- keep_df %>%
  mutate(keep = TRUE,
         # keep = ifelse(area < hole_thresh, FALSE, keep))
         keep = ifelse(hole == TRUE  & area < hole_thresh, FALSE, keep),
         keep = ifelse(hole == FALSE & area < feat_thresh, FALSE, keep))

keep_ids <- keep_df1$id[keep_df1$keep]

bathy_shp <- bathy_shp_orig

bathy_shp@polygons[[1]]@Polygons <- bathy_shp@polygons[[1]]@Polygons[keep_ids]
bathy_shp@polygons[[1]]@plotOrder <- 1:length(bathy_shp@polygons[[1]]@Polygons)
bathy_shp@plotOrder <- 1:length(bathy_shp@polygons)

### need to repair comments... ugh
slot(bathy_shp, "polygons") <- lapply(slot(bathy_shp, "polygons"), checkPolygonsHoles)

# # will assign comments, seen by: 
# lapply(slot(bathy_shp, "polygons"), comment) 
# 
# # or remove them by: 
# slot(bathy_shp, "polygons") <- lapply(slot(bathy_shp, "polygons"), 
#    "comment<-", NULL)

### convert to spatial polygons dataframe
bathy_shp1 <- bathy_shp %>%
  SpatialPolygonsDataFrame(data = data.frame('depth' = 200))

### check new map against original map
tm_shape(bathy_shp_orig) + 
  tm_polygons(col = 'blue', lwd = 0) + 
  tm_shape(bathy_shp1) + 
  tm_polygons(col = 'red', 
              border.col = 'red3', 
              lwd = 0, alpha = .8)

writePolyShape(bathy_shp1,
               file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200/bathy_app_simple'))
file.copy(from = file.path(dir_git, 'clip_depth/ne_10m_bathymetry_K_200/ne_10m_bathymetry_K_200.prj'),
          to   = file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200/bathy_app_simple.prj'),
          overwrite = FALSE)
# rgdal::writeOGR(bathy_shp,
#                 dsn    = path.expand(file.path(dir_shiny, 'data/ne_10m_bathymetry_K_200')),
#                 layer  = 'bathy_app_simple',
#                 driver = 'ESRI Shapefile',
#                 overwrite_layer = TRUE)

```

***

``` {r prov_footer, results = 'asis'}
prov_wrapup(commit_outputs = FALSE)
```