### 2017-04-21 case_study_mpa/x_mpa_gap_in_time.Rmd
initialize dummy branch

---
title: 'MPA Gap Analysis: IUCN and AquaMaps'
output: html_document
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

# Libraries and Paths

library(raster)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)

dir_N <- c('Windows' = '//neptune.nceas.ucsb.edu/data_edit',
           'Darwin'  = '/Volumes/data_edit',
           'Linux'   = '/var/data/ohi')[[ Sys.info()[['sysname']] ]]

dir_git <- '~/github/IUCN-AquaMaps'

dir_anx <- file.path(dir_N, 'git-annex/globalprep/SPP_ICO')
dir_fig <- file.path(dir_git, 'figures')
dir_data <- file.path(dir_git, 'data')
dir_am_data <- file.path(dir_N, 'git-annex/globalprep/_raw_data/aquamaps/v2015/csv')

if(basename(getwd()) != 'IUCN-AquaMaps') setwd(dir_git)

### When knitting this, it automatically sets WD to be this directory...
### the 'setwd()' is there for running by chunk
```

``` {r Analysis using WDPA 2015 and AM 2015, echo = FALSE, eval = FALSE}
### Use cell_df_des to capture only DESIGNATED protected areas.

area_mpa <- cell_df_des %>%
  summarize(marine_area  = sum(marine_area), 
            mpa_tot_1990 = sum(y1990_area), 
            mpa_tot_1995 = sum(y1995_area), 
            mpa_tot_2000 = sum(y2000_area), 
            mpa_tot_2005 = sum(y2005_area), 
            mpa_tot_2010 = sum(y2010_area), 
            mpa_tot_2014 = sum(y2014_area), 
            mpa_tot_2015 = sum(y2015_area))

### Assemble the AM species-cell df with info on marine area and 
### protected area per cell
mpa_spp_2015 <- am_spp_cells %>%
  left_join(cell_df_des, by = 'loiczid')

### filter to the appropriate threshold...
thresh <- 0.50

mpa_spp_2015 <- mpa_spp_2015 %>%
  filter(probability >= thresh)

### group by speciesid, then summarize sum(marine area), 
### sum(protected area), ratio
mpa_spp_2015_sum <- mpa_spp_2015 %>%
  group_by(speciesid) %>%
  summarize(cellarea_tot = sum(cellarea), 
            ocean_range  = sum(oceanarea), 
            marine_range = sum(marine_area),
            mpa1990 = sum(y1990_area),
            mpa1995 = sum(y1995_area),
            mpa2000 = sum(y2000_area),
            mpa2005 = sum(y2005_area),
            mpa2010 = sum(y2010_area),
            mpa2014 = sum(y2014_area),
            mpa2015 = sum(y2015_area),
            pct1990 = mpa1990/marine_range,
            pct1995 = mpa1995/marine_range,
            pct2000 = mpa2000/marine_range,
            pct2005 = mpa2005/marine_range,
            pct2010 = mpa2010/marine_range,
            pct2014 = mpa2014/marine_range,
            pct2015 = mpa2015/marine_range)

mpa_spp_2015_sum <- mpa_spp_2015_sum %>%
  mutate(cat2015 = ifelse(pct2015 == 0,                       'gap', NA),
         cat2015 = ifelse(pct2015 > 0    & pct2015 <= 0.02,  '0-2%', cat2015),
         cat2015 = ifelse(pct2015 > 0.02 & pct2015 <= 0.05,  '2-5%', cat2015),
         cat2015 = ifelse(pct2015 > 0.05 & pct2015 <= 0.10, '5-10%', cat2015),
         cat2015 = ifelse(pct2015 > 0.10,                    '>10%', cat2015))

mpa_spp_2015_sum2 <- mpa_spp_2015_sum %>%
  group_by(cat2015) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2015 = factor(cat2015, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2015)

knitr::kable(mpa_spp_2015_sum2)

# |cat2015 | n_tot| n_cat| pct_cat|
# |:-------|-----:|-----:|-------:|
# |gap     | 22885|   802|     3.5|
# |0-2%    | 22885| 11035|    48.2|
# |2-5%    | 22885|  8703|    38.0|
# |5-10%   | 22885|  1656|     7.2|
# |>10%    | 22885|   689|     3.0|

```

| study            | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :--------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015 | IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, wdpa2014 (recreate Klein) | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| am2015, wdpa2014 (update Klein)   | IUCN I-IV, Sp. Mar |   2,915,745  | 1.7% | 45.7% | 42.7% |  7.2% | 2.7% |
| am2015, wdpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 3.5% | 48.2% | 38.0% |  7.2% | 3.0% |

So: for analysis, just keep using the Klein WDPA shapefile... avoid that particular variable!

``` {r check results for 1990 through 2010 at 50%, echo = FALSE, eval = FALSE}

### Save this for the time series analysis...

mpa_spp_2015_sum <- mpa_spp_2015_sum %>%
  mutate(cat1990 = ifelse(pct1990 == 0,                       'gap', NA),
         cat1990 = ifelse(pct1990 > 0    & pct1990 <= 0.02,  '0-2%', cat1990),
         cat1990 = ifelse(pct1990 > 0.02 & pct1990 <= 0.05,  '2-5%', cat1990),
         cat1990 = ifelse(pct1990 > 0.05 & pct1990 <= 0.10, '5-10%', cat1990),
         cat1990 = ifelse(pct1990 > 0.10,                    '>10%', cat1990),
         cat1995 = ifelse(pct1995 == 0,                       'gap', NA),
         cat1995 = ifelse(pct1995 > 0    & pct1995 <= 0.02,  '0-2%', cat1995),
         cat1995 = ifelse(pct1995 > 0.02 & pct1995 <= 0.05,  '2-5%', cat1995),
         cat1995 = ifelse(pct1995 > 0.05 & pct1995 <= 0.10, '5-10%', cat1995),
         cat1995 = ifelse(pct1995 > 0.10,                    '>10%', cat1995),
         cat2000 = ifelse(pct2000 == 0,                       'gap', NA),
         cat2000 = ifelse(pct2000 > 0    & pct2000 <= 0.02,  '0-2%', cat2000),
         cat2000 = ifelse(pct2000 > 0.02 & pct2000 <= 0.05,  '2-5%', cat2000),
         cat2000 = ifelse(pct2000 > 0.05 & pct2000 <= 0.10, '5-10%', cat2000),
         cat2000 = ifelse(pct2000 > 0.10,                    '>10%', cat2000),
         cat2005 = ifelse(pct2005 == 0,                       'gap', NA),
         cat2005 = ifelse(pct2005 > 0    & pct2005 <= 0.02,  '0-2%', cat2005),
         cat2005 = ifelse(pct2005 > 0.02 & pct2005 <= 0.05,  '2-5%', cat2005),
         cat2005 = ifelse(pct2005 > 0.05 & pct2005 <= 0.10, '5-10%', cat2005),
         cat2005 = ifelse(pct2005 > 0.10,                    '>10%', cat2005),
         cat2010 = ifelse(pct2010 == 0,                       'gap', NA),
         cat2010 = ifelse(pct2010 > 0    & pct2010 <= 0.02,  '0-2%', cat2010),
         cat2010 = ifelse(pct2010 > 0.02 & pct2010 <= 0.05,  '2-5%', cat2010),
         cat2010 = ifelse(pct2010 > 0.05 & pct2010 <= 0.10, '5-10%', cat2010),
         cat2010 = ifelse(pct2010 > 0.10,                    '>10%', cat2010),
         cat2014 = ifelse(pct2014 == 0,                       'gap', NA),
         cat2014 = ifelse(pct2014 > 0    & pct2014 <= 0.02,  '0-2%', cat2014),
         cat2014 = ifelse(pct2014 > 0.02 & pct2014 <= 0.05,  '2-5%', cat2014),
         cat2014 = ifelse(pct2014 > 0.05 & pct2014 <= 0.10, '5-10%', cat2014),
         cat2014 = ifelse(pct2014 > 0.10,                    '>10%', cat2014))

mpa_spp_1990 <- mpa_spp_2015_sum %>%
  group_by(cat1990) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat1990 = factor(cat1990, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat1990)

mpa_spp_1995 <- mpa_spp_2015_sum %>%
  group_by(cat1995) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat1995 = factor(cat1995, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat1995)

mpa_spp_2000 <- mpa_spp_2015_sum %>%
  group_by(cat2000) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2000 = factor(cat2000, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2000)

mpa_spp_2005 <- mpa_spp_2015_sum %>%
  group_by(cat2005) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2005 = factor(cat2005, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2005)

mpa_spp_2010 <- mpa_spp_2015_sum %>%
  group_by(cat2010) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2010 = factor(cat2010, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2010)

mpa_spp_2014 <- mpa_spp_2015_sum %>%
  group_by(cat2014) %>%
  summarize(n_tot   = nrow(mpa_spp_2015_sum), 
            n_cat   = n(), 
            pct_cat = round(n_cat/n_tot * 100, 1)) %>%
  mutate(cat2014 = factor(cat2014, levels = c('gap', '0-2%', '2-5%', '5-10%', '>10%'))) %>%
  arrange(cat2014)

#   cat1990 n_tot n_cat pct_cat
#       gap 22889   925     4.0
#      0-2% 22889 21569    94.2
#      2-5% 22889   376     1.6
#     5-10% 22889    18     0.1
#      >10% 22889     1     0.0
# 
#   cat1995 n_tot n_cat pct_cat
#       gap 22889   878     3.8
#      0-2% 22889 21111    92.2
#      2-5% 22889   817     3.6
#     5-10% 22889    70     0.3
#      >10% 22889    13     0.1
#
#   cat2000 n_tot n_cat pct_cat
#       gap 22889   801     3.5
#      0-2% 22889 20973    91.6
#      2-5% 22889  1004     4.4
#     5-10% 22889    92     0.4
#      >10% 22889    19     0.1
#
#   cat2005 n_tot n_cat pct_cat
#       gap 22889   608     2.7
#      0-2% 22889 19828    86.6
#      2-5% 22889  2174     9.5
#     5-10% 22889   205     0.9
#      >10% 22889    74     0.3
# 
#   cat2010 n_tot n_cat pct_cat
#       gap 22889   578     2.5
#      0-2% 22889 18670    81.6
#      2-5% 22889  3278    14.3
#     5-10% 22889   288     1.3
#      >10% 22889    75     0.3

argh <- am_spp_mpas_filt %>%
  select(-(mpa1990:mpa2015)) %>%
  gather(key = year, value = pct, pct1990:pct2015) %>%
  mutate(year = str_replace(year, 'pct', '')) %>%
  group_by(year)

hist_plot <- ggplot(data = argh, aes(x = pct, color = year)) + 
  geom_density() +
  scale_x_continuous(limits = c(0, .05))
#  scale_x_log10()

hist_plot
ggsave(file.path(dir_git, 'figures/mpa_analysis_dist.png'))
```

### Original study (Klein 2015) vs new analyses

| study           | Protected areas    | Area (km^2^) | Gap  |  0–2% |  2–5% | 5–10% | >10% |
| :-------------: | :----------------- | :----------: | :--: | :---: | :---: | :---: | :--: |
| Klein et al 2015| IUCN I-IV, Sp. Mar |   2,967,898  | 1.4% | 42.3% | 46.8% |  6.9% | 2.6% |
| am2014, mpa2014 | IUCN I-IV, Sp. Mar |   2,915,745  | 1.6% | 44.6% | 44.6% |  6.8% | 2.4% |
| .               | .                  | .            | .    | .     | .     | .     | .    |
| am2015, mpa2015 | IUCN I-IV, Sp. Mar |   2,889,758  | 2.4% | 50.3% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2014 | IUCN I-IV, Sp. Mar |   2,885,465  | 2.4% | 50.4% | 39.1% |  6.4% | 1.7% |
| am2015, mpa2010 | IUCN I-IV, Sp. Mar |   1,137,142  | 2.5% | 81.6% | 14.3% |  1.3% | 0.3% |
| am2015, mpa2005 | IUCN I-IV, Sp. Mar |     932,817  | 2.7% | 86.6% |  9.5% |  0.9% | 0.3% |
| am2015, mpa2000 | IUCN I-IV, Sp. Mar |     748,172  | 3.5% | 91.6% |  4.4% |  0.4% | 0.1% |
| am2015, mpa1995 | IUCN I-IV, Sp. Mar |     532,098  | 3.8% | 92.2% |  3.6% |  0.3% | 0.1% |
| am2015, mpa1990 | IUCN I-IV, Sp. Mar |     351,592  | 4.0% | 94.2% |  1.6% |  0.1% | 0.0% |

max protected species = Fis-160882